\chapter{Martin-L\"of's Dependent Type Theory}
\label{chap:type-theory}%

Dependent type theory is a formal system to organize all mathematical objects, structure, and knowledge. Dependent type theory is about types, or more generally dependent types, and their elements. There are many ways to think about type theory, types, and its elements. Types can be interpreted as sets, i.e., there is an interpretation of type theory into Zermelo-Fraenkel set theory, but there are some important differences between type theory and set theory, and the interpretation of types as sets has significant limitations. One of the differences is that in type theory, every element comes equipped with its type. We will write $a:A$ for the judgment that $a$ is an element of type $A$. This leads us to a second important difference between type theory and set theory. Set theory is axiomatized in the formal system of first order logic, whereas type theory is its own formal system. Types and their elements are constructed by following the rules of this formal system, and the only way to construct an element is to construct it as an element of a previously constructed type. The expression $a:A$ is therefore not considered to be a proposition, i.e., something which one can assert about an arbitrary element and an arbitrary type, but it is considered to be a judgment, i.e., an assessment that is part of the construction of the element $a:A$.

In type theory there is a much stronger focus on equality of elements than there is in set theory. It is said that a type is not fully understood until (i) one understands how to construct an element of the type and (ii) one understands precisely how to show that two elements of the type are equal. Equality in type theory is governed by the identity type. Unlike in classical set theory, where equality is a decidable proposition of first order logic, the \emph{type} $x=y$ of identifications of two elements $x,y:A$ is itself a type, and therefore it could possess intricate further structure. 

Dependent type theory is built up in several stages. At the first stage we give structural rules, which express the general theory of type dependency. There is no ambient deductive system of first order logic in type theory. Type theory is its own deductive system, and the structural rules are at the heart of this system. The basic operations that are governed by the structural rules are substitution and weakening operations. After the general system of dependent type theory has been set up, we introduce the ways in which we can form types. The most fundamental class of types are dependent function types, or $\Pi$-types. They are used for practically everything. Next, we introduce the type of natural numbers, where we use type-dependency to formulate a type-theoretic version of the induction principle. By the type-theoretic nature of this induction principle, it can be used in two ways: it can be used to construct the many familiar operations on $\N$, such as addition and multiplication, and it can also be used to prove properties about those operations.

The next idea is that we can consider induction principles for many other types as well. This leads to the idea of more general inductive types. In \cref{sec:inductive} we introduce the unit type, the empty type, the booleans, coproducts, dependent pair types, and cartesian products. All of these are examples of inductive types, and their induction principles can be used to construct the basic operations on them, as well as to prove properties about those operations.

Then we come to the most characteristic ingredient of Martin L\"of's dependent type theory: the identity type. The identity type $x=_Ay$ is an example of a \emph{dependent} type, because it is indexed by $x,y:A$, and it is inductively generated by the reflexivity element $\refl{x}:x=_Ax$. The catch is, however, that the identity type $x=_Ay$ is just another type, and it could potentially have many different elements.

The last class of types that we introduce are universes. Universes are type families that are closed under the operations of type theory: $\Pi$-types, $\Sigma$-types, identity types, and so on. Universes play a fundamental role in the theory. One important reason for introducing universes is that they can be used to define type families over inductive types via their induction principles. For example, this allows us to define the ordering relations $\leq$ and $<$ on the natural numbers. We will also use the universes to show the Peano axioms asserting that $\succN$ is injective, and that $\zeroN$ is not a successor.

In the final two sections of this chapter, we start developing mathematics in type theory. In \cref{sec:modular-arithmetic} we study the Curry-Howard interpretation, and use it to develop modular arithmetic in type theory. In \cref{sec:decidability} we study the concept of decidability, and use it to obtain basic theorems in elementary number theory, such as the well-ordering theorem, the construction of the greatest common divisor, and the infinitude of primes. Both of these sections can be viewed as tutorials in type theory, designed to give you some practical experience with type theory before diving into the intricacies of the univalent foundations of mathematics.

\input{dtt}
\input{pi}
\input{nat}
\input{inductive}
\input{identity}
\input{universes}
\input{modular-arithmetic}
\input{number-theory}
%\input{integers}

\cleardoublepage
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "hott-intro"
%%% End:
