\documentclass[12pt]{article}

\usepackage[colorlinks=true
,breaklinks=true
,urlcolor=blue
,anchorcolor=blue
,citecolor=blue
,filecolor=blue
,linkcolor=blue
,menucolor=blue
,linktocpage=true]{hyperref}
\hypersetup{
bookmarksopen=true,
bookmarksnumbered=true,
bookmarksopenlevel=10
}



\input{preamble}

\input{macros}

\usepackage[noBBpl,sc]{mathpazo}
\linespread{1.05}
\usepackage[papersize={6.9in, 10.0in}, left=.5in, right=.5in, top=1in, bottom=.9in]{geometry}
\sloppy
\raggedbottom
\usepackage[small]{titlesec}

% these include amsmath and that can cause trouble in older docs.
\input{../helpers/cmrsum}
\input{../helpers/fix-underbrace.tex}
\begin{document}

\title{A general definition of dependent type theories}

\author{Andrej Bauer
  \and Philipp G.~Haselwarter
  \and Peter LeFanu Lumsdaine}

\date{September 11, 2020}

\maketitle

\extrafootertext{The authors benefited greatly from visits funded by the COST Action \href{https://eutypes.cs.ru.nl}{\emph{EUTypes}} CA15123.}
\extrafootertext{This material is based upon work supported by the U.S.~Air Force Office of Scientific Research under award number FA9550-17-1-0326, grant number 12595060.}


\begin{abstract}
  We define a general class of dependent type theories, encompassing Martin-Löf’s intuitionistic type theories and variants and extensions.
  %
  The primary aim is pragmatic: to unify and organise their study, allowing results and constructions to be given in reasonable generality, rather than just for specific theories.
  %
  Compared to other approaches, our definition stays closer to the direct or naïve reading of syntax, yielding the traditional presentations of specific theories as closely as possible.
  
  Specifically, we give three main definitions: \emph{raw type theories}, a minimal setup for discussing dependently typed derivability; \emph{acceptable type theories}, including extra conditions ensuring well-behavedness; and \emph{well-presented type theories}, generalising how in traditional presentations, the well-behavedness of a type theory is established step by step as the type theory is built up.
  %
  Following these, we show that various fundamental fitness-for-purpose metatheorems hold in this generality.

  Much of the present work has been formalised in the proof assistant Coq.
\end{abstract}

\newpage
\tableofcontents
\newpage

\newpage

\input{introduction.tex}
\input{preliminaries.tex}
\input{raw-syntax.tex}
\input{raw-type-theories.tex}
\input{metatheorems.tex}
\input{well-founded-theories.tex}
\input{discussion.tex}

\bibliographystyle{plainnat}
\bibliography{general-type-theories.bib}

\appendix

\input{formalisation}
% \input{app-sequential-contexts.tex}



\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
