\documentclass[prl,onecolumn,superscriptaddress,12pt]{revtex4-2}

\usepackage{graphics}
\usepackage{graphicx}
\usepackage{amsthm} % for theorems
\usepackage{amsbsy} % for bold symbols (\bf doesn't work for Greek letters)
\usepackage{amsmath,amsfonts,amssymb,times,color}
%\usepackage{dsfont}
%\usepackage{pdflscape}
%\usepackage{lscape}
\usepackage{enumerate}
%\usepackage{fullpage}
\usepackage{natbib}
\usepackage{braket}
%\usepackage{rotating}
%\special{! /setdistillerparams where {pop}{userdict /setdistillerparams {pop} put}ifelse
% <</AutoRotatePages /None>> setdistillerparams}

%\usepackage[noBBpl,sc]{mathpazo}
\usepackage[noBBpl,sc]{mathpazo}
\linespread{1.1}
%\usepackage{amsmath,amsthm,amscd,amssymb}
\usepackage[papersize={6.9in, 10.0in}, left=.5in, right=.5in, top=.7in, bottom=.9in]{geometry}
\sloppy
\raggedbottom
\pagestyle{plain}


% these include amsmath and that can cause trouble in older docs.
\input{../helpers/fix-underbrace.tex}
\input{../helpers/cmrsum}

\usepackage{microtype}

\usepackage{xcolor}

\usepackage[colorlinks=true,
,urlcolor=blue
,anchorcolor=blue
,citecolor=blue
,filecolor=blue
,linkcolor=blue
,menucolor=blue,
,pdfborder=false
,linktocpage=true]{hyperref}
\hypersetup{
bookmarksopen=true,
bookmarksnumbered=true,
bookmarksopenlevel=10
}


%\usepackage[dotinlabels]{titletoc}
%\titlelabel{{\thetitle}.\quad}
%\input{../helpers/psu-plain-titles.tex}
%\input{../helpers/psu-sc-headers.tex}
\input{../helpers/fix-revtex-12.tex}
%\DeclareSymbolFont{CMlargesymbols}{OMX}{cmex}{m}{n}
%\DeclareMathSymbol{\sum}{\mathop}{CMlargesymbols}{"50}

\newtheorem{thm}{Theorem}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{conj}[thm]{Conjecture}
\theoremstyle{remark}
\newtheorem{rem}[thm]{Remark}
\theoremstyle{definition}
\newtheorem{dfn}[thm]{Definition}

% Environment commands
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\ba}{\begin{eqnarray}}
\newcommand{\ea}{\end{eqnarray}}
\newcommand{\ban}{\begin{eqnarray*}}
\newcommand{\ean}{\end{eqnarray*}}
\newcommand{\bit}{\begin{itemize}}
\newcommand{\eit}{\end{itemize}}
\newcommand{\ben}{\begin{enumerate}}
\newcommand{\een}{\end{enumerate}}


% Command aliases: notation
\newcommand{\tr}{\mathrm{tr}\,} % Trace
\newcommand{\br}[1]{\ensuremath{\left\langle #1 \right.}}
\newcommand{\bk}[2]{\br{{#1}}\ket{{#2}}}
\newcommand{\modu}[1]{\ensuremath{\left| {#1} \right|}}
\newcommand{\magn}[1]{\ensuremath{\modu{#1}^2}}
\newcommand\abs[1]{\left|#1\right|}
%\newcommand{\born}[2]{\ensuremath{\magn{\bk{{#1}}{{#2}}}}}
\newcommand{\born}[2]{\abs{\braket{#1|#2}}^2}
\newcommand{\half}{\ensuremath{\frac{1}{2}}}
\def\ketbra[#1]#2{\mathinner{\vert{#1}\rangle\langle{#2}}\vert}
\newcommand{\oepsilon}[2]{\epsilon(#1,#2)}

\newcommand{\C}{\mathbbm C} % Complex space
\newcommand{\id}{\mathds{1}}
\newcommand{\Hilb}{\mathbf{Hilb}}
\newcommand{\nCob}{\boldsymbol{n}\mathbf{Cob}}
\newcommand{\de}[1]{\,\mathrm{d}#1}
\newcommand{\sth}{\textsuperscript{th}}
%\newcommand{\bra}[1]{\left\langle#1\right|} % } These resize
%\newcommand{\ket}[1]{\left|#1\right\rangle} % }
%\newcommand{\bracket}[2]{\left\langle #1 \,\vrule\, #2 \right\rangle}
%\newcommand{\ketbra}[2]{\mathinner{#1\rangle\langle#2}}


% Command aliases: affiliations
%\newcommand{\oxford}{\affiliation{...University of Oxford...}}
%\newcommand{\wuerzburg}{\affiliation{Fakult{\"{a}}t f\"{u}r Physik und Astronomie, Universit\"{a}t W\"{u}rzburg, Am Hubland, 97074 W\"{u}rzburg, Germany}}

% Macros for comments (choose any color you want)
\definecolor{cautionred}{rgb}{1.0,0,0}
\newcommand{\note}[1]{{\marginpar{\vspace{1em}\color{red} \fontfamily{sfdefault}\fontsize{9pt}{9pt}\selectfont #1}}}
\newcommand{\add}[1]{{\color{blue} #1}}
\newcommand{\commentrl}[1]{{\color{green} #1}}
\newcommand{\caution}[1]{{\bf \color{cautionred} #1}}
%\newcommand{\cmt}[1]{{\color{red} #1}}

\newcommand{\red}{\color{red}}
\newcommand{\mg}{\color{magenta}}
\newcommand{\blk}{\color{black}}
\newcommand{\blu}{\color{blue}}
\definecolor{maroon}{rgb}{0.7,0,0}
\newcommand{\mr}{\color{maroon}}
\definecolor{ngreen}{rgb}{0.3,0.7,0.3}
\newcommand{\grn}{\color{ngreen}}
\definecolor{golden}{rgb}{0.8,0.6,0.1}
\newcommand{\gold}{\color{golden}}
\newcommand{\egc}[1]{{\color{maroon} \small \bf [[{#1}]]}}

	
\begin{document}
\title{No $\psi$-epistemic model can fully explain the indistinguishability of quantum states}
\author{Jonathan Barrett}
\affiliation{Department of Computer Science, University of Oxford, UK}
\author{Eric G. Cavalcanti}
\affiliation{School of Physics, The University of Sydney, Australia}
\affiliation{Department of Computer Science, University of Oxford, UK}
\author{Raymond Lal}
\affiliation{Department of Computer Science, University of Oxford, UK}
\author{Owen J. E. Maroney}
\affiliation{Faculty of Philosophy, University of Oxford, UK}
\pacs{03.65.Ta, 03.67.-a, 03.65.-w}
\date{30 Oct 2013 }

\begin{abstract}

According to a recent no-go theorem (M.~Pusey, J.~Barrett and T.~Rudolph, Nature Physics {\bf 8} 475 (2012)), models in which quantum states correspond to probability distributions over the values of some underlying physical variables must have the following feature: the distributions corresponding to distinct quantum states do not overlap. This is significant because if the distributions do not overlap, then the quantum state itself is encoded by the physical variables. In such a model, it cannot coherently be maintained that the quantum state merely encodes information about underlying physical variables. The theorem, however, considers only models in which the physical variables corresponding to independently prepared systems are independent. This work considers models that are defined for a single quantum system of dimension $d$, such that the independence condition does not arise. We prove a result in a similar spirit to the original no-go theorem, in the form of an upper bound on the extent to which the probability distributions can overlap, consistently with reproducing quantum predictions. In particular, models in which the quantum overlap between pure states is equal to the classical overlap between the corresponding probability distributions cannot reproduce the quantum predictions in any dimension $d \geq 3$. The result is noise tolerant, and an experiment is motivated to distinguish the class of models ruled out from quantum theory.
\end{abstract}
\maketitle

\newpage

%%%%%%%%NEW SECTION%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{mainbody}
\acknowledgements{
\textit{Acknowledgements.} We would like to thank Andrew Briggs, Matthew Leifer, Stephanie Simmons and Christopher Timpson for helpful discussions during the development of this work. EGC received support from an Australian Research Council grant DE120100559. OJEM, RL are supported by the John Templeton Foundation. This work is supported by the CHIST-ERA DIQIP project, and an FQXi Large Grant ``Time and the Structure of Quantum Theory''.
}
%\bibliographystyle{revtex}
\bibliography{epistemic_refs}

\newpage

\appendix*
\input{appendix1}
\input{appendix2}

\end{document}
