% !TEX root = runners-in-action.tex

\section{Related work}
\label{sect:relatedwork}

Comodels and (ordinary) runners have been used as a natural
model of stateful top-level behaviour.
%
For instance, Plotkin and Power~\cite{Plotkin:TensorsOfModels} have given a treatment of operational 
semantics using the tensor product of a model and a comodel.
%
Recently, Katsumata, Rivas, and Uustalu have generalised this interaction of models and comodels
to monads and comonads~\cite{Katsumata:InteractionLaws}.
%
An early version of \pl{Eff}~\cite{Bauer:AlgebraicEffects} implemented \emph{resources},
which were a kind of stateful runners, although they lacked satisfactory theory.
%
Uustalu~\cite{Uustalu:Runners} has pointed out that runners are the additional
structure that one has to impose on state
to run algebraic effects statefully.
%
Møgelberg and Staton's~\cite{Mogelberg:LinearUsageOfState} linear-use state-passing
translation also relies on equipping the state with a comodel
structure for the effects at hand. Our runners arise
when their setup is specialised to a certain Kleisli adjunction.

Our use of kernel state is analogous to the use
of parameters in parameter-passing 
handlers~\cite{Plotkin:HandlingEffects}: their 
$\tmkw{return}$ clause also provides a form of finalisation, as the 
final value of the parameter is available.
There is however no guarantee of
finalisation happening because handlers need not use the
continuation linearly.

The need to tame the excessive generality of handlers, and willingness to
give it up in exchange for efficiency and predictability, has recently
been recognised by \pl{Multicore OCaml}'s implementors, who
have observed that in practice most handlers resume 
continuations precisely once~\cite{Dolan:MulticoreOCaml}. In exchange for impressive efficiency, they require
continuations to be used linearly by default, whereas discarding and copying
must be done explicitly, incurring additional cost.
%
Leijen~\cite{Leijen:Finalisation} has extended 
handlers in \pl{Koka} with a $\tmkw{finally}$ clause, whose 
semantics ensures that finalisation happens whenever a handler
discards its continuation.
Leijen also added an $\tmkw{initially}$ clause to
parameter-passing handlers, which is used to compute the initial value of the parameter before handling, but that 
gets executed again every time the handler resumes its continuation.


%%% Local Variables:
%%% mode: latex
%%% TeX-master: "runners-in-action"
%%% End:
