\documentclass[11pt]{jarticle} %%% アンコメント
\ifx\nameofpaper\undefined %%% もし main.tex が読まれていなければ
  \usepackage{macro_natsuume} %%% 自分用のマクロファイルを読み込み，
  \def\beginsection{\section*}
  \def\endofsection{\end{document}} %%% \endofsection = \bye と定義
  \input draft_header.tex
\else %%% そうでなければ
  \def\beginsection{\chapter}
  \def\endofsection{\endinput} %%% \endofsection = \endinput と定義
\fi


%%%%%%%%%
\section{Appendix: Review of gauge theory}\label{sec:YM}
%%%%%%%%%

\newcommand{\Am}{A_\mu}

%%%
\head{$U(1)$ gauge theory}
%%%

First, recall a $U(1)$ gauge theory. Consider a complex scalar field:
\be
%
\calL = - \del_\mu \phi^* \del^\mu \phi -m^2 \phi^*\phi~.
%\label{eq:}
%
\ee
The theory is invariant under the global $U(1)$ transformation
\be
%
\phi \rightarrow U \phi~, \quad (U:= e^{i\alpha})~.
%\label{eq:}
%
\ee
The theory is not invariant under the local $U(1)$ transformation though:
\be
%
\phi \rightarrow U(x) \phi~.
%\label{eq:}
%
\ee
The local symmetry can be incorporated by introducing a $U(1)$ gauge field $\Am$ and the covariant derivative $D_\mu$:
\be
%
D_\mu \phi = (\del_\mu - i\Am)\phi~,
%\label{eq:}
%
\ee
where $\Am$ transforms as 
\be
%
\Am \rightarrow \Am + \del_\mu \alpha = \Am - i(\del_\mu U)U^{-1}~.
%\label{eq:}
%
\ee
The covariant derivative transforms as 
\be
%
D_\mu \phi \rightarrow U D_\mu \phi~, 
\quad \text{or} \quad
D_\mu \rightarrow U D_\mu U^{-1}~.
%\label{eq:}
%
\ee
One makes the invariant action using covariant (or invariant) quantities such as $D_\mu\phi$ and $F_{\mu\nu} = \del_\mu A_\nu - \del_\nu A_\mu$. The gauge field action is given by
\be
%
\calL = - \frac{1}{4e^2} F^{\mu\nu}F_{\mu\nu}~.
%\label{eq:}
%
\ee
We normalized the gauge field so that the charge $e$ (coupling constant) appears only as the overall normalization. The usual form is obtained by
\be
%
\Am \rightarrow e \Am~.
%\label{eq:}
%
\ee
Then, the transformation law and the covariant derivative become
\begin{align}
%
\Am & \rightarrow \Am + \frac{1}{e}\del_\mu \alpha~,\\
D_\mu \phi &= (\del_\mu - ie\Am)\phi~.
%\label{eq:}
%
\end{align}

%%%
\head{$U(\Nc)$ gauge theory}
%%%

Now, let us consider a scalar field with $\Nc$ components $\phi_i = (\phi_1, \phi_2, \cdots)$. Consider the Lagrangian given by
\be
%
\calL = - (\del_\mu \phi)^\dag (\del^\mu \phi) -m^2 \phi^\dag\phi~.
%\label{eq:}
%
\ee
The theory has a global $U(\Nc)$ symmetry:
\be
%
\phi_i \rightarrow U_i^{~j} \phi_j~,
%\label{eq:}
%
\ee
where $U$ is a unitary $\Nc\times\Nc$ matrix. Again, one can promote the symmetry to a local symmetry by  introducing a gauge field $\Am$ and the covariant derivative $D_\mu$ for $\phi$:
\be
%
D_\mu \phi = (\del_\mu - i\Am)\phi~,
%\label{eq:}
%
\ee
where $\Am$ is a Hermitian $\Nc\times\Nc$ matrix $(\Am)_i^{~j}$ and transforms as 
\be
%
\Am \rightarrow U \Am U^{-1} - i(\del_\mu U)U^{-1}~.
\label{eq:A_transf}
%
\ee
Then, the covariant derivative transforms as 
\begin{align}
%
D_\mu \phi &\rightarrow (\del_\mu U) \phi + U\del_\mu \phi 
- i\{ U \Am U^{-1} - i(\del_\mu U)U^{-1} \} U \phi \\
&= U (\del_\mu - i\Am) \phi~,
%\label{eq:}
%
\end{align}
so
\be
%
D_\mu \phi \rightarrow U D_\mu \phi~, 
\quad \text{or} \quad
D_\mu \rightarrow U D_\mu U^{-1}~.
%\label{eq:}
%
\ee

%One calls that the field $\phi_i$ belongs the fundamental representation of the gauge group. So far, we consider a matter in the fundamental representation, but matter fields in the adjoint representation often appear in AdS/CFT. An adjoint scalar $(\phi_A)_i^{~j}$ transforms as
So far, we consider the matter field $\phi_i$, a field in the fundamental representation, but matter fields in the adjoint representation often appear in AdS/CFT. An adjoint scalar $(\phi_A)_i^{~j}$ transforms as
\be
%
\phi \rightarrow U\phi_A U^{-1}~.
%\label{eq:}
%
\ee
Namely, it transforms as $\phi_i (\phi^\dag)^j$. The gauge field $(\Am)_i^{~j}$ also belongs to the adjoint representation although $\Am$ transformation \eqref{eq:A_transf} itself has an inhomogeneous term $(\del_\mu U)U^{-1}$. The definition of the covariant derivative depends on the fields it acts. Following the same procedure as $\phi_i$, one can show that the covariant derivative of $\phi_A$ is given by
\be
%
D_\mu \phi_A = \del_\mu \phi_A - i [\Am,\phi_A]~.
%\label{eq:}
%
\ee

One can show
\be
%
[D_\mu, D_\nu]\phi = \left\{ \del_\mu A_\nu - \del_\nu A_\mu - i [A_\mu, A_\nu] \right\} \phi~.
%\label{eq:}
%
\ee
Note that $D_\mu$ is a differential operator, but its commutator is no longer a differential operator. So, we define the field strength as
\be
%
F_{\mu\nu} =  -i[D_\mu, D_\nu] = \del_\mu A_\nu - \del_\nu A_\mu - i [A_\mu, A_\nu]~.
%\label{eq:}
%
\ee
Because $D_\mu \rightarrow U D_\mu U^{-1}$, $F_{\mu\nu}$ transforms as
\be
%
F_{\mu\nu} \rightarrow U F_{\mu\nu} U^{-1}~.
%\label{eq:}
%
\ee
In the $U(1)$ gauge theory, $F_{\mu\nu}$ is an invariant quantity, but here $F_{\mu\nu}$ is a covariant quantity. We define the gauge theory action as
\be
%
\calL = - \frac{1}{2g_\text{YM}^2} \text{tr} (F^{\mu\nu}F_{\mu\nu})~.
%\label{eq:}
%
\ee

Write the transformation matrix $U$ as
\begin{align}
%
U &= e^{iH}~, \\
H &= \alpha \cdot 1 + \alpha^a t^a~. \quad (a=1,\ldots, \Nc^2-1)
%\label{eq:}
%
\end{align}
The parameter $\alpha$ represents $U(1)$ transformation, and the parameters $\alpha^a$ represent $SU(\Nc)$ transformations. The matrices $t^a$ are called $SU(\Nc)$ generators. Because $U$ is a unitary matrix, $t^a$ are traceless Hermitian matrices. In general, the generators satisfy the commutation relation
\be
%
\left[ t^a, t^b \right] = if_{abc} t^c~,
%\label{eq:}
%
\ee
where $f_{abc}$ is called the structure constant of the gauge group. We normalize the generators so that
\be
%
\text{tr}\left( t^a t^b \right) = \frac{1}{2} \delta^{ab}~.
%\label{eq:}
%
\ee
Using $t^a$, one can decompose matrix fields such as $\Am$ and $F_{\mu\nu}$, \eg, 
\be
%
(\Am)_i^{~j} = \Am^a (t^a)_i^{~j}~.
%\label{eq:}
%
\ee
Then, the action becomes
\be
%
\calL = - \frac{1}{2g_\text{YM}^2} \text{tr} (F^{\mu\nu}F_{\mu\nu})
= - \frac{1}{4g_\text{YM}^2} F^{a\mu\nu}F^a_{\mu\nu}~.
%\label{eq:}
%
\ee

\endofsection
