%\documentclass[11pt]{jarticle} %%% アンコメント
\ifx\nameofpaper\undefined %%% もし main.tex が読まれていなければ
  \usepackage{macro_natsuume} %%% 自分用のマクロファイルを読み込み，
  \def\beginsection{\section*}
  \def\endofsection{\end{document}} %%% \endofsection = \bye と定義
  \input draft_header.tex
\else %%% そうでなければ
  \def\beginsection{\chapter}
  \def\endofsection{\endinput} %%% \endofsection = \endinput と定義
\fi


%%%%%%%%%
\beginsection{Other AdS spacetimes}\label{chap:others}
%%%%%%%%%

%\begin{flushright}
%        \today\\
%\end{flushright}

\begin{quote}
We discuss various spacetimes which often appear in AdS/CFT. We describe charged AdS black holes, the Schwarzschild-AdS black hole in the other dimensions, various branes (M-branes and D$p$-branes), and some other examples. Many of them are asymptotically AdS spacetimes, but some are not.
\end{quote}

%%%%%%%%%
\section{Overview of other AdS spacetimes
%そのほかのAdS時空
}\label{sec:others}
%%%%%%%%%

So far we discussed the AdS$_5$ spacetime and the Schwarzschild-AdS$_5$ (SAdS$_5$) black hole, but various other asymptotically AdS spacetimes appear in the literature. Here, we briefly discuss them. We discuss some of the geometries further in \sect{others_details}. Currently, one cannot directly compute QCD in AdS/CFT. So, one analyzes, \eg, the black hole which is dual to the \Nfour\ SYM, but the results are not the ones for QCD. Therefore, when one computes a physical quantity, the result may be theory-dependent. It is desirable to compute a physical quantity in various theories in order to know the theory dependence.  
%これまでAdS$_5$時空やSAdS$_5$ブラックホールを扱ってきたが，ほかにもさまざまな漸近的AdS時空が議論されている．ここではそれらを簡単に紹介する．AdS/CFTでは（現在のところ）直接QCDを計算できるわけではない．したがって，${\cal N}=4$ゲージ理論（に対応するブラックホール）などを使って解析するが，それは本来ほしいQCDの結果ではない．このため，なにかの物理量を計算したときに，結果がどの程度理論の詳細によるかを知るためにも，さまざまな理論で計算することが望ましい．

% added
There are various approaches below, but in all cases, the symmetry of both the gauge theory and the gravitational theory plays an important role. 

%%-------------------------------------
\subsection*{Charged AdS black holes
%荷電AdSブラックホール
}%\label{sec:}
%%-------------------------------------

As in the asymptotically flat spacetime, one can consider charged black holes in AdS spacetime. The Einstein-Maxwell theory gives a simple example: 
%漸近的平坦な時空と同様に，荷電AdSブラックホールを考えることができる．まず，アインシュタイン-マクスウェル理論を考える：
\be
%
\action =  \frac{1}{16\pi G_5} \int d^{5}x\, \sqrt{-g} \left(R - 2\Lambda - \frac{L^2}{4}F_{MN}^2 \right)~.
%\label{eq:}
%
\ee
The bulk Maxwell field $A_M$ corresponds to the boundary current $J^\mu$. So, the electric charge case, $A_0 \neq 0$, corresponds to a finite chemical potential or a finite charge density. The solution is known as the \keyword{Reissner-Nordstr\"{o}m AdS black hole} (RN-AdS black hole hereafter).
%バルクにゲージ場$A_M$がある場合は，AdS境界にカレント$J^\mu$が存在する場合に相当するので，$A_0 \neq 0$の場合は有限化学ポテンシャル／有限密度に相当する．この場合の解は，{\bf ライスナー-ノルドストリュームAdS}（RN-AdS）{\bf ブラックホール}\index{らいすなーのるとすとりゆーむふらつくほーる@ライスナー-ノルドストリュームAdSブラックホール}と呼ばれる．

One generalization of the charged AdS black hole is the \keyword{R-charged black hole} \cite{Behrndt:1998jd,Kraus:1998hv,Cvetic:1999xp}. The R-charged black hole is dual to the \Nfour\ SYM at a finite chemical potential. In this case, the bulk gravitational action is complicated, but it schematically takes the form
%また，荷電ブラックホールの1つの一般化として，{\bf R荷電ブラックホール}\index{Rかてんふらつくほーる@R荷電ブラックホール}が知られている．この場合のバルク作用は複雑なので，形式的に書くと，
\be
%
{\cal L}_5 =  \text{(Gravity)} + \text{(three $U(1)$ gauge fields)} + \text{(two real scalars)}~.
%\label{eq:}
%
\ee
The \Nfour\ SYM is dual to the AdS$_5 \times S^5$ spacetime. An R-charged black hole is obtained by adding an angular momentum along $S^5$. The sphere $S^5$ has the $SO(6)$ symmetry which is rank three, so one can add at most three independent angular momenta.
%もともとの10次元の立場では，${\cal N}=4$ゲージ理論はAdS$_5 \times S^5$時空に対応したが，この場合は $S^5$方向に角運動量を入れたことに相当する．$S^5$の対称性，$SO(6)$ はランク3なので，3つの独立な角運動量を与えることができる．

If one compactifies $S^5$, one gets Kaluza-Klein gauge fields \index{Kaluza-Klein gauge field} from the compactified metric%
\footnote{See \sect{vector} for the simple $S^1$ compactification.}.
The angular momenta along $S^5$ become the charges under the Kaluza-Klein gauge fields from the five-dimensional point of view. The symmetry of $S^5$ corresponds to the internal symmetry of the dual gauge theory, R-symmetry, so such a charge is called a R-charge\index{R-charge}.
%5次元の立場では，コンパクト化によって，コンパクト方向の計量からゲージ場が出てくる（カルツァ-クライン・ゲージ場\index{かるつあくらいんけーしは@カルツァ-クライン・ゲージ場}）%
%\footnote{簡単な$S^1$コンパクト化については，\sect{vector}で議論している．}．
%$S^5$方向の角運動量は，5次元の立場ではこれらゲージ場に対する電荷である．$S^5$の対称性はR対称性に対応したので，このような電荷をR電荷\index{Rてんか@R電荷}と呼ぶ．

When all three charges are equal, the R-charged black hole reduces to the RN-AdS black hole above. However, the R-charged black hole in general differs from the RN-AdS black hole. This is because scalar fields are nonvanishing for the R-charged black hole. For asymptotically flat black holes, a scalar must vanish, and the existence of a scalar does not affect a black hole solution \cite{Bekenstein:1971hc}. This is because of the no-hair theorem\index{no-hair theorem}. But this is not the case for the R-charged black hole. Namely, the no-hair theorem is often violated in the AdS spacetime. 
%R荷電ブラックホールで3つの電荷が等しい場合は，上のRN-AdSブラックホールに帰着する．R電荷が1つの場合は，RN-AdSブラックホールとは違う解である．これはスカラー場が存在するからである．もっとも，漸近的平坦な時空の場合，スカラー場はゼロにしかならずブラックホール解は変わらない．これはノー・ヘア定理\index{のーへあていり@ノー・ヘア定理}によるものである．つまり，AdS時空ではノー・ヘア定理は必ずしも成り立つわけではない．

The R-charged black hole undergoes a second-order phase transition \cite{Cai:1998ji,Cvetic:1999ne}.

%%-------------------------------------
\subsection*{Schwarzschild-AdS black holes in arbitrary dimensions
%一般次元のAdSブラックホール
}%\label{sec:}
%%-------------------------------------

From the gravitational theory point of view, one may consider the SAdS$_{p+2}$ black hole as a simple extension of the SAdS$_5$ black hole. The dual gauge theory should be a $(p+1)$-dimensional conformal theory. The details of the dual gauge theory are unclear though because we do not go through the brane argument such as \sect{D3_brane}. But when $p=2$ and 5, there are brane realizations, the M2-brane and the M5-brane below.
%重力理論の立場では，簡単な一般化として一般次元のSAdS$_{p+2}$ブラックホールを考えることができる．境界理論としては，$(p+1)$次元のコンフォーマルな理論に相当する．ただし，\sect{D3_brane}のようなブレーンを経由していないので，これ以上の詳細は不明だが，$p=2, 5$ の場合はそれぞれ以下のM2, M5ブレーンとして考えることができる．

%%-------------------------------------
%\subsection{そのほかのブレーン}\label{sec:other_branes}
%%-------------------------------------

%%-------------------------------------
\subsection*{Other branes (M-branes)
%そのほかのブレーン（Mブレーン）
}%\label{sec:M_brane}
%%-------------------------------------

In string theory, there appear various branes, and the AdS/CFT duality based on these branes has been discussed (see, \eg, Ref.~\cite{Marolf:2011zs} for a review of these branes).
%超弦理論ではさまざまなブレーンが登場し，それらを使ったAdS/CFT双対性も議論されている．

The \keyword{M-branes} appear in the 11-dimensional theory known as the \keyword{M-theory}. The M-theory does not yet have a microscopic formulation unlike string theory, but we know several clues about the theory:
%{\bf Mブレーン}\index{Mふれーん@Mブレーン}は，{\bf M理論}\index{Mりろん@M理論}と言う11次元理論にあらわれるブレーンである．しかし，超弦理論と違い，M理論の微視的な定式化はまだない．とは言え，いくつかはっきりしていることがある．
\begin{itemize}

\item First, the M-theory is not a theory of strings. For a superstring theory to be consistent, the spacetime dimensionality must be 10.
%まず，M理論はストリングの理論ではない．ストリングが無矛盾であるためには，10次元時空であることが要求されるからである．

\item At low energy, the M-theory is described by the 11-dimensional supergravity. This is just as  superstring theories are described by 10-dimensional supergravities at low energy. Just as the D-brane is a solution of a supergravity [\cf, \eq{D3_extreme}], the M-brane is a solution of the 11-dimensional supergravity. What we know about M-branes mostly comes from the 11-dimensional supergravity.
%また，低エネルギーでは11次元の超重力理論であらわされる．ちょうど超弦理論の低エネルギー有効理論が10次元超重力理論だというのと同じである．Dブレーンが超重力理論の解であったように［\eq{D3_extreme}を参照］，Mブレーンも超重力理論の解として議論できる．

\item Although the M-theory is not a string theory, it is related to string theory. The $S^1$-compactification of the M-theory becomes the so-called \keyword{type IIA superstring theory}. As a result, the M-branes and the D-branes are related to each other.
%さらに，M理論はストリングの理論ではないものの，超弦理論とは無関係ではない．M理論を1次元コンパクト化したものは，{\bf タイプIIA超弦理論}\index{たいぷIIA@タイプIIA超弦理論}と呼ばれる超弦理論に帰着する．このため，MブレーンとDブレーンのあいだにも関係がある．

\end{itemize}

There are two kinds of M-branes: the M2-brane and the M5-brane. The M2-brane reduces to the AdS$_4 \times S^7$ spacetime in the near-horizon limit. From the symmetry of the geometry, the dual gauge theory must be a $(2+1)$-dimensional conformal theory with R-symmetry $SO(8)$. Also, the M2-brane at finite temperature reduces to the SAdS$_4 \times S^7$ spacetime. But the full details of the dual gauge theory are still unclear. Unlike string theory, the M-theory does not yet have a microscopic formulation, so we do not have a microscopic tool such as the D-brane. 
%Mブレーンとしては，M2ブレーンとM5ブレーンがある．M2ブレーンは，near-horizon 極限でAdS$_4 \times S^7$ 時空に帰着する．対称性から対応するゲージ理論は，$(2+1)$次元のコンフォーマル理論で，R対称性 $SO(8)$をもつはずである．また有限温度では，シュワルツシルドAdS$_4 \times S^7$時空に帰着する．しかし，M2 ブレーンに対応するゲージ理論が何かはまだはっきりしていない．超弦理論と違い，M理論には微視的な定式化がないため，Dブレーンのように微視的な立場からの解析ができない．このため，超重力理論以外の手がかりがないという点もある．

Moreover, AdS/CFT predicts a characteristic behavior of thermodynamic quantities. For example,  the energy density of the M2-brane behaves as $\varepsilon \propto \Nc^{3/2} T^3$. The temperature dependence $T^3$ is natural and comes from the Stefan-Boltzmann law \index{Stefan-Boltzmann law} in $(2+1)$-dimensions. But the $\Nc$-dependence is not easy to understand. For gauge theories, it is natural to have the dependence $\Nc^2$ like the D3-brane, but the M2-brane does not have such a dependence. The power 3/2 implies that the M2-brane is not described by a simple gauge theory but is described by a highly nontrivial gauge theory. The dual gauge theory is not completely understood, but there are important progresses in recent years, and the power 3/2 has been understood to some extent \cite{Aharony:2008ug,Drukker:2010nc}.
%さらに，AdS/CFTの予言では，熱力学量は $\varepsilon \propto N_c^{3/2} T^3$ という振るまいをする．温度依存性$T^3$ は $(2+1)$ 次元のステファン-ボルツマンの法則\index{すてふあんほるつまんのほうそく@ステファン-ボルツマンの法則}で自然である．しかし，$N_c$ 依存性が奇妙である．ゲージ理論の場合， D3 ブレーンのように $N_c^2$ が自然であるが，そうはなっていない．$N_c^{3/2}$ というベキは，M2 ブレーンが単純なゲージ理論ではなく，きわめてノントリビアルな理論だということを示唆する．M2ブレーンに対応するゲージ理論については，完全には決着がついていないが，近年大きな進展があり，$N_c^{3/2}$ のベキについても，ある程度理解されてきている\cite{Aharony:2008ug,Drukker:2010nc}．

On the other hand, the M5-brane reduces to the AdS$_7 \times S^4$ spacetime in the near-horizon limit. Then, the dual gauge theory must be a $(5+1)$-dimensional conformal theory with R-symmetry $SO(5)$. Also, the M5-brane at finite temperature reduces to the SAdS$_7 \times S^4$ spacetime. Again, thermodynamic quantities have a characteristic behavior in $\Nc$, \eg, $\varepsilon \propto \Nc^3 T^6$.
%一方，M5ブレーンは，near-horizon 極限でAdS$_7 \times S^4$ 時空に帰着する．このため対応するゲージ理論は，$(5+1)$次元のコンフォーマル理論で，R対称性 $SO(5)$をもつはずである．また有限温度では，シュワルツシルドAdS$_7 \times S^4$時空に帰着する．やはり，熱力学量は $\varepsilon \propto N_c^3 T^5$ と$N_c$ 依存性が奇妙な振るまいをする．

%%-------------------------------------
\subsection*{Other branes (D$p$-branes)
%そのほかのブレーン（D$p$ブレーン）
}%\label{sec:Dp_brane}
%%-------------------------------------

One can consider the D$p$-brane \index{D$p$-brane} with $p \neq 3$ \cite{Itzhaki:1998dd}. As in the D3-brane case, the D$p$-brane is dual to the $(p+1)$-dimensional SYM. The geometry has the \poincare\ invariance $ISO(1,p)$ and R-symmetry $SO(9-p)$. But the geometry does not reduce to the AdS spacetime in the near-horizon limit, so it does not have a conformal invariance. In particular, there is no scale invariance%
\footnote{The theory is scale invariant up to an overall scaling of the metric. An extension of such a geometry is known as the \keyword{hyperscaling violating geometry} \cite{Huijse:2011ef}.} 
because the dilaton is nontrivial (\sect{SUGRA}). From the gauge theory point of view, the $(p+1)$-dimensional gauge theory does not have a scale invariance because the coupling constant is dimensionful as $[g_\text{YM}^2] = \text{L}^{p-3}$. 
%D3ブレーン以外のD$p$ブレーン\index{Dp@D$p$ブレーン}を考えることもできる．D3ブレーン同様に，これらは$(p+1)$次元超対称ゲージ理論に対応する．D$p$ブレーンは，ポアンカレ対称性$ISO(1,p)$とR対称性$SO(9-p)$をもつ．しかし，near-horizon 極限でAdS時空には帰着しない．したがって，コンフォーマル対称性をもたない．実際，ディラトンがノントリビアルなので（\sect{SUGRA}），AdS時空のようなスケール対称性はない．ゲージ理論の立場では，$(p+1)$次元のゲージ理論では結合定数が $[g_\text{YM}^2] = \text{L}^{p-3}$ と次元をもつため，スケール対称性がないことに対応する．

The energy density behaves as%
\footnote{One normally considers the $p<5$ case because the heat capacity diverges for $p=5$ and becomes negative for $p>5$.}
%熱力学量を計算すると，たとえばエネルギー密度は%
%\footnote{比熱は $p=5$ で発散，$p>5$ で負になるので，通常 $p<5$を考える．}
\be
%
\varepsilon \propto (g_\text{YM}^2 N_c)^{(p-3)/(5-p)} N_c^2 T^{2(7-p)/(5-p)}~.
\label{eq:energy_Dp}
%
\ee
In this case, we do not have the $(p+1)$-dimensional Stefan-Boltzmann law because we have a dimensionful quantity $g_\text{YM}$ other than the temperature. Also, thermodynamic quantities depend on the \thooft\ coupling $\lambda$ unlike the D3-brane and the M-branes.
%この場合は，温度以外に次元をもつ量 $g_\text{YM}$ があるため，$(p+1)$次元のステファン-ボルツマンの法則にはならない．また，D3ブレーンやMブレーンとは異なり，トホーフト結合定数 $\lambda$ にも依存している．

However, there are exceptions. For $p=1$, $\varepsilon \propto \lambda^{-1/2} N_c^2 T^3 \propto N_c^{3/2} T^3$, which is the same behavior as the M2-brane. Similarly, for $p=4$, $\varepsilon \propto \lambda N_c^2 T^3 \propto N_c^{3} T^6$, which is the same behavior as the M5-brane. These behaviors are not just coincidences. They arise because some D-branes have their origins in the 11-dimensional M-branes although we do not discuss the details here.
%しかし，$p=1$ で $\varepsilon \propto \lambda^{-1/2} N_c^2 T^3 \propto N_c^{3/2} T^3$ となり，M2ブレーンの振るまいと同じである．同様に，$p=4$ で $\varepsilon \propto \lambda N_c^2 T^3 \propto N_c^{3} T^6$ となり，M5ブレーンの振るまいと一致する．これらの振るまいは偶然ではない．ここでは詳細に議論しないが，Dブレーンの一部は11次元のMブレーンに起源をもつからである．

%%-------------------------------------
\subsection*{Extensions from the gauge theory point of view
%Deformations of \Nfour\ SYM
%${\cal N}=4$ゲージ理論の deformation
}%\label{sec:}
%%-------------------------------------

Starting from the \Nfour\ SYM, one can obtain new gauge theories by adding deformations. \index{deformations} For example, the \Nfour\ SYM has various matter fields in the adjoint representation, and one can add deformations to make these fields massive. The resulting theories have less supersymmetry compared to the \Nfour\ SYM. Also, the \Nfour\ SYM is scale invariant, so the theory does not have the confining phase but has only the plasma phase. The deformations of the \Nfour\ SYM are generally not scale invariant and some have a dynamical confinement like QCD. Some examples of dual geometries are 
\begin{itemize}
\item Pilch-Warner geometry (${\cal N}=2^*$ theory) \index{Pilch-Warner geometry} \cite{Pilch:2000ue,Buchel:2003ah},
\item Polchinski-Strassler geometry \index{Polchinski-Strassler geometry}  (${\cal N}=1^*$ theory) \cite{Polchinski:2000uf}, 
\item Klebanov-Strassler geometry \index{Klebanov-Strassler geometry} \cite{Klebanov:2000hb,Gubser:2001ri}.
%\item Maldacena-Nunez geometry \index{Maldacena-Nunez geometry} \cite{Maldacena:2000yy,Buchel:2001qi,Gubser:2001eg},
\end{itemize}
This approach has both the advantage and the disadvantage:
\begin{itemize}
\item Advantage: the dual gauge theory is clear. 
\item Disadvantage: the resulting gravitational theory and the metric are often very complicated and difficult to analyze.
\end{itemize}

%${\cal N}=4$ゲージ理論にはさまざまな随伴表現が登場したが，たとえばこれらの場に質量をもたせるような摂動（deformation）を加えることも可能である．このようにしてできた理論は，もとの${\cal N}=4$ゲージ理論と比べて超対称性が小さい．また，${\cal N}=4$ゲージ理論ではスケール対称性から閉じこめが起きず，プラズマ相しか存在しなかった．しかし，このような理論はスケール対称性を保たず，QCDのようにダイナミカルな閉じこめが起こるものもある．このような理論に対応する時空はいくつか知られており，代表的なものとしてKlebanov-Strassler 時空\index{Klebanov-Strassler@Klebanov-Strassler 時空}や，Polchinski-Strassler 時空\index{Polchinski-Strassler@Polchinski-Strassler 時空}などがある\cite{Klebanov:2000hb,Polchinski:2000uf}．

%このアプローチの利点・欠点としては，
%\begin{itemize}
%\item 利点：対応する場の理論が明確．
%\item 欠点：重力理論や計量がしばしば複雑で解析が困難．
%\end{itemize}

%%-------------------------------------
\subsection*{Extensions from the gravitational theory point of view
%Other examples
%対応が未知のもの
}%\label{sec:}
%%-------------------------------------

When one adds a deformation to the \Nfour\ SYM, one starts with the gauge theory point of view. In contrast, one could start with the gravitational theory point of view. There are at least two approaches:
%deformationの立場ではゲージ理論から出発して考えるが，対照的に重力理論から出発して考える立場もある．このアプローチには少なくとも2つのパターンがある．

\begin{enumerate}

\item \textit{Start with a gravitational theory which is easy to handle.} When one adds a deformation, the resulting gravitational action is often highly complicated and is hard to analyze. To avoid the problem, one could choose a theory which is natural or simple from the gravitational theory point of view. The SAdS$_{p+2}$ or RN-AdS black holes are the examples. The \textit{holographic superconductor}\index{holographic superconductors} is another typical example (\sect{H^3}). The holographic superconductor has the action 
%{\bf 重力理論として解析しやすい理論から出発する．}deformationの立場では，重力理論が複雑でしばしば手に負えないので，重力理論として自然なもの，解析しやすいものを選ぶ方法である．代表的なものに，ホログラフィック超伝導\index{ほろくらふいつくちようてんとう@ホログラフィック超伝導}が挙げられる（\sect{H^3}）．ホログラフィック超伝導は，
% normalization changed
%\mathindent=0zw % <-- [1]
\begin{align}
%
& \action = \int d^{p+2}x \sqrt{-g} \left[ R - 2\Lambda - \frac{1}{4} F_{MN}^2  
- |D_M\Psi|^2 -V(|\Psi|) \right]~, \\
& D_M := \nabla_M - i \qm A_M~,
%\label{eq:H^3_action}
%
\end{align}
%\mathindent=2zw % <-- [2] デフォルト
where $\Psi$ is a complex scalar field.
%という系である．$\Psi$ は複素スカラー場である．

\item \textit{Start with the symmetry of the metric.} In this case, one first writes down the metric with the desired symmetry and then tries to find a gravitational theory which admits such a metric as a solution. A typical example is the \keyword{Lifshitz geometry} \cite{Kachru:2008yh}. The metric 
%{\bf 計量のもつ対称性から出発する．}代表的なものにリフシッツ時空\index{りふしつつしくう@リフシッツ時空}が挙げられる\cite{Kachru:2008yh}．これは境界理論が，これまで議論したスケール対称性にかわって
%という時間方向と空間方向で非等方のスケール対称性をもつ場合である（AdS時空の場合は$z=1$に相当する）．このようなスケーリングは，\sect{exponents}の動的臨界現象に関連して議論されている．この対称性に対応して，
\be
%
\frac{ds^2}{L^2} = -r^{2z}dt^2 + r^2 d\bmx_p^2 + \frac{dr^2}{r^2}
\label{eq:Lifshitz}
%
\ee
is invariant under the anisotropic scaling 
\be
%
t \rightarrow a^z t~, \bmx \rightarrow a \bmx~, r \rightarrow r/a
%\label{eq:}
%
\ee
instead of the scaling discussed so far which corresponds to $z=1$. Such an anisotropic scaling is partly motivated by the dynamic critical phenomena in \sect{exponents}.
%という計量が議論されている．計量は$t \rightarrow a^z t~, \bmx \rightarrow a \bmx, r \rightarrow r/a$の下で不変である．

This approach starts with the metric, so the first questions one would ask are what kind of gravitational theory admits such a solution and whether such a gravitational theory is natural from the string theory point of view. Also, the metric \eqref{eq:Lifshitz} has a spacetime singularity at $r=0$ (when $z \neq 1$)%
\footnote{Unlike the Schwarzschild black hole, a curvature invariant such as $R^{MNPQ}R_{MNPQ}$ does not diverge at this singularity, but the tidal force diverges there. Some branes have similar spacetime singularities. A spacetime singularity is called a s.p. (scalar polynomial) singularity \index{scalar polynomial singularity} if a curvature invariant has a divergence and is called a p.p. (parallelly propagated) singularity \index{parallelly propagated singularity} if the tidal force diverges. There are various kinds of spacetime singularities and one cannot completely classify them. Thus, in general relativity, a spacetime singularity is defined operationally as geodesic incompleteness. \index{geodesically incomplete} }. 
The curvature becomes strong at the origin, and $\alpha'$-corrections could become important, so physical quantities one computes may not be reliable. 

%この場合，計量から出発するので，この計量が解となる重力理論は何か，またその重力理論は超弦理論の立場として自然かがまず問題になる．さらに計量 \eqref{eq:Lifshitz} は$r=0$で時空特異点をもつ（$z\neq1$の場合）%
%\footnote{時空特異点と言っても，曲率テンソルで作る不変量$R^{MNPQ}R_{MNPQ}$などに発散がなく，潮汐力が発散するという特異点である．$R^{MNPQ}R_{MNPQ}$などが発散するものをs.p.（scalar polynomial）特異点\index{scalar polynomialとくいてん@scalar polynomial特異点}，潮汐力が発散するものをp.p.（parallelly propagated）特異点\index{parallelly propagatedとくいてん@parallelly propagated特異点}と言う．一部のブレーンも同様の特異点をもつ．実は，時空特異点にはさまざまな種類があり，完全に分類することはできない．このため，一般相対論では時空特異点を測地線不完備（geodesically incomplete）\index{そくちせんふかんひ@測地線不完備}として操作主義的に定義する．測地線不完備とは，少なくとも1つの測地線が有限の「固有時間」（有限アフィン・パラメータ）で終端を迎え，先には延ばせないことを言う．}．
%曲率が大きくなるので，（$\alpha'$補正が大きく効いて）計算する物理量が信頼できるのかといった問題がありうる．

\end{enumerate}
These approaches also have both the advantage and the disadvantage: 
\begin{itemize}
\item Advantage: one can choose a simple gravitational theory. One can use various gravitational theories. 
\item Disadvantage: the dual field theory is not very clear. In particular, one lacks the precise AdS/CFT dictionary which relates the parameters of the gravitational theory and the field theory, so quantitative predictions may be difficult.
\end{itemize}

%どちらのパターンも利点・欠点としては，
%\begin{itemize}
%\item 利点：重力理論として単純なものを選べる．さまざまな重力理論を使うことが可能．
%\item 欠点：対応する場の理論が明確ではない．とくに重力理論と場の理論のパラメータを関係づけるAdS/CFT辞書が欠けているので，定量的な予言が難しい．
%\end{itemize}

%%-------------------------------------
\subsection*{Other holographic theories}%\label{sec:}
%%-------------------------------------

%So far we described generalizations of AdS/CFT where underlying bulk theories are string theory or are assumed to be string theory. But various variations of AdS/CFT have been proposed in the literature. For those variations, underlying bulk theories are not (necessarily) string theory. Examples include
Finally, various generalizations of AdS/CFT have been proposed in the literature. They are holographic theories in a broad sense. Examples include
\begin{itemize}
\item dS/CFT correspondence (see Ref.~\cite{Spradlin:2001pw} for a review).
\item Kerr/CFT correspondence (see Ref.~\cite{Bredberg:2011hp} for a review).
\item Correspondence between the Rindler space \eqref{eq:rindler} and an incompressible fluid \cite{Bredberg:2010ky}.
\item Higher-spin holography \cite{Klebanov:2002ja}.
%(correspondence between ``higher-spin gauge theory" (see Ref.~\cite{Vasiliev:2014vwa} for a review) in the AdS$_4$ spacetime and the three-dimensional $O(N)$ vector model).
\end{itemize}


%%------------------
\titlenewterms
%%------------------

\begin{multicols}{2}
\noindent
Reissner-Nordstr\"{o}m AdS black hole\\
R-charged black hole\\
Kaluza-Klein gauge field\\
%R-charge\\
%no-hair theorem\\
M-branes\\
M-theory\\
%type IIA superstring theory\\
%Stefan-Boltzmann law\\
D$p$-brane\\
deformations
%Klebanov-Strassler geometry\\
%Polchinski-Strassler geometry\\
%Maldacena-Nunez geometry\\
%N=2* geometry@${\cal N}=2^*$ geometry\\
%hyperscaling geometry\\
%holographic superconductors\\
%Lifshitz geometry\\
%scalar polynomial singularity\\
%parallelly propagated singularity\\
%geodesically incomplete\\
%Chern-Simons term\\
%current anomaly
\end{multicols}

\endofsection
