%%%%%%%%%%%%%%%%%%%%%%preface.tex%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% sample preface
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%

\preface

This book describes ``real-world" applications of the AdS/CFT duality for beginning graduate students in particle physics and for researchers in the other fields. 

The AdS/CFT duality is a powerful tool for analyzing strongly-coupled gauge theories using classical gravitational theories. The duality originated from string theory, so it has been actively investigated in particle physics. In recent years, however, the duality has been discussed beyond theoretical particle physics.  In fact, the original AdS/CFT paper by Maldacena has been cited in all physics arXivs. This is because the duality is becoming a powerful tool to analyze the ``real world." 
%Examples are QCD, nuclear physics, nonequilibrium physics, and condensed-matter physics. 
For example, it turns out that one prediction of AdS/CFT is indeed close to the experimental results of the real quark-gluon plasma. Since then, the  duality has been applied to various fields of physics; examples are QCD, nuclear physics, nonequilibrium physics, and condensed-matter physics.

In order to carry out such researches, one has to know many materials such as string theory, general relativity, nuclear physics, nonequilibrium physics, and condensed-matter physics. The aim of this book is to provide these background materials as well as some key applications of the AdS/CFT duality in a single volume. The emphasis throughout the book is on a pedagogical and intuitive approach focusing on the underlying physical concepts. Yet it also includes step-by-step computations for important results which are useful for beginners. Most of them are contained in the appendices.

Following conventions of many textbooks, I often do not refer to original research papers and refer only to the other textbooks and reviews that may be more useful to readers. Also, the choice of references reflects my knowledge, and I apologize in advance for possible omissions.

Initially, this project was begun for a book that was published in Japanese (Saiensu-sha Co., Ltd, 2012), and this is the ``translated" one. But I used this opportunity to improve many explanations and to add more materials to the Japanese edition. So, this book is the ``second edition" in this sense. 

% Springer
I would like to thank many people who helped me with this book. This book is based on review talks at various conferences and on %intensive and regular 
courses I taught at various graduate schools (Tohoku University, Ochanomizu University, Graduate University of Advanced Studies, and Rikkyo University). I thank the organizers and the participants of the conferences and the courses. I also like to thank Elena C\'{a}ceres, Koji Hashimoto, Tetsuo Hatsuda, Gary Horowitz, and Joe Polchinski for encouraging me to write this English edition. 
I  also thank Tetsufumi Hirano, Akihiro Ishibashi, and Takeshi Morita for useful comments and discussion.
%I also thank colleagues for reading parts (or all in some cases) of the draft and for many useful suggestions for improvement. These include...
I especially would like to thank Takashi Okamura who clarified my understanding of the subjects in this book through collaboration for many years. He also gave many suggestions for improvement. I thank my editor Hisako Niko at Springer Japan and the staff of the Lecture Notes in Physics. 
% Springer
Of course, the responsibility for any remaining mistake is solely mine. I will be happy to receive comments on this book. Please send them to \href{mailto:makoto.natsuume@icloud.com}{makoto.natsuume@icloud.com}. 
% arXiv
%I will be happy to receive comments on this book, in particular, conceptual, computational, and English errors. Please send them to makoto.natsuume@icloud.com. 

% Springer
An updated list of corrections will be posted on my website. The tentative address is \url{http://research.kek.jp/people/natsuume/ads-real-world.html}. Even if the address changes in future, you can probably easily search the website because my family name is rather rare. 

I hope that this book will help readers to explore new applications of the AdS/CFT duality.

% Springer
%\vspace{\baselineskip}
%\begin{flushright}\noindent
%Tsukuba, May 2014
%\hfill {\it Makoto Natsuume}\\
%\end{flushright}

%Place(s),\hfill {\it Firstname  Surname}\\
%month year\hfill {\it Firstname  Surname}\\

%%% old %%%

%The AdS/CFT duality is an idea which originated from superstring theory, so it had been actively studied in superstring community. However, in recent years, the AdS/CFT duality has been discussed beyond theoretical particle physics. This is because the duality is becoming a powerful tool to analyze the ``real-world." Examples are QCD, nuclear physics, nonequilibrium physics, and condensed-matter physics. In fact, the original AdS/CFT paper by Maldacena has been cited in all physics arXivs. The AdS/CFT duality claims that one can solve strongly-coupled gauge theories by analyzing curved spacetime. So, the duality has been applied, \eg, to QCD and nuclear physics, where a strongly-coupled gauge theory indeed appears.

%Most of the contents in this book was originally published in Japanese (Saiensu-sha Co., Ltd, 2012) and this book is the ``translated" one although I tried to improve many explanations and add more materials to the Japanese edition. The Japanese edition is based on intensive and regular courses I taught at various graduate schools (Tohoku University, Ochanomizu University, and Graduate University of Advanced Studies). I would like to thank organizers and students of these courses. 

%本書は，大学院生や他分野の研究者向けに書いた超弦理論のAdS/CFT双対性の「応用」についての本である．

%AdS/CFTは，ある意味，素粒子物理でもっとも活発に議論された考えだと言って差しつかえないだろう．素粒子物理学を中心とした論文のデータベースINSPIRE/SPIRESによると\cite{topcites}，歴代論文のなかでもAdS/CFTを提唱した論文 \cite{Maldacena:1997re} が引用件数トップである．この論文は，2008年には，標準模型でCPの破れを説明した小林-益川論文を抑え2位に，2011年には，標準模型の電弱理論を確立したワインバーグ論文を抑えてトップに躍り出た．

%もともとAdS/CFTは超弦理論の考えなので，超弦理論の分野で盛んに議論されてきた．しかし，近年は様相が変わってきており，素粒子論の枠を超えて一般相対論，原子核物理，物性物理，非平衡統計力学などの分野で議論されてきている．それは，AdS/CFTが「現実世界」を解析するうえで強力な手法になりつつあるからである．AdS/CFTは，5次元の曲がった時空の解析によって，4次元の強結合ゲージ理論を解くことができると主張する．このため，実際に強結合ゲージ理論が登場する原子核物理をはじめとした分野で盛んに議論されている．

%このようなAdS/CFTの応用では，超弦理論，一般相対論，原子核理論，物性理論，非平衡統計力学といった多分野にわたる話が出てくる．そこで，本書では初学者でも学びやすいように，これらの分野の基礎を含めて紹介する．また，このように多分野にわたる話では，わからない点があってもなかなか専門家に相談する機会も少ないかもしれない．そこで，初心者がつまずきやすいと思われる点をなるべく丁寧に議論するよう心がけた．一方，教科書というのは，それだけではなかなか実際の研究に進むことができないのが常である．本書では，読者がAdS/CFTの研究を実践できるよう，結果だけを紹介するのではなく，なるべく細かい計算も含め，また（ごく簡単とは言え）なるべく進んだ話題，論文にもふれることにした．本書をもとに，もしもさまざまな分野の方々が新たなAdS/CFTの応用をみつけることができれば幸いである．

%執筆にあたり，草稿を読んで有益なコメントを寄せていただいた岡村隆さん（関西学院大学），中村真さん（京都大学）に感謝します．また本書は，東北大学天文学教室，お茶の水女子大学素粒子論研究室での集中講義がもとになっており，講義に参加していただいた方，講義を企画していただいた方々（二間瀬敏史さん，菅本晶夫さん・曹基哲さん）にも感謝します．とくに，講義のわかりにくい点や疑問点などを多く挙げていただいたお茶の水女子大学の大学院生のみなさん（大野慶子さん，小俣巴芳さん，齋藤麻由美さん，長岡友美さん，穂谷野訓子さん）に感謝します．最後に編集部の伊崎修道さん，平勢耕介さんには，長年にわたってSGCライブラリへの執筆を勧めていただき，また実際の執筆に際してもお世話になりました．感謝いたします． 



