\documentclass[12pt]{article}
\usepackage{amsmath,amssymb,bm}
\usepackage{textcomp,gensymb}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage[comma,authoryear]{natbib}
\usepackage{color}
\usepackage{calrsfs}
\usepackage[english]{babel}
\usepackage{setspace}
\usepackage{gensymb}
\usepackage{xfrac}

\usepackage{amsmath}
\usepackage{amsfonts}


% restart equation numbering after every section:
\numberwithin{equation}{section}

%\usepackage{amsmath,amsthm,amscd,amssymb}
\usepackage[noBBpl,sc]{mathpazo}
\usepackage[papersize={6.8in, 10.0in}, left=.5in, right=.5in, top=.6in, bottom=.9in]{geometry}
\linespread{1.1}
\sloppy
\raggedbottom
\pagestyle{plain}

% these include amsmath and that can cause trouble in older docs.
\input{../helpers/cmrsum}
\input{../helpers/fix-underbrace.tex}

\usepackage[small]{titlesec}
\titlelabel{\thetitle.\quad}

\usepackage{microtype}

% hyperref last because otherwise some things go wrong.
\usepackage{hyperref}
\hypersetup{colorlinks=true
,breaklinks=true
,urlcolor=blue
,anchorcolor=blue
,citecolor=blue
,filecolor=blue
,linkcolor=blue
,menucolor=blue
,linktocpage=true}
\hypersetup{
bookmarksopen=true,
bookmarksnumbered=true,
bookmarksopenlevel=10
}

% make sure there is enough TOC for reasonable pdf bookmarks.
\setcounter{tocdepth}{3}

%\usepackage[dotinlabels]{titletoc}
%\titlelabel{{\thetitle}.\quad}
%\input{../helpers/psu-plain-titles.tex}
%\input{../helpers/psu-sc-headers.tex}
%\input{../helpers/fix-revtex-12.tex}
%\DeclareSymbolFont{CMlargesymbols}{OMX}{cmex}{m}{n}
%\DeclareMathSymbol{\sum}{\mathop}{CMlargesymbols}{"50}
%\pdfbookmark[1]{Introduction}{Introduction}

\def\bbar{{\mathchar'26\mkern-9mu b}}

% Helper functions to enable commenting on the text:
\newcommand{\mike}{\color{blue}}
\newcommand{\michael}{\color{green}}
\newcommand{\michel}{\color{red}}

\begin{document}

\title{Putting probabilities first. \\ How Hilbert space generates and constrains them.\footnote{This preprint, which we originally posted in October 2019, has since been further developed into the book: \emph{Understanding Quantum Raffles: Quantum Mechanics on an Informational Approach: Structure and Interpretation} (Springer, Cham, 2021, ISBN: \href{https://www.springer.com/gp/book/9783030859381}{978-3-030-85938-1}). The following are the most notable changes between this preprint and the book: Section 5 (now Chapter 6 of \emph{Understanding Quantum Raffles}), especially Section 5.5 on measurement, has been largely rewritten and greatly expanded. The upshot of that Section has also been strengthened: Chapter 6 of \emph{Understanding Quantum Raffles} defends the view that \emph{both} what we call the ``small'' and the ``big'' measurement problems are not bugs but features of quantum mechanics. Section 1 has also been re-written and expanded in the book (where it is now Chapter 1) and parts of Section 2 (which has now been refactored into two separate chapters) have been rewritten to be more easily digestible by general readers. Section 6 (Chapter 7 in the book) now includes a comparison of our interpretation of quantum mechanics with the Everett interpretation (which like our interpretation adds nothing to the quantum-mechanical formalism). Finally, \emph{Understanding Quantum Raffles} includes a foreword by Jeffrey Bub.}$~\;^{\raisebox{-.1em}{\normalsize ,}}\,$\footnote{Dedicated to the memory of Itamar Pitowsky (1950--2010) and William Demopoulos (1943--2017).}}
%Polytopes and elliptopes: classical and quantum constraints on correlations

\author{Michael Janas,\footnote{School of Physics and Astronomy, University of Minnesota; E-mail: jana0030@umn.edu, janss011@umn.edu} \, Michael E. Cuffaro,\footnote{Rotman Institute of Philosophy, University of Western Ontario; Munich Center for Mathematical Philosophy, Ludwig-Maximilians-Universit\"at M\"unchen; E-mail: mike@michaelcuffaro.com} \, Michel Janssen$^\dagger$}
\date{}

\maketitle


\begin{abstract}
\noindent 
We use correlation arrays, the workhorse of Bub's (2016) \emph{Bananaworld}, to analyze the correlations found in an experimental setup due to \citet{Mermin 1981} for measurements on the singlet state of a pair of spin-$\frac12$ particles. Adopting an approach pioneered by \citet{Pitowsky 1989b} and promoted in \emph{Bananaworld}, we geometrically represent the class of correlations allowed by quantum mechanics in this setup as an elliptope in a non-signaling cube.  To determine which of these quantum correlations are allowed by local hidden-variable theories, we investigate which ones we can simulate using raffles with baskets of tickets that have the outcomes for all combinations of measurement settings printed on them. The class of correlations found this way can be represented geometrically by a tetrahedron contained within the elliptope. We use the same Bub-Pitowsky framework to analyze a generalization of the Mermin setup for measurements on the singlet state of two particles with higher spin. The class of correlations allowed by quantum mechanics in this case is still represented by the elliptope; the subclass of those whose main features can be simulated with our raffles can be represented by polyhedra that, with increasing spin, have more and more vertices and facets and get closer and closer to the elliptope. We use these results to advocate for Bubism (not to be confused with QBism), an interpretation of quantum mechanics along the lines of \emph{Bananaworld}. Probabilities and expectation values are primary in this interpretation. They are determined by inner products of vectors in Hilbert space. Such vectors do not themselves represent what is real in the quantum world. They encode families of probability distributions over values of different sets of observables. As in classical theory, these values ultimately represent what is real in the quantum world. Hilbert space puts constraints on possible combinations of such values, just as Minkowski space-time puts constraints on possible spatio-temporal constellations of events. Illustrating how generic such constraints are, the constraint derived in this paper, the equation for the elliptope, is a general constraint on correlation coefficients that can be found in older literature on statistics and probability theory. \citet{Yule 1896} already stated the constraint. De Finetti (1937) already gave it a geometrical interpretation sharing important features with its interpretation in Hilbert space.
\end{abstract}

\noindent
\emph{Keywords}: Correlation arrays, correlation polytopes, Bubism, Bell inequalities, Tsirelson bound, Born rule, kinematical constraints, principle theories, measurement.

\tableofcontents

%SECTION 1 (label: 0)
\section{Introduction}\label{0}
\input{Section1}

%SECTION 2 (labels: 1, 1.1 through 1.6)
\section{Representing distant correlations by correlation arrays and polytopes} \label{1}
\input{Section2}
%SECTION 2.5
\input{Section2.5}
%SECTION 2.6
\input{Section2.6}

%SECTION 3 (label: 3}
\section{Generalization to the singlet state of two particles with higher spin} \label{2}
\input{Section3-intro}

%SUBSECTION 3.1 (labels: 3.1, 3.1.1 through 3.1.X)
\subsection{The quantum correlations} \label{2.1}
\input{Section3.1-Quantum}

%SUBSECTION 3.2 (labels: 3.2, 3.2.1 through 3.2.4)
\subsection{Designing raffles to simulate the quantum correlations} \label{2.2}
\input{Section3.2-classical}
%SECTION 3.2.4
\input{Section3.2.4-classical}

%SECTION 4
\section{Correlation arrays, polytopes and the CHSH inequality} \label{3}
\input{Section4-CHSH}

%SECTION 5
\section{Interpreting quantum mechanics} \label{4}
\input{Section5}

%SECTION 6
\section{Conclusion} \label{5}
\input{Conclusion}


\section*{Acknowledgments}

Thanks, above all, to Jeff Bub for his enthusiastic support of our efforts and for his patience in explaining his information-theoretic interpretation of quantum mechanics to us. The three of us came to this project from different directions. Janssen (a historian of science) is a recovering Everettian who has been defending Bub's information-theoretic interpretation with the zeal of the converted. Cuffaro (a philosopher of science) has been working on information-theoretic interpretations for some time and has tried to keep Janssen's proselytizing at bay. Janas (a theoretical physicist) was and remains a Bohm sympathizer but got intrigued by Pitowsky's polytopes. One thing the three of us share is a broadly Kantian outlook. This paper has been in the making for some time and we have presented preliminary results in a variety of settings. We thank Laurent Taudin for drawing chimps and bananas both for more recent presentations and for this paper. For their feedback on our presentations, we are grateful to members of the Physics Interest Group (PIG) of the Minnesota Center for Philosophy of Science; to students at the University of Minnesota and Washburn High in Minneapolis; to audiences at a physics colloquium at Minnesota State University Mankato, a lunchtime talk at the Center for Philosophy of Science at the University of Pittsburgh, and seminars in the School of Mathematics at the University of Minnesota and at the \emph{Max-Planck-Institut f\"ur Wissenschaftsgeschichte} in Berlin; to participants in a workshop on interpreting quantum mechanics at \emph{Politecnico di Milano}; and, finally, to participants in the 2019 edition of the conference series, \emph{New Directions in the Foundations of Physics}, in Viterbo, Italy. Especially Wayne Myrvold's intervention in the Q\&A following our presentation in Viterbo, our only joint performance, greatly improved this paper (see notes \ref{Myrvold 1} and \ref{Myrvold 2} in Section \ref{1.6}). For further discussion we thank Tony Duncan, Sam Fletcher, Louisa Gilder, Peter Grul, Luc Janssen, Christian Joas, David Kaiser, Jim Kakalios, Alex Kamenev, Marius Krumm, Christoph Lehner, Charles Marcus, Markus M\"uller, Max Niedermayer, Sergio Pernice, J\"urgen Renn, Serge Rudaz, David Russell, Rob ``Ryno'' Rynasiewicz, Rob Spekkens and Brian Woodcock.  Instead of dedicating our paper to them we would have loved to discuss it with Bill Demopoulos and Itamar Pitowsky. Cuffaro gratefully acknowledges support from the Rotman Institute of Philosophy and from the Foundational Questions Institute (FQXi). Janssen gratefully acknowledges support from the \emph{Alexander von Humboldt Stiftung} and the \emph{Max-Planck-Institut f\"ur Wissenschaftsgeschichte}. Janas thanks the staff of Al's Breakfast in Dinkytown.

\renewcommand{\refname}{Bibliography}

\begin{thebibliography}{}

\bibitem[Accardi and Fedullo(1982)]{Accardi and Fedullo 1982}  Accardi, L.\ and Fedullo, A.\ (1982). On the statistical meaning of complex numbers in quantum mechanics. \emph{Lettere al Nuovo Cimento} 34: 161--172.

\bibitem[Acu\~na(2014)]{Acuna 2014} Acu\~na, P.\ (2014). On the empirical equivalence between special relativity and Lorentz's ether theory. {\em Studies in History and Philosophy of Modern Physics} 46: 283--302.

\bibitem[Albert(1992)]{Albert 1992} Albert, D.\ Z.\ (1992). \emph{Quantum Mechanics and Experience.} Cambridge, MA: Harvard University Press.

\bibitem[Albert(2018)]{Albert 2018} Albert, D.\ Z.\ (2018). Quantum's leaping lizards (Review of \citet{Becker 2018}) \emph{The New York Review of Books} 65, no.\ 7 (April 19, 2018).

\bibitem[Aldrich(1998)]{Aldrich 1998} Aldrich, J.\ (1998). Doing least squares: perspectives from Gauss and Yule. \emph{International Statistical Review} 66: 61--81.

\bibitem[Bacciagaluppi(2017)]{Bacciagaluppi 2017} Bacciagaluppi, G.\ (2017). Bohr's slit and Hermann's microscope. In G.\ Bacciagaluppi and E.\ Crull, Eds., \emph{Grete Hermann: Between Physics and Philosophy} (pp. 135--148). Dordrecht: Springer.

\bibitem[Bacciagaluppi, Crull and Maroney(2017)]{BCM 2017} Bacciagaluppi, G., Crull, E.\ and Maroney, O.\ J.\ E.\ (2017). Jordan's derivation of blackbody fluctuations. \emph{Studies in History and Philosophy of Modern Physics} 60: 23--34.

\bibitem[Balashov and Janssen(2003)]{Balashov  and Janssen 2003} Balashov, Y.\ and Janssen, M.\ (2003). Presentism and relativity.  {\it British Journal for the Philosophy of Science}
 54: 327--346.

\bibitem[Ball(2018)]{Ball 2018} Ball, P.\ (2018). \emph{Beyond Weird: Why Everything You Thought You Knew About Quantum Physics is Different}. London: The Bodley Head.

%\bibitem[Barrett \emph{et al.}(2005)]{Barrett 2005} Barrett, J., Linden, N., Massar, S., Pironio, S., Popescu, S., and Roberts, D.\ (2005). Non-local correlations as an information theoretic resource. \emph{Physical Review} A 71: 022101.
%xxx--yyy see bananaworld, p. 120, note 2.

\bibitem[Baym(1969)]{Baym 1969} Baym, G.\ (1969). \emph{Lectures on Quantum Mechanics.} Reading, MA: Addison-Wesley.

\bibitem[Becker(2018)]{Becker 2018} Becker, A.\ (2018). \emph{What is Real? The Unfinished Quest for the Meaning of Quantum Physics}. New York: Basic Books.

\bibitem[Bell(1964)]{Bell 1964} Bell, J.\ S.\ (1964). On the Einstein-Podolsky-Rosen paradox. \emph{Physics} 1: 195--200. Reprinted in \citet[pp.\ 14--21]{Bell 1987}.

\bibitem[Bell(1984)]{Bell 1984} Bell, J.\ S.\ (1984). Beables for quantum field theory. CERN-TH.4035/84. Reprinted in \citet[pp.\ 159--166]{Bell 1987}

\bibitem[Bell(1987)]{Bell 1987} Bell, J.\ S.\ (1987). \emph{Speakable and Unspeakable in Quantum Mechanics}. Cambridge: Cambridge University Press. Page references to second edtion (2004).

\bibitem[Bell(1990)]{Bell 1990} Bell, J.\ S.\ (1990). Against measurement. \emph{Physics World} 8: 33--40.

\bibitem[Ben-Menahem(2017)]{Ben-Menahem 2017} Ben-Menahem, Y.\ (2017). The PBR theorem: Whose side is it on? \emph{Studies in History and Philosophy of Modern Physics} 57: 80--88.

\bibitem[Berkovitz(2012)]{Berkovitz 2012} Berkovitz, J.\ (2012). The world according to de {F}inetti. In Y.\ Ben-Menahem and M.\ Hemmo, Eds., \emph{Probability in Physics} (pp. 249--280). Heidelberg: Springer.

\bibitem[Berkovitz(2019)]{Berkovitz 2019} Berkovitz, J.\ (2019). On de {F}inetti's instrumentalist philosophy of probability. \emph{European Journal for the Philosophy of Science} 9: 25.

\bibitem[Bingham(2000)]{Bingham 2000} Bingham, N.\ H.\ (2000). Studies in the history of probability and statistics XLVI. Measure into probability: From Lebesgue to Kolmogorov. \emph{Biometrika} 87: 145--156.

\bibitem[Birkhoff and Kreyszig(1984)]{Birkhoff and Kreyszig 1984} Birkhoff, G.\ and Kreyszig, E.\ (1984). The establishment of functional analysis. \emph{Historia Mathematica} 11: 258--321.

\bibitem[Bohr(1935)]{Bohr 1935} Bohr, N.\ (1935). Can quantum-mechanical description of physical reality be considered complete? \emph{Physical Review} 48: 696--702.

\bibitem[Bohr(1937)]{Bohr 1937} Bohr, N.\ (1937). Causality and complementarity. \emph{Philosophy of Science} 4: 289--298.

\bibitem[Bohr(1958)]{Bohr 1958} Bohr, N.\ (1958). Quantum physics and philosophy. In R.\ Klibansky, Ed., \emph{Philosophy in the Mid-Century: A Survey} (pp.\ 308--314). Firenze: La Nuova Italia Editrice. Page reference to reprint in: J.\ Kalckar, Ed., \emph{Niels Bohr: Collected Works}, Vol.\ 7 (pp.\ 385--394). Amsterdam: Elsevier.

%\bibitem[Bohm(1951)]{Bohm 1951} Bohm, D.\ (1951). \emph{Quantum Theory.} Englewood Cliffs, NJ: Prentice-Hall.
%pp. 614--623

%\bibitem[Born(1971)]{Born 1971} Born, M., Ed.\  (1971). \emph{The Born-Einstein Letters.} New York: Walker.

\bibitem[Born, Heisenberg and Jordan(1926)]{dreimaenner} Born, M., Heisenberg, W.\ and Jordan, P.\ (1926). Zur Quantenmechanik II. \emph{Zeitschrift f\"{u}r Physik}  35: 557--615. English translation in \citet[pp.\ 321--385]{Van der Waerden}.

\bibitem[Bowler(2003)]{Bowler 2003} Bowler, P.\ J.\ (2003). \emph{Evolution. The History of an Idea}. 3rd revised and expanded edition. Berkeley, Los Angeles, London: University of California Press.

\bibitem[Bravais(1846)]{Bravais 1846} Bravais, A.\ (1846). Analyse math\'ematique sur les probabilit\'es des erreurs de situation d'un point. \emph{M\'emoires de l'Acad\'emie Royale des Sciences de l'Institut de France} 9: 255--332.

\bibitem[Brown(2005)]{Brown 2005} Brown, H.\ R.\ (2005). \emph{Physical Relativity. Space-time Structure from a Dynamical Perspective}, Oxford: Oxford University Press.

\bibitem[Brown and Timpson(2006)]{Brown and Timpson 2006}  Brown, H.\ R.\ and Timpson, C.\ G.\ (2006). Why special relativity should not be a template for a fundamental reformulation of quantum mechanics. In W.\ Demopoulos and I.\ Pitowsky, Eds., \emph{Physical Theory and its Interpretation} (pp.\ 29--42). Berlin: Springer.  

\bibitem[Bub(2004)]{Bub 2004} Bub, J.\ (2004). Why the quantum? \emph{Studies in History and Philosophy of Modern Physics} 35: 241--266.

\bibitem[Bub(2010)]{Bub 2010} Bub, J.\ (2010). Von Neumann's `no hidden variables' proof: A re-appraisal.  \emph{Foundations of Physics} 40: 1333--1340.

\bibitem[Bub(2016)]{Bub 2016} Bub, J.\ (2016). \emph{Bananaworld. Quantum Mechanics for Primates}. Oxford: Oxford University Press. Slightly revised paperback edition: 2018.

\bibitem[Bub(2019)]{Bub 2019} Bub, J.\ (2019). Interpreting the quantum world: old questions and new answers (essay review of \citet{Freire 2015}, \citet{Becker 2018} and \citet{Ball 2018}). \emph{Historical Studies in the Natural Sciences} 49: 226--239.

\bibitem[Bub and Demopoulos(2010)]{Bub and Demopoulos 2010} Bub, J.\ and Demopoulos, W.\ (2010). Itamar Pitowsky 1950--2010. \emph{Studies in History and Philosophy of Modern Physics} 41: 85.

\bibitem[Bub and Pitowsky(2010)]{Bub and Pitowsky 2010} Bub, J.\ and Pitowsky, I.\ (2010). Two dogmas about quantum mechanics.  In \citet[pp.\ 433--459]{Many Worlds 2010}.

\bibitem[Bub and Bub(2018)]{Bub and Bub 2018} Bub, T.\ and Bub, J.\ (2018). \emph{Totally Random. Why Nobody Understands Quantum Mechanics. A Serious Comic on Entanglement.} Princeton: Princeton University Press.

\bibitem[Busch(2003)]{Busch 2003} Busch, P.\ (2003). Quantum states and generalized observables: A simple proof of Gleason's theorem. \emph{Physical Review Letters} 91: 120403.

\bibitem[Butler(1968)]{Butler 1968} Butler, J.\ W.\ (1968). On the Trouton-Noble experiment. \emph{American Journal of Physics} 36: 936--941.

\bibitem[Cabello(2019)]{Cabello 2019} Cabello, A.\ (2019). Quantum correlations from simple assumptions. \emph{Physical Review A} 100: 032120.

\bibitem[Caulton(2015)]{Caulton 2015} Caulton, A. (2015). The role of symmetry in the interpretation of physical theories. \emph{Studies in History and Philosophy of Modern Physics} 52: 153--162.

\bibitem[Chiribella, D'Ariano and Perinotti(2010)]{Chiribella et al 2010} Chiribella, J., D'Ariano, G.\ M.\ and Perinotti, P.\  (2010). Probabilistic theories with purification. \emph{Physical Review A} 81:
062348.

\bibitem[Chiribella and Ebler(2019)]{Chiribella and Ebler 2019} Chiribella, J.\ and Ebler, D.\, Quantum speedup in the identification of cause-effect relations. \emph{Nature Communications} 10: 1472.

\bibitem[Chiribella and Spekkens(2016)]{Chiribella and Spekkens 2016} Chiribella, G.\ and Spekkens, R.\ (2016). \emph{Quantum Theory: Informational Foundations and Foils}. Dordrecht: Springer.

\bibitem[Cirel'son(1980)]{Cirel'son 1980} Cirel'son, B.\ S.\ (1980). Quantum generalizations of Bell's inequality. \emph{Letters in Mathematical Physics} 4: 93--100.

\bibitem[Clauser and Freedman(1972)]{Clauser and Freedman 1972} Clauser, J.\ F.\ and Freedman, S.\ J.\ (1972). Experimental test of local hidden-variable theories. \emph{Physical Review Letters} 28: 938--941.

\bibitem[Clauser, Horne, Shimony and Holt(1969)]{CHSH} Clauser, J.\ F., Horne, M.\ A., Shimony, A.\ and Holt, R.\ A.\ (1969). Proposed experiment to test local hidden-variable theories. \emph{Physical Review} 23: 880--884.

\bibitem[Clifton, Bub and Halvorson(2003)]{CBH} Clifton, R., Bub, J.\ and Halvorson, H.\ (2003). Characterizing quantum theory in terms of information-theoretic constraints. \emph{Foundations of Physics} 33: 1561--1591.

\bibitem[Cuffaro(2018)]{Cuffaro 2018} Cuffaro, M.\, E.\ (2018). Information causality, the Tsirelson bound, and the `being-thus' of things. \emph{Studies in History and Philosophy of Modern Physics}. In press.

\bibitem[Cuffaro and Fletcher(2018)]{Cuffaro and Fletcher 2018} Cuffaro, M.\ E.\  and Fletcher, S.\ C., Eds.\ (2018). \emph{Physical Perspectives on Computation, Computational Perspectives on Physics.} Cambridge: Cambridge University Press.

\bibitem[Curiel(2019)]{Curiel 2019} Curiel, E.\ (2019). Schematizing the observer and the epistemic content of theories. Preprint, {arXiv:1903.02182}.

\bibitem[Dantzig and Thapa(1997)]{Dantzig and Thapa 1997}  Dantzig, G.\ B.\ and Thapa, M.\ N.\ (1997). \emph{Linear programming 1: Introduction}. New York: Springer.

\bibitem[Dantzig and Thapa(2003)]{Dantzig and Thapa 2003} Dantzig, G.\ B.\ and Thapa, M.\ N.\ (2003). \emph{Linear programming 2: Theory and Extensions.} New York: Springer.

\bibitem[De Finetti(1930)]{De Finetti 1930} De Finetti, B.\ (1930). Spazi astratti metrici ($\mathcal{D}_M$). \emph{Atti della Pontificia Accademia delle Scienze Nuovi Lincei} 83: 248--256.

\bibitem[De Finetti(1937)]{De Finetti 1937} De Finetti, B.\ (1937). A proposito di ``correlazione." \emph{Supplemento Statistico ai Nuovi problemi di Politica, Storia ed Economia}, 3:  41--57. Page reference to English translation in \emph{Journal \'Electronique d'Histoire des Probabilit\'es et de la Statistique} (www.jehps.net) 4 (4) (2008): 1--15.

\bibitem[Demopoulos(2010)]{Demopoulos 2010}  Demopoulos, W.\ (2010). Effects and propositions. \emph{Foundations of Physics} 40: 368--389.

\bibitem[Demopoulos(2018)]{Demopoulos 2018} Demopoulos, W.\ (2018). \emph{On Theories}. Unpublished book manuscript.

\bibitem[Deza and Laurent(1997)]{Deza and Laurent 1997} Deza, M.\ M.\ and Laurent, M.\  (1997). \emph{Geometry of Cuts and Metrics}. Berlin: Springer.

\bibitem[Dieks(2017)]{Dieks 2017} Dieks, D.\ (2017). Von Neumann's impossibility proof: Mathematics in the service of rhetorics. \emph{Studies in History and Philosophy of Modern Physics} 60: 136--148.

\bibitem[Dirac(1926)]{Dirac 1926} Dirac, P.\ A.\ M.\ (1926). The fundamental equations of quantum mechanics. \emph{Proceedings of the Royal Society of London} A109: 642--653. Reprinted in \citet[pp.\ 307--320]{Van der Waerden}

\bibitem[Dirac(1927)]{Dirac 1927} Dirac, P.\ A.\ M.\  (1927).  The physical interpretation of the quantum dynamics. \emph{Proceedings of the Royal Society of London. Series A} 113: 621--641. 

\bibitem[Dirac(1930)]{Dirac 1930} Dirac, P.\ A.\ M.\  (1930). \emph{Principles of Quantum Mechanics.} Oxford: Clarendon.

\bibitem[Doob(1934)]{Doob 1934} Doob, J.\ L.\ (1934). Probability and statistics. \emph{Transactions of the American Mathematical Society} 36: 759--775.

\bibitem[Duncan and Janssen(2007)]{Duncan and Janssen 2007} Duncan, A.\ and Janssen, M.\  (2007). On the verge of \emph{Umdeutung} in Minnesota: Van Vleck and the correspondence principle. 2 Pts. \emph{Archive for History of Exact Sciences} 61: 553--624, 625--671.

\bibitem[Duncan and Janssen(2008)]{Duncan and Janssen 2008} Duncan, A.\ and Janssen, M.\   (2008). Pascual Jordan's resolution of the conundrum of the wave-particle duality of light. \emph{Studies in History and Philosophy of Modern Physics} 39: 634--666.

\bibitem[Duncan and Janssen(2013)]{Duncan and Janssen 2013} Duncan, A.\ and Janssen, M.\  (2013). (Never) mind your $p$'s and $q$'s: Von Neumann versus Jordan on the foundations of quantum theory. \emph{The European Physical Journal H} 38: 175--259.

\bibitem[Duncan and Janssen(2014)]{Duncan and Janssen 2014} Duncan, A.\ and Janssen, M.\  (2014). The trouble with orbits: the Stark effect in the old and the new quantum theory. \emph{Studies in History and Philosophy of Modern Physics} 48: 68--83.

\bibitem[Duncan and Janssen(2015)]{Duncan and Janssen 2015} Duncan, A.\ and Janssen, M.\  (2015).  The Stark effect in the Bohr-Sommerfeld theory and in Schr\"odinger's wave mechanics. In F.\ Aaserud and H.\ Kragh, Eds., \emph{One Hundred Years of the Bohr Atom} (pp. 217--271). Copenhagen: Det Kongelige Danske Videnskabernes Selskab.

\bibitem[Duncan and Janssen(2019)]{Duncan and Janssen 2019} Duncan, A.\ and Janssen, M.\  (2019). \emph{Constructing Quantum Mechanics}, Vol.\ 1, \emph{The Scaffold, 1900--1923}. Oxford: Oxford University Press.

\bibitem[Eckert(2013)]{Eckert 2013} Eckert, M.\ (2013). \emph{Arnold Sommerfeld. Science, Life and Turbulent Times.} New York: Springer,.

\bibitem[Ehrenfest(1925)]{Ehrenfest 1925} Ehrenfest, P.\ (1925). Energieschwankungen im Strahlungsfeld oder Kristallgitter bei Superposition quantisierter Eigenschwingungen. \emph{Zeitschrift f\"ur Physik}, 34, 362--373.

\bibitem[Einstein(1909a)]{Einstein 1909a} Einstein, A.\ (1909a). Zum gegenw\"artigen Stand des Strahlungsproblems. \emph{Physikalische Zeitschrift} 10: 185--193. Reprinted in Einstein (1987--2018, Vol.\ 2, Doc.\ 56).

\bibitem[Einstein(1909b)]{Einstein 1909b} Einstein, A.\ (1909b). \"Uber die Entwicklung unserer Anschauungen \"uber das Wesen und die Konstitution der Strahlung. \emph{Physikalische Zeitschrift} 10: 817--825. Reprinted in Einstein (1987--2018, Vol.\ 2, Doc.\ 60).

\bibitem[Einstein(1919)]{Einstein 1919} Einstein, A.\ (1919). Time, space, and gravitation. \emph{The London Times}, November 28, 1919. Reprinted in Einstein (1987--2018, Vol.\ 7, Doc.\ 26).
%M.\ Janssen \emph{et al}., Eds. \emph{The Collected Papers of Albert Einstein}. Vol.\ 7. \emph{The Berlin Years: Writings, 1918--1921} (Doc.\ 26). Princeton: Princeton University Press, 2002. 
Reprinted under a different title (What is the theory of relativity?) in A.\ Einstein, \emph{Ideas and Opinions} (pp.\ 227--232). New York: Crown Publishers, 1954.

\bibitem[Einstein(1987--2018)]{Einstein 1987--2018} Einstein, A.\ (1987--2018). \emph{The Collected Papers of Albert Einstein} (15 Vols).  J.\ Stachel, \emph{et al.} (Eds.). Princeton: Princeton University Press.

\bibitem[Einstein, Podolsky and Rosen(1935)]{EPR} Einstein, A., Podolsky, B.\ and Rosen, N.\ (1935). Can quantum-mechanical description of physical reality be considered complete? \emph{Physical Review} 47: 777--780.

\bibitem[Eisberg and Resnick(1985)]{Eisberg and Resnick 1985} Eisberg, R.\ and Resnick, R.\ (1985). \emph{Quantum Physics of Atoms, Molecules, Solids, Nuclei, and Particles}. 2nd ed.\ (1st ed.: 1974). New York: Wiley.   

\bibitem[Epstein(1916)]{Epstein 1916} Epstein, P.\ (1916). Zur Theorie des Starkeffektes. \emph{Annalen der Physik} 50: 489--521.

\bibitem[Epstein(1926)]{Epstein 1926} Epstein, P.\ (1926). The Stark effect from the point of view of Schr\"odinger's quantum theory. \emph{Physical Review} 28: 695--710.

%\bibitem[Fine(1982a)]{Fine 1982a}  Fine, A.\ (1982a). Hidden variables, joint probability, and the Bell inequalities. \emph{Physical Review Letters} 48:
% (5): 291--294.

%\bibitem[Fine(1982b)]{Fine 1982b}  Fine, A.\ (1982b). Joint distributions, quantum correlations, and commuting observables. \emph{Journal of Mathematical Physics}  23: 
%(7): 1306--1310.

\bibitem[Fisher(1915)]{Fisher 1915}  Fisher, R.\ A.\ (1915). Frequency distribution of the values of the correlation coefficient in samples from an indefinitely large population. \emph{Biometrika} 10: 507--521.

\bibitem[Fisher(1924)]{Fisher 1924}  Fisher, R.\ A.\ (1924). The distribution of the partial correlation coefficient. \emph{Metron} 3: 329--332. 

\bibitem[Forster(1942)]{Forster 1942} Forster, E.\ M.\ (1942). \emph{Virginia Woolf. The Rede Lecture 1941}. Cambridge: Cambridge University Press.

\bibitem[Franklin \emph{et al.}(2008)]{Franklin et al 2008}  Franklin, A., Edwards, A.\ W.\ F., Fairbanks, D.\ J., Hartl, D.\ L.\ and Seidenfeld, T.\ (2008).  \emph{Ending the Mendel-Fisher Controversy}. Pittsburgh: University of Pittsburgh Press.

\bibitem[Frauchiger and Renner(2018)]{Frauchiger and Renner 2018} Frauchiger, D.\ and Renner, R.\ (2018). Quantum theory cannot
consistently describe the use of itself. \emph{Nature Communications} 9: 3711.

\bibitem[Freire(2015)]{Freire 2015} Freire, O.\ Jr.\ (2015). \emph{The Quantum Dissidents: Rebuilding the Foundations of Quantum Mechanics (1950--1990)}. Heidelberg: Springer.

\bibitem[Fristedt and Gray(1997)]{Fristedt and Gray 1997}  Fristedt, B.\ E.\ and Gray, L.\ F.\ (1997). \emph{A Modern Approach to Probability Theory}. Boston, Basel, Berlin: Birkh\"auser. 
%Fristedt, Bert E., Gray, Lawrence F. 

\bibitem[Ghirardi(2018)]{Ghirardi 2018} Ghirardi, G.\ (2018). Collapse Theories. In E.\ N.\ Zalta (ed.), \emph{The Stanford Encyclopedia of Philosophy} (Fall 2018 Edition). URL =
\textless\url{plato.stanford.edu/archives/fall2018/entries/qm-collapse/}\textgreater

\bibitem[Gilder(2008)]{Gilder 2008} Gilder, L.\ (2008). \emph{The Age of Entanglement. When Quantum Physics Was Reborn.} New York: Knopf. Reprinted: New York: Vintage, 2009.

\bibitem[Giovanelli(2014)]{Giovanelli 2014} Giovanelli, M.\ (2014). `But one must not legalize the mentioned sin': Phenomenological vs.\ dynamical treatments of rods and clocks in Einstein's thought. \emph{Studies in History and Philosophy of Modern Physics} 48: 20--44.

\bibitem[Gleason(1957)]{Gleason 1957} Gleason, A.\ M.\ (1957). Measures on the closed subspaces of a Hilbert space. \emph{Journal of Mathematics and Mechanics} 6: 885--893.

\bibitem[Greenberger, Horne and Zeilinger(2014)]{GHZ} Greenberger, D.\ M., Horne, M.\ A.\ and Zeilinger, A. (1998). Going beyond Bell's theorem. In \citet[69--72]{Kafatos 1989}

%\bibitem[Griffiths(2005)]{Griffiths 2005} Griffiths, D.~J.\ (2005). \emph{Introduction to Quantum Mechanics}. 2nd ed. Upper Saddle River, NJ: Pearson Prentice Hall.

\bibitem[Hald(1998)]{Hald 1998} Hald, A.\ (1998).\emph{A History of Mathematical Statistics from 1750 to 1930}. New York: Wiley.

\bibitem[Handsteiner \emph{et al.}(2017)]{Handsteiner 2017} Handsteiner, J.\ \emph{et al.} (2017). Cosmic Bell test: measurement settings from milky way stars. \emph{Physical Review Letters} 118: 060401-1--7.

\bibitem[Hardy(2007)]{Hardy 2007} Hardy, L.\ (2007). Towards quantum gravity: A framework for probabilistic theories with non-fixed causal structure. \emph{Journal of Physics A} 30: 3081.

\bibitem[Hardy(2012)]{Hardy 2012} Hardy, L.\ (2012). The operator tensor formulation of quantum theory. \emph{Philosophical Transactions of the Royal Society A} 370: 3385--3417.

\bibitem[Harrigan and Spekkens(2010)]{Harrigan and Spekkens 2010} Harrigan, N.\ and Spekkens, R.\ W.\ (2010). Einstein, incompleteness, and the epistemic view of quantum states. \emph{Foundations of Physics} 40: 125--157.

\bibitem[Healey(2016)]{Healey 2016} Healey, R.\ (2016). A pragmatist view of the metaphysics of entanglement. \emph{Synthese} (in press). Advance access version available via \emph{Springer Netherlands online}.

\bibitem[Healey(2019)]{Healey 2019} Healey, R.\ (2019). The mysteries of quantum mechanics in graphic-novel form (review of Bub and Bub, 2018). \emph{Physics Today}  72 (9): 56--57. 

\bibitem[Heisenberg(1925)]{Heisenberg 1925} Heisenberg, W.\ (1925). \"Uber die quantentheoretische Umdeutung kinematischer und mechanischer Beziehungen. \emph{Zeitschrift f\"ur Physik} 33: 879--893. English translation in \citet[pp.\ 261--276]{Van der Waerden}.

%B.\ L.\ van der Waerden, Ed.\ (1968). \emph{Sources of Quantum Mechanics} (pp.\ 261--276) New York: Dover.

\bibitem[Janssen(2009)]{Janssen 2009} Janssen, M.\ (2009). Drawing the line between kinematics and dynamics in special relativity. \emph{Studies in History and Philosophy of Modern Physics} 40: 26--52.

\bibitem[Janssen(2019)]{Janssen 2019} Janssen, M.\ (2019). Arches and scaffolds: bridging continuity and discontinuity in theory change. In A.\ C.\ Love and W.\ C.\ Wimsatt, Eds., \emph{Beyond the Meme. Development and Structure in Cultural Evolution.} (pp.\ 95--199). Minneapolis: University of Minnesota Press.

\bibitem[Janssen and Lehner(2014)]{Janssen and Lehner 2014} Janssen, M.\ and Lehner, C., Eds.\ (2014). \emph{The Cambridge Companion to Einstein}. Cambridge: Cambridge University Press.

\bibitem[Joas and Lehner(2009)]{Joas and Lehner 2009} Joas, C.\ and Lehner, C.\ (2009). The classical roots of wave mechanics: Schr\"odinger's transformations of the optical-mechanical analogy. \emph{Studies in History and Philosophy of Modern Physics} 40: 338--351.

\bibitem[Jordan(1927a)]{Jordan 1927a} Jordan, P.\ (1927a).  \"Uber eine neue Begr\"undung der Quantenmechanik. \emph{Zeitschrift f\"ur Physik} 40: 809--838. 

\bibitem[Jordan(1927b)]{Jordan 1927b} Jordan, P.\  (1927b).  \"Uber eine neue Begr\"undung der Quantenmechanik II. \emph{Zeitschrift f\"ur Physik} 44: 1--25.

\bibitem[Kafatos(1989)]{Kafatos 1989} Kafatos, M., Ed., (1989).  \emph{Bell's Theorem, Quantum Theory and Conceptions of the Universe}. Dordrecht: Kluwer. 

\bibitem[Kaiser(2011)]{Kaiser 2011} Kaiser, D.\ (2011). \emph{How the Hippies Saved Physics. Science, Counterculture, and the Quantum Revival.} New York: Norton.

\bibitem[Kakalios(2009)]{Kakalios 2009} Kakalios, J.\ (2009). \emph{The Physics of Superheroes}. 2nd ed. New York: Gotham Books

\bibitem[Kakalios(2018)]{Kakalios 2018} Kakalios, J.\ (2018). A graphic tale of entanglement (review of Bub and Bub, 2018). \emph{Physics World} 31 (12): 37--38. 

\bibitem[Kendall(1961)]{Kendall 1961} Kendall, M.~G.\ (1961). \emph{A Course in the Geometry of $n$ Dimensions.} New York: Hafner. 

\bibitem[Koberinski and M\"uller(2018)]{Koberinski and Mueller 2018} Koberinski, A.\ and M\"uller, M.\ (2018). Quantum theory as a principle theory: Insights from an information theoretic reconstruction. In \citet[pp. 257--281]{Cuffaro and Fletcher 2018}. 

\bibitem[Kramers and Heisenberg(1925)]{Kramers and Heisenberg 1925} Kramers, H.\ A.\ and Heisenberg, W.\ (1925). \"{U}ber die Streuung von Strahlung durch Atome. \emph{Zeitschrift f\"{u}r Physik} 31: 681--707. English translation in \citet[pp. 223--252]{Van der Waerden}.

\bibitem[Landau(1988)]{Landau 1988} Landau, L.\ J.\ (1988). Empirical two-point correlation functions. \emph{Foundations of Physics} 18: 440--460.

\bibitem[Lassez and Lassez(1992)]{Lassez and Lassez 1992} Lassez, C.\ and Lassez, J.-L.\ (1992). Quantifier elimination for conjunctions of linear constraints via a convex hull algorithm. In B.\ R.\ Donald, D.\ Kapur, J.\ L.\ Mundy, Eds., \emph{Symbolic and Numerical Computation for Artificial Intelligence} (pp.\ 103--119). San Diego: Academic Press, 1992.

\bibitem[Laurence, Hwang and Barone(2008)]{Laurence 2008} Laurence, P., Hwang, T-H.\ and Barone, L.\ (2008). Geometric properties of multivariate correlation in de Finetti's approach to insurance theory. \emph{Journal \'Electronique d'Histoire des Probabilit\'es et de la Statistique} (www.jehps.net) 4 (2): 1--13.

%\bibitem[Lehner(2014)]{Lehner 2014} Lehner, C.\ (2014). Einstein's realism and his critique of quantum mechanics. In M.\ Janssen and C.\ Lehner, \emph{The Cambridge Companion to Einstein} (pp.\ 306--353). Cambridge: Cambridge University Press.

\bibitem[Masanes, Galley and M\"uller(2019)]{Masanes et al 2019} Masanes, L., Galley, T.~D. and M\"uller, M.~P.\ (2019). The measurement postulates of quantum mechanics are operationally redundant. \emph{Nature Communications} 10: 1361.

\bibitem[McGrayne(2011)]{McGrayne 2011}  McGrayne, S.\ B.\ (2011). \emph{The Theory That Would Not Die: How Bayes' Rule Cracked the Enigma Code, Hunted Down Russian Submarines, \& Emerged Triumphant from Two centuries of Controversy}. New Haven: Yale University Press. 

\bibitem[Mensing and Pauli(1926)]{Mensing and Pauli 1926} Mensing, L.\ and Pauli, W.\ (1926). \"Uber die Dielektrizit\"atskonstante von Dipolgasen nach der Quantenmechanik. \emph{Physikalische Zeitschrift} 27: 509--512.

\bibitem[Mermin(1981)]{Mermin 1981} Mermin, N.\ D.\ (1981). Quantum mysteries for everyone. \emph{Journal of Philosophy} 78: 397--408. Page references to reprint in \citet[pp.\ 81--94]{Mermin 1990}.

\bibitem[Mermin(1988)]{Mermin 1988} Mermin, N.\ D.\ (1988). Spooky actions at a distance: mysteries of the quantum theory.  Page references to reprint in \citet[pp.\ 110--176]{Mermin 1990}.

\bibitem[Mermin(1990)]{Mermin 1990} Mermin, N.\ D.\ (1990). \emph{Boojums All the Way Through. Communicating Science in a Prosaic Age.} Cambridge: Cambridge University Press.

\bibitem[Messiah(1962)]{Messiah 1962} Messiah, A.\ (1962). \emph{Quantum Mechanics.} 2 Vols. New York: Wiley. 

\bibitem[Midwinter and Janssen(2013)]{Midwinter and Janssen 2013} Midwinter, C.\ and Janssen, M.\ (2013). Kuhn losses regained: Van Vleck from spectra to susceptibilities. In M.\ Badino and J.\ Navarro, Eds., \emph{Research and Pedagogy: A History of Early Quantum Physics Through Its Textbooks} (pp.\ 137--205). Berlin: Edition Open Access.

\bibitem[Myrvold(2010)]{Myrvold 2010} Myrvold, W.\ (2010). From physics to information theory and back. In A.\ Bokulich and G.\ Jaeger, Eds., \emph{Philosophy of Quantum Information and Entanglement} (pp.\ 181--207). Cambridge: Cambridge University Press.

\bibitem[Nielsen and Chuang(2016)]{Nielsen and Chuang 2016} Nielsen, M.\ A.\ and Chuang, I.\ L.\ (2016). \emph{Quantum Computation and Quantum Information.} 10th ann.\ ed.\ Cambridge: Cambridge University Press.

\bibitem[Pauli(1921)]{Pauli 1921} Pauli, W.\ (1921). Zur Theorie der Dieelektrizit\"atskonstante zweiatomiger Dipolgase. \emph{Zeitschrift f\"ur Physik} 6: 319--327.

\bibitem[Pauli(1994)]{Pauli 1994} Pauli, W.\ (1994). \emph{Writings on Physics and Philosophy.} C.\ P.\ Enz and K.\ von Meyenn, Eds. Berlin: Springer.

\bibitem[Pawlowski \emph{et al}(2009)]{Pawlowski et al 2009} Paw{\l}owski, M., Paterek, T., Kaszlikowski, D., Scarani, V., Winter, A.\ and {\.Z}ukowski, M.\  (2009). Information causality as a physical principle. \emph{Nature} 461: 1101--1104.

\bibitem[Pearson(1896)]{Pearson 1896} Pearson, K.\ (1896). Mathematical contributions to the theory of evolution. III. Regression, heredity and panmixia. \emph{Philosophical Transactions of the Royal Society, London, Series A} 186: 343--414.

\bibitem[Pearson(1916)]{Pearson 1916} Pearson, K.\ (1916). On some novel properties of partial and multiple Correlation in a universe of manifold characteristics. \emph{Biometrika} 11: 231--238.

\bibitem[Pearson(1921)]{Pearson 1921} Pearson, K.\ (1921). Notes on the history of correlation. \emph{Biometrika} 13: 25--45. Page reference to reprint in: E.\ S.\ Pearson and M.\ G.\ Kendall, Eds., \emph{Studies in the History of Statistics and Probability} (pp.\ 185--205). London: Griffin.  

\bibitem[Pitowsky(1986)]{Pitowsky 1986} Pitowsky, I.\ (1986). The range of quantum probabilities. \emph{Journal of Mathematical Physics} 27: 1556--1566.

\bibitem[Pitowsky(1989a)]{Pitowsky 1989a} Pitowsky, I.\ (1989a). From George Boole to John Bell: The Origins of Bell's inequalities. In \citet[pp.\ 37--49]{Kafatos 1989}.

\bibitem[Pitowsky(1989b)]{Pitowsky 1989b} Pitowsky, I.\ (1989b).   (Lecture Notes in Physics, No.\ 321). Heidelberg: Springer.

\bibitem[Pitowsky(1991)]{Pitowsky 1991} Pitowsky, I.\ (1991). Correlation polytopes, their geometry and complexity. \emph{Mathematical Programming} A 50: 395--414.

\bibitem[Pitowsky(1994)]{Pitowsky 1994} Pitowsky, I.\ (1994). George Boole's `conditions of possible experience' and the quantum puzzle. \emph{The British Journal for the Philosophy of Science} 45 95--125.

\bibitem[Pitowsky(2006)]{Pitowsky 2006} Pitowsky, I.\ (2006). Quantum mechanics as a theory of probability.  In W.\ Demopoulos and I.\ Pitowsky, Eds., \emph{Physical Theory and its Interpretation. Essays in Honor of Jeffrey Bub} (pp. 213--240). New York: Springer.

\bibitem[Pitowsky(2008)]{Pitowsky 2008} Pitowsky, I.\ (2008). Geometry of quantum correlations. \emph{Physical Review} A 77: 062109. 
%xxx--yyy see bananaworld, p. 120, note 1.

\bibitem[Pokorny \emph{et al.}(2019)]{Pokorny et al 2019} Pokorny, F., Zhang, C., Higgins, G., Cabello, C., Kleinmann, M.\ and Hennrich, M.\  (2019). Tracking the dynamics of an ideal quantum measurement. \emph{arXiv}: 1903.10398v1.

\bibitem[Popescu(2016)]{Popescu 2016} Popescu, S.\ (2016). Foreword. In \citet[pp.\ v--vi]{Bub 2016}

\bibitem[Popescu and Rohrlich(1994)]{Popescu and Rohrlich 1994} Popescu, S.\ and Rohrlich, D.\ (1994). Quantum nonlocality as an axiom. \emph{Foundations of Physics} 24: 370--385.

\bibitem[Pusey, Barrett and Rudolph(2012)]{PBR} Pusey, M.\ F., Barrett, J.\ and Rudolph, T.\ (2012). On the reality of the quantum state. \emph{Nature Physics} 8: 475--478

\bibitem[Quine(1951)]{Quine 1951} Quine, W.\ V.\ O.\ (1951). Two dogmas of empiricism. \emph{The Philosophical Review} 60: 20--43.

\bibitem[Rastall(1995)]{Rastall 1995} Rastall, P.\ (1995). Locality, Bell's theorem, and quantum mechanics. \emph{ Foundations of Physics} 15: 963--972.

%\bibitem[Renn(2007)]{Renn 2007} Renn, J., Ed.\ (2007). \emph{The Genesis of General Relativity.} 4 Vols. New York, Berlin: Springer.

\bibitem[Sainz \emph{et al.}(2018)]{Sainz et al 2018} Sainz, A.\ B., Guryanova, Y\, Ac\'in, A., and Navascu\'es, M.\  (2018). Almost-quantum correlations violate the no-restriction hypothesis. \emph{Physical Review Letters} 120: 200402.

\bibitem[Saunders \emph{et al.}(2010)]{Many Worlds 2010} Saunders, S., Barrett, J., Kent, A.\ and D.\ Wallace, D.,Eds.\ (2010). \emph{Many Worlds? Everett, Quantum Theory, and Reality}. Oxford: Oxford University Press.

%\bibitem[Schlosshauer (2007)]{Schlosshauer 2007} Schlosshauer, M.\ (2007). \emph{Decoherence and the Quantum-to-Classical Transition.} Berlin: Springer.

\bibitem[Schr\"odinger(1926)]{Schroedinger 1926} Schr\"odinger, E.\ (1926).  Quantisierung als Eigenwertproblem.  Dritte Mitteilung: St\"orungstheorie, mit Anwendung auf den Starkeffekt der Balmerlinien. \emph{Annalen der Physik}  80: 437--490. 

%\bibitem[Schr\"odinger(1935)]{Schroedinger 1935} Schr\"odinger, E.\ (1935). Discussion of probability relations between separated systems. \emph{Mathematical Proceedings of the Cambridge Philosophical Society} 31: 555--563.

\bibitem[Schwarzschild(1916)]{Schwarzschild 1916} Schwarzschild, K.\ (1916). Zur Quantenhypothese. \emph{K\"oniglich Preussischen Akademie der Wissenschaften} (Berlin)  \emph{Sitzungsberichte}: 548--568.

\bibitem[Sommerfeld(1919)]{Sommerfeld 1919} Sommerfeld, A.\ (1919). \emph{Atombau und Spektrallinien.} Braunschweig: Vieweg.

\bibitem[Spekkens(2007)]{Spekkens 2007} Spekkens, R.\ W.\, Evidence for the epistemic view of quantum states: A toy theory. \emph{Physical Review A} 75: 032110.

%\bibitem[Stachel(1994)]{Stachel 1994} Stachel, J.\ (1994). Changes in the concepts of space and time brought about by relativity. In C.\ C.\ Gould and R.\ S.\ Cohen, Eds., \emph{Artifacts, Representation and Social Practice} (pp.\ 141--162). Dordrecht: Kluwer.

\bibitem[Stark(1913)]{Stark 1913} Stark, J.\ (1913). Beobachtungen \"uber den Effekt des elektrischen Feldes auf Spektrallinien. {\it K\"oniglich Preussischen Akademie der Wissenschaften} (Berlin). {\it Sitzungsberichte}: 932-946.

\bibitem[Stone(1932)]{Stone 1932} Stone, M.\ H.\  ( 1932). \emph{Linear Transformations in Hilbert Space}. New York: American Mathematical Society.

\bibitem[Tanabashi \emph{et al.}(2018)]{Tanabashi et al 2018} Tanabashi, M.\ \emph{et al}.\ (2018). Review of Particle Physics (Particle Data Group). \emph{Physical Review} D 98, 030001.

\bibitem[Teukolsky(1996)]{Teukolsky 1996} Teukolsky, S.\ A.\ (1996). The explanation of the Trouton-Noble experiment revisited. \emph{American Journal of Physics}   64: 1104--1106. 

\bibitem[Timpson(2010)]{Timpson 2010} Timpson, C.\ (2010). Rabid dogma? Comments on Bub and Pitowsky [2010]. In \citet[pp.\ 460--466]{Many Worlds 2010}.

\bibitem[van der Waerden(1968)]{Van der Waerden} Van der Waerden, B.\ L., Ed.\ (1968). \emph{Sources of quantum mechanics.} New York: Dover.

\bibitem[van Fraassen(1991)]{Van Fraassen 1991} Van Fraassen, B.\ C.\ (1991). \emph{Quantum Mechanics. An Empiricist View.} Oxford: Clarendon.

\bibitem[Van Vleck(1926)]{Van Vleck 1926} Van Vleck, J.\ H.\ (1926). Magnetic susceptibilities and dielectric constants in the new quantum mechanics. \emph{Nature} 118: 226--227.

\bibitem[Van Vleck(1928)]{Van Vleck 1928} Van Vleck, J.\ H.\ (1928). The new quantum mechanics. \emph{Chemical Reviews} 5: 467--507.

\bibitem[Van Vleck(1932)]{Van Vleck 1932} Van Vleck, J.\ H.\ (1932). \emph{The Theory of Electric and Magnetic Susceptibilities}. Oxford University Press, Oxford.

\bibitem[von Mises(1928)]{von Mises 1928} Von Mises, R.\ (1928). \emph{Wahrscheinlichkeit, Statistik und Wahrheit}. Vienna: Springer.  

\bibitem[von Neumann(1927a)]{von Neumann 1927a} Von Neumann, J.\ (1927a). Mathematische Begr\"undung der Quantenmechanik.  \emph{K\"onigliche Gesellschaft der Wissenschaften zu G\"ottingen. Mathematisch-physikalische Klasse. Nachrichten} 1--57.

\bibitem[von Neumann(1927b)]{von Neumann 1927b} Von Neumann, J.\ (1927b). Wahrscheinlichkeitstheoretischer Aufbau der Quantenmechanik.  \emph{K\"onigliche Gesellschaft der Wissenschaften zu G\"ottingen. Mathematisch-physikalische Klasse. Nachrichten} 245--272.

\bibitem[von Neumann(1927c)]{von Neumann 1927c} Von Neumann, J.\ (1927c). Thermodynamik quantenmechanischer Gesamtheiten. \emph{K\"onigliche Gesellschaft der Wissenschaften zu G\"ottingen. Mathematisch-physikalische Klasse. Nachrichten} 273--291.

\bibitem[von Neumann(1932)]{von Neumann 1932} Von Neumann, J.\ (1932). \emph{Mathematische Grundlagen der Quantenmechanik.} Berlin: Springer.

%\bibitem[Wallace(2016)]{Wallace 2016} Wallace, D.\ (2016). What is orthodox quantum mechanics? \emph{arXiv}: 1604.05973v1.

\bibitem[Wallace(2019)]{Wallace 2019} Wallace, D.\ (2019). On the plurality of quantum theories: Quantum theory as a framework, and its implications for the quantum measurement problem. Preprint, \emph{PhilSci Archive}. To be published in: S. French and J. Saatsi, Eds., \emph{Realism and the Quantum}. Oxford, Oxford University Press, forthcoming.

\bibitem[Wigner(1931)]{Wigner 1931} Wigner, E.\ (1931). \emph{Gruppentheorie und ihre Anwendung auf die Quantenmechanik der Atomspektren}. Braunschweig: Vieweg. 
%Reprinted in facsimile by Authority of the Alien Property Custodian  in 1944.
%under License No.\ A-281.

\bibitem[Yule(1897)]{Yule 1896} Yule, G.\ U.\ (1897). On the significance of Bravais' formulae for regression, \&c., in the case of skew correlation. \emph{Proceedings of the Royal Society of London} 60: 477--489. 
%(issues: 359�367). Udny Yule.
%In 1846 he published a paper on the statistical concept of correlation, and arrived at a definition of the correlation coefficient before Karl Pearson.
%Stephen Mack Stigler. 

  

\end{thebibliography}

\end{document}
