\documentclass[10pt, nofootinbib]{revtex4}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                         %%   Version revised on 23/10/08            %%                                         %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%\newcommand{\be}{\begin{equation}}\newcommand{\ee}{\end{equation}}\newcommand{\bea}{\begin{eqnarray}}\newcommand{\eea}{\end{eqnarray}}\newcommand{\bean}{\begin{eqnarray*}}\newcommand{\eean}{\end{eqnarray*}}\newcommand{\Ref}[1]{(\ref{#1})}\newcommand{\p}{\partial}\begin{document}\title{\Large\bf What is a particle?}

\author{Daniele Colosi${}^a$, Carlo Rovelli${}^b$}

\affiliation{
\small\it ${}^a$Instituto de Matem\'aticas, UNAM,
Campus Morelia, C.P.\,58190, Morelia, Michoac\'an, Mexico\\
\small\it ${}^b$Centre de Physique Th\'eorique de Luminy\footnote{Unit\'e mixte de recherche (UMR 6207) du CNRS et des Universit\'es de Provence (Aix-Marseille I), de la M\'editerran\'ee (Aix-Marseille II) et du Sud (Toulon-Var); laboratoire affili\'e \`a la FRUMAM (FR 2291).}, Case 907, F-13288 Marseille, EU \vspace{0.2cm}
} 



\bigskip\date{\small\today} \begin{abstract} \noindent \noindent Theoretical developments related to the gravitational interaction 
have questioned thenotion of \emph{particle} in quantum field theory (QFT).  Forinstance, uniquely-defined particle states do not exist in general, 
in QFT on a curved spacetime. More in general, particlestates are difficult to define in a background-independent quantum 
theory of gravity.  These difficultieshave lead some to suggest that in general QFT should not beinterpreted in terms of particle states, but rather in terms ofeigenstates of {local} operators.  Still, it is notobvious how to reconcile this view with the empirically-observedubiquitous particle-like behavior of quantum fields, apparent
for instance in experimental high-energy physics, or ``particle"-physics.  Here we offeran element of clarification by observing that already in flat spacethere exist --strictly speaking-- {two} distinct notions of particles:globally defined $n$-particle Fock-states and \emph{local particlestates}.  The last describe the physical objects detected byfinite-size particle detectors and  {are}  eigenstates of localfield operators.  In the limit in which the particle detectors areappropriately large, global and localparticle states converge in a weak topology (but not in norm).  Thisobservation has little relevance for flat-space 
theories --it amounts to a reminder that there are boundaryeffects in realistic detectors--; but is relevant for gravity. It reconciles the two points ofview mentioned above. More importantly, it provides a definition of{\em local} particle state that remains well-defined even when the conventionalglobal particle states are not defined.  This definition plays animportant role in quantum gravity.\end{abstract}
\maketitle\section{Introduction}Is a particle a local or a global object?  On the one hand, a particleis a local object detected by a local apparatus, such as aphotoelectric detector or a high-energy-experiment bubble chamber.  Onthe other hand, the $n$-particle states of quantum field theory (QFT),namely the eigenstates of the particle-number operator in Fock space,
have a well-known nonlocal character; for instance, they are noteigenstates of local operators.  There is a tension between these twofacts.This tension becomes acute when QFT is defined on acurved spacetime \cite{bd,wald}.  In this contest, the definition ofparticle states depends on the choice of a spacelike foliation ``allover the universe", a choice that has no physical meaning.  In flat 
space, Poincar\'e invariance selects preferred foliations and particle states are defined by decomposingthe field into modes and distinguishing positive and negativefrequencies.  On curved spacetime, in general there is no 
symmetry group, no preferred set of 
modes and no preferred decomposition intopositive and negative frequency.  As a consequence, there is nopreferred vacuum state, and the interpretation of the field states interm of particles appears to be difficult.  

In fact, it is well known that the Poincar\'{e} group plays a centralrole in the particle interpretation of the states of the field:Wigner's celebrated analysis \cite{wigner} has shown that the Fock
particle states are the irreducible representations of the Poincar\'{e} groupin the QFT state space.  The defining properties of the particles,mass and spin (or helicity), are indeed the invariants of thePoincar\'e group.  Now, strictly speaking we do not live in aPoincar\'e invariant region of spacetime: does this means that,strictly speaking, the world around us has no particles?
Such arbitrariness and ambiguity of the particle concept have led sometheoreticians like Davies to affirms that ``particles do not exist"\cite{Davies}, a view shared by several relativists.  The idea is thatQFT should be interpreted in terms of local quantities, such as theintegral of energy-momentum-tensor components over finite regions, asmaintained for instance by Wald \cite{wald}.  But this view is notshared by many other theoreticians, especially (not surprisingly!)  comingfrom the ``particle"-physics tradition, who hold that QFT isfundamentally a formalism for describing processes involvingparticles, such as scattering or decays.  A typical example of thisposition is Weinberg \cite{weinberg}, who cannot certainly be suspectedof ignoring general relativity.

These difficulties become serious in a background-independent quantum 
context (see for instance \cite{book}). For instance, in loop quantum gravity
\cite{book,lqg}
quantum states of the gravitational field are described in terms of a 
spin network basis. Can we talk about gravitons, or other particle states, 
in loop quantum gravity \cite{gravitons}?  A common view among relativists is that we 
cannot, unless we consider the asymptotically flat context.  But there 
should well be a way of describing what a finite-size 
detector detects, even in a local background-independent theory! 
Indeed, a recent line of development in loop quantum gravity aims at 
computing transition amplitudes between particle states \cite{transitions},
using only {\em finite} spacetime regions, using a formalism 
developed in  \cite{boundary} and in \cite{book}. 
What are those particle states? What is a particle, in a context in which
there is no Poincar\'e invariance and no preferred foliation of a background 
spacetime?
Here we present an observation which may contribute to bring someclarity, and reconciliate the two points of view.  We address tworelated questions: (i) The problem of the local/global nature ofparticles.  More precisely: how can an apparatus localized inspacetime detect a particle state, if a particle state is not aneigenstate of a local field operator?  And (ii) how can we understanda local apparatus detecting physical particles in the context ofcurved-spacetime QFT, or even in the context of background independentquantum gravity, where the standard global construction of particlestates is ambiguous?To address these questions, we observe that if the mathematicaldefinition of a particle appears somewhat problematic, its operationaldefinition is clear: particles are the objects revealed by detectors,tracks in bubble chambers, or discharges of a photomultiplier.  Now,strictly speaking a particle detector is a measurement apparatus thatcannot detect a $n$-particle Fock state, precisely because it islocalized.  A particle detector measures a local observable fieldquantity (for instance the energy of the field, or of a fieldcomponent, in some region).  This observables quantity is representedby an operator that in general has discrete spectrum.  \emph{Theparticles observed by the detector are the quanta of this localoperator.} Our key observation is that the eigenstates of thisoperator are states of the quantum field that are \emph{similar, butnot identical, to the Fock particle states defined globally}.Therefore, strictly speaking there are two distinct notions ofparticle in QFT. \emph{Local particle states} correspond to the realobjects observed by finite size detectors.  They are eigenstates oflocal operators. On the other hand,  \emph{global particle states}, 
such as the Fock particles, namely the eigenstates of the number 
operator in Fock space, can be defined only under certain conditions.  
Global particle states are simpler to define and they approximate well 
the local particle states detected by local measurements.  Therefore 
the global particle states, {\em when they are available}, give a goodapproximate description of the physics of the ``real" particles detectedby the detectors. In the paper we illustrate the difference between these two classes ofstates, and discuss their relation.  The precise sense in which globalstates approximate local particle states is subtle.  We show belowthat (contrary to what we expected at first) the convergence is not inthe Hilbert space norm, but only in a weaker topology defined by localobservables themselves.We only deal here with free fields.  This is not necessarily a trivialcontext even in flat space, as illustrated for instance by the Unruheffect \cite{Unruh:db}.  This effect can be understood in terms of abasis of particle states different from the standard Minkowski one,indicating that even in flat space there can be ambiguities in thedefinition of the notion of particle.  We expect our conclusions tohave general validity also for an interacting theory, but we do not 
venture here into generalizations. 
In flat space, and for inertial observers, the distinction betweenglobal and local particle states is needlessly punctilious, sincephysically it boils down to exponentially small correlation effects atthe detector's boundary.  But the distinction is conceptuallyimportant because it indicates that {\em particle states thatdescribe the physical particles we observe are equally well 
defined in flat space as is curved spacetime, and even in the 
absence of spacetime in a full quantum gravity context}.  The distinction shows that the global features of the Fock particle 
states have nothing to do with the real observed particles: they 
are an artifact of the simplification taken by approximating a 
truly observed local particle state with easier-to-deal-with 
Fock particles.
This conclusion opens the pathfor discussing particle states in a background-independent context.  
Therefore present paper provides the conceptual and technical 
justification for the use of particle states in the research line in the
references \cite{transitions} and, more in general in the boundary 
formulation of quantum field theory \cite{boundary,book}.  The notion of 
particle used in this context is  not 
inconsistent with the standard QFT notion of particle, in spite of 
the local character of the first and the global character of the second.   In the paper, we first introduce the distinction between localparticles and global particles using a very simple model: two coupledharmonic oscillators (Section 2).  Then we extend theconstruction to field theory in two steps.  First we consider asequence of a large number of coupled oscillators (Section 3,4), then wediscuss field theory in Section 5.  In all these cases, we defineglobal and local particle states and we discuss their relations.  InSection 6 we summarize our results and we give a general discussion ofthe notion of particle in QFT.\section{Two oscillators}To begin with, consider two weakly coupled harmonic oscillators $q_{1},q_{2}$, with unit mass and with the same angular frequency $\omega$;the dynamics is governed by the hamiltonian%\bea H_{0}&=& H_{1}+H_{2} + V = \frac{1}{2}\left(p_1^2 + \omega^2\; {q}_1^2\right) + \frac{1}{2}\left(p_2^2 + \omega^2\; {q}_2^2 \right)+\lambda\, q_{1}q_{2}, \eea%where $p_{1}, p_{2}$ are the momenta conjugate to $q_{1}, q_{2}$ and,say, $\lambda\ll\omega^2$.  The state space of the system is ${\calH}=L_{2}[R^2,dq_{1},dq_{2}]$ formed by the functions$\psi(q_{1},q_{2})$.  We can define an orthonormal basis in thisHilbert space by diagonalizing a complete set of commutingself-adjoint operators.  Let us choose the set formed by $H_{1}$ and$H_{2}$.  Call $E_1$ and $E_2$ the eigenvalues of the operators$H_{1}$ and $H_{2}$ respectively, and $|n_{1},n_{2}\rangle_{\rm loc}$their common eigenstates.  The reason for the suffix ``loc" will beclear in a moment.  The integers $n_{1}$ and $n_{2}$ are the quantumnumbers of $E_1$ and $E_2$ and we can interpret them as the number ofquanta in the first and in the second oscillator respectively.  Moreprecisely, if we measure the energy $H_{1}$ of the first oscillator weobserve that the measurement outcome is quantized:$E_{1}=\hbar\omega(n_{1}+1/2)$ and $n_{1}$ can be interpreted as thenumber of quanta in $q_{1}$.  It is suggestive to call these quanta``particles".  Call $N_{12}=n_{1}+n_{2}$ the total particle number,and call $n$-particle states the eigenstates of $H_{1}+H_{2}$. Introducing a Fock-like notation, we can write the state withno-particles also as\begin{equation} |0\rangle_{\rm loc}= |0,0\rangle_{\rm loc};\label{120}\end{equation}the two one-particle states with particles localized on each oscillator as\begin{eqnarray} |1\rangle_{\rm loc}= |1,0\rangle_{\rm loc}, \label{1loc}\\|2\rangle_{\rm loc}= |0,1\rangle_{\rm loc}, \end{eqnarray}where the state $|1\rangle_{\rm loc}$ represents a particle on thefirst oscillator and the state $|2\rangle_{\rm loc}$ represent aparticle on the second oscillator; and so on.  Notice that, accordingto standard Fock-space terminology, any linear combination ofone-particle states\begin{equation} |\psi\rangle_{\rm loc}= c_{1}|1\rangle_{\rm loc}+c_{2}|2\rangle_{\rm loc} \label{psiloc}\end{equation}is also called a one-particle state. Of course the states $|n_{1},n_{2}\rangle_{\rm loc}$ are notstationary states.  In a perturbation theory in $\lambda$, forinstance, we can compute the probability amplitude for theparticles to ``jump from one oscillator to the other", and so on.  Ifwe are interested in the stationary states, we need the normal modes ofthe system.  These are%\bea q_a =\frac{q_{1}+q_{2}}{\sqrt{2}}, \hspace{3em}q_b = \frac{q_{1}-q_{2} }{ \sqrt{2}}, \label{modes}\eea%with eigenfrequencies%\bea\omega_{{a}}^2 = \omega^2 + \lambda, \hspace{3em} \omega_{{b}}^2 = \omega^2 - \lambda.  \eea%In terms of these, the hamiltonian factorizes as%\beaH = H_{a} + H_{b} =\frac{1}{2}\left(p_a^2 +\omega_{a}q_a^2\right)  +\frac{1}{2}\left(p_b^2 +\omega_{b}q_b^2\right)  .\eea%Let $E_{a}$ ($E_{b}$) be the eigenvalues of $H_{a}$ $(H_{b})$, anddenote $|n_{a},n_{b}\rangle_{\rm }$ the common eigenstates of$H_{a}$ and $H_{b}$.  The number $n_{a}$ ($n_{b}$) is the number ofquanta (or ``particles") in the mode $a$ ($b$).  Call$N_{ab}=n_{a}+n_{b}$ the total number of these particles in thesystem.  For instance the no-particle state is%\begin{equation} |0\rangle_{\rm }=  |0,0\rangle_{\rm };\label{ab0}\end{equation}%the two one-particle states with particles localized on each\emph{mode} are\begin{eqnarray} | a\rangle &=& |1,0\rangle, \label{a}\\| b\rangle &=& |0,1\rangle. \end{eqnarray}A generic one-particle state is a state of the form%\begin{equation} |\psi\rangle = c_{a} |a\rangle + c_{b} |b\rangle.\label{psiglob}\end{equation}%What is the relation between the one-particle states$|\psi\rangle_{\rm loc}$ defined in (\ref{psiloc}) and the one particlestates $|\psi\rangle$ defined in (\ref{psiglob})?One may be naively tempted to say that they are the same states,namely that the two one-particle states $|1\rangle_{\rm loc}$ and$|2\rangle_{\rm loc}$ (single excitations of the oscillators) are justlinear combinations of the two one-particle states $|a\rangle$ and$|b\rangle$ (single excitations of the modes).  But this is not thecase.  In the classical theory, $q_{1}$ can be expressedas the linear combination of the two modes by inverting (\ref{modes}):%\bea q_1 =\frac{q_{a}+q_{b}}{\sqrt{2}};\eea%accordingly, we can choose $c_{a}=c_{b}=1/\sqrt{2}$ in(\ref{psiglob}), and we obtain a one-particle state which is maximallyconcentrated on the first oscillator. Denote it %\begin{equation} |1\rangle= \frac{1}{\sqrt{2}} |a\rangle +\frac{1}{\sqrt{2}} |b\rangle.\label{1glo}\end{equation}%Is this state equal to $|1\rangle_{\rm loc}$?  No, it is not.  If$\lambda$ is small the two states differ only a little, but they dodiffer.  Both states are, in some sense, ``one particle states" and inboth states the ``particle" is on the first oscillator.  However, theyare distinct states.  We illustrate their difference in two ways.  First, we can simplywrite both of them explicitly in the coordinate basis.  It is a simpleexercise to show that%\bea \langle q_{1},q_{2}|1\rangle_{\rm loc} &=& {\textstyle{\sqrt{\frac{2\omega^3}{\pi}}}}\  q_{1}\ e^{ -\frac{\omega}{2}(q_1^2+q_2^2) } \label{1loc11}\eea%while%\bea\langle q_{1},q_{2}|1\rangle_{\rm } &=& \textstyle{\frac{(\omega_a \omega_b)^{1/4}}{\sqrt{2\pi}} }\ \left( \frac{\sqrt{\omega_a}+\sqrt{\omega_b}}{2}q_{1}+ \frac{\sqrt{\omega_a}-\sqrt{\omega_b}}{2}q_{2}\right)\ e^{-\frac{1}{2}\left(\frac{\omega_{a}+\omega_{b}}{2}(q_{1}^2+q_{2}^2)+\frac{\omega_{a}-\omega_{b}}{2} q_{1}q_{2}\right)}.\label{1glo11}\eea%If $\lambda$ is small, $\omega_{a}\sim\omega_{b}\sim\omega$ and thetwo states are similar.  In fact, it is easy to compute that theirscalar product is%\bea{}_{\rm loc}\langle 1|1\rangle &=& 1 - O(\lambda^2). \label{ss}\eea%which means that, in a sense, the two states are indistinguishableeven at first order in $\lambda$.  Second, we can compare them usingperturbation theory in $\lambda$.  This is instructive because we willbe able to do the same in the context of field theory.  Let us take$H_{0}=H_{1}+H_{2}$ as unperturbed hamiltonian.  The two states$|1\rangle_{\rm loc}$ and $|2\rangle_{\rm loc}$ span a degenerate eigenspace of $H_{0}$. We must therefore diagonalize $V$ on this eigenspace to start perturbation theory. Clearly $V$ is diagonalized in this subspace by the two states \begin{eqnarray} |a\rangle_{0} &=& \frac{|1\rangle_{\rm loc}+|2\rangle_{\rm loc}}{\sqrt 2} , \\|b\rangle_{0} &=& \frac{|1\rangle_{\rm loc}-|2\rangle_{\rm loc}}{\sqrt 2}.  \end{eqnarray}We can compute the first order correction to these states using first order perturbation theory. It is convenient to use creation and annihilation operators \begin{eqnarray} q_{1,2} &=& \frac{1}{\sqrt{2\omega}}(a_{1,2}+ a^\dagger_{1,2}),\\p_{1,2} &=& \frac{-i\sqrt{\omega}}{\sqrt{2}}(a_{1,2}-a^\dagger_{1,2})\end{eqnarray}in terms of which the perturbation reads\begin{equation}V = \frac{\lambda}{2\omega} (a^\dagger_{1}a^\dagger_{2}+a_{1}a_{2}+a^\dagger_{1}a_{2}+a_{1}a^\dagger_{2}). \end{equation}Notice that the term $a^\dagger_{1}a^\dagger_{2}$ brings out from theone particle sector, giving the non-vanishing matrix elements\begin{eqnarray}\langle 2,1 |V| a \rangle = \hskip4pt\langle 1,2 |V| a \rangle &=& \frac{\lambda}{2 \omega}\\\langle 2,1 |V| b \rangle = - \langle 1,2 |V| b \rangle &=& \frac{\lambda}{2 \omega}\end{eqnarray}To first order in $\lambda$, the hamiltonian eigenstates $| a\rangle$ and $| b \rangle$ are therefore\begin{equation} |a\rangle\ =\ |a\rangle_{0}\ +\ \frac{\langle 2,1 |V| a\rangle}{E_{a}-E_{2,1}}\ |2,1\rangle \ +\ \frac{\langle 1,2 |V| a\rangle}{E_{a}-E_{2,1}}\ |1,2\rangle\ =\  |a\rangle_{0}\ -\frac{\lambda}{4\omega^2}\| 2, 1 \rangle \ -\ \frac{\lambda}{4\omega^2}\ | 1, 2 \rangle\end{equation}and\begin{equation} |b\rangle\ =\ |b\rangle_{0}\ +\ \frac{\langle 2,1 |V| b\rangle}{E_{a}-E_{2,1}}\ |2,1\rangle\ -\ \frac{\langle 1,2 |V| b\rangle}{E_{a}-E_{2,1}}\ |1,2\rangle\ =\ |b\rangle_{0}\ -\frac{\lambda}{4\omega^2}\| 2, 1 \rangle \ +\ \frac{\lambda}{4\omega^2}\ | 1, 2 \rangle\end{equation}And therefore, to first order in $\lambda$%\be |1\rangle\ =\ |1\rangle_{\rm loc}\ -\\frac{\lambda}{\sqrt{8}\, \omega^2}\ |2,1\rangle.\ee%Thus, the two states (\ref{1loc}) and (\ref{1glo}) are both ``one-particlestates" in which the ``particle" is concentrated on the oscillator$q_{1}$, but they are distinct states.  They represent two distinctkinds of one-quantum states, or two distinct kinds of quanta.  We call$|1\rangle_{\rm loc}$ a \emph{local} particle state, and$|1\rangle_{\rm }$ a \emph{global} particle state.  They represent thesimplest example of the distinction between these two classes ofstates.More in general, we call ``global particle states" theeigenstates of the ``global" number operator%\bea N_{ab}\ |n_{a},n_{b}\rangle_{\rm } &=& (n_{a}+n_{b})\|n_{a},n_{b}\rangle\ \ , \eea%and we call ``local particle states" the eigenstates of the ``local"number operator%\bea N_{1}\ |n_{1},n_{2}\rangle_{\rm loc} &=& n_{1}\|n_{1},n_{2}\rangle_{\rm loc} \ \ .\eea%Let us illustrate the different properties that these states have. The state $|1\rangle_{\rm loc}$ is an eigenstate of $H_{1}$, which isan observable that depends just on $q_{1}$ and its momentum, namelyjust on the variable associated to the first oscillator.  If we wantto measure how many local particles are in the first oscillator,namely to measure $n_{1}$, we can make a measurement that involvessolely variables of the $q_{1}$ oscillator.  In this sense$|1\rangle_{\rm loc}$ is ``local".The state $|1\rangle_{\rm }$, on the other hand, describes a singleparticle ``on the first oscillator", but is not an eigenstate ofobservables that depend on variables of the sole first oscillator. This can be seen from the fact that it is a state in which the twooscillators are (weakly) correlated. The source of thesecorrelations can be traced to the vacuum state: local particles areexcitation over the local vacuum (\ref{120}) which has nocorrelations:%\bea \langle q_{1},q_{2}|0\rangle_{\rm loc} &=& {\textstyle{\sqrt{\frac{\omega}{\pi}}}}\ e^{ -\frac{\omega}{2}q_1^2 } \ e^{ -\frac{\omega}{2}q_2^2 } = \psi_{0}(q_{1})\,  \psi_{0}(q_{2})\label{0loc}\eea%while global  particles are excitations over the global vacuum (\ref{ab0})%\bea\langle q_{1},q_{2}|0\rangle_{\rm } &=& \textstyle{\frac{(\omega_a \omega_b)^{1/4}}{\sqrt{\pi}} }\ e^{-\frac{1}{2}\frac{\omega_{a}+\omega_{b}}{2}q_{1}^2}\ e^{-\frac{1}{2}\frac{\omega_{a}+\omega_{b}}{2}q_{2}^2}\ e^{-\frac{1}{2}\frac{\omega_{a}-\omega_{b}}{2} q_{1}q_{2}},\label{0glo}\eea%which does not factorize, and therefore represents vacuum correlationsbetween the two oscillators.  In Appendix A we give a more precise andquantitative expression of this correlation.Notice that $|1\rangle_{\rm loc}$ is not an energy eigenstate, becauseof the interaction term $V$, but $|1\rangle_{\rm }$ isn't an energyeigenstate either, because $ |1,0\rangle_{\rm }$ and $|0,1\rangle_{\rm}$ have different energies.  Its defining property is just the fact ofbeing a linear combination of one-quantum excitations of the normalmodes of the system.  What is then the physical relevance of the state$|1\rangle_{\rm }$?  It is the following: the one-particle Fock statesof QFT are precisely states of the same kind as $|1\rangle_{\rm }$. To see this, consider a Fock particle localized in a region $R$.  Thisstate can be described by means of a function $f(x)$ with compactsupport in $R$, as%\bea|f\rangle_{\rm } &=& \int dk \ \tilde f(k)\ |k\rangle. \label{fock1}\eea%where $\tilde f(k)$ is the Fourier transform of $f(x)$ and the states$|k\rangle$ are the one-particle Fock states with momentum $k$.  Theyare energy eigenstates (with different energies) and they aresingle-particle excitations of the normal modes of the system. Therefore they are analog to the states $|1,0\rangle_{\rm }$ and$|0,1\rangle_{\rm }$ of the two-oscillators model.  The linearcombination (\ref{fock1}) is the analog of the linear combination(\ref{1glo}), which picks the one-particle global state maximallyconcentrated in the region chosen (the oscillator $q_{1}$ in themodel, the region $R$ in the QFT).  Thus, Fock particles are globalparticles.  No measurement in a finite region $R$ can count thoseparticles, because Fock particles are not eigenstates of local fieldoperators, precisely in the same sense in which $|1\rangle_{\rm }$ isnot an eigenstate of an observable localized on the $q_{1}$oscillator.  If we make a measurement with an apparatus located in theregion $R$, we can count the number of particles the apparatus detect. However, these particles are not global particles.  They are localparticles, that can be described by appropriate QFT states which areclose, but not identical, to $n$-particle Fock states, like$|1\rangle_{\rm loc}$ is close, but not identical to $|1\rangle_{\rm}$.  Later on, we discuss local particle states, analog to the$|n_{1},n_{2}\rangle_{\rm loc}$ states, in the context of QFT.Suppose now the state of the system is $|0\rangle_{\rm }$ and wemeasure whether a particle is on the first oscillator by measuring theenergy $E_{1}$.  The probability of not seeing any particle is notdetermined by the sole scalar product (\ref{ss}), because we are infact tracing over $n_{2}$.  Rather, it is given by%\be {\cal P} \ =\  \left| \sum_{n_{2}}\ {}_{\rm loc}\langle 0, n_{2}|0\rangle_{\rm } \right|^2 \ =\ {}_{\rm }\langle 0|P_{0_{\rm loc}}|0\rangle_{\rm }\ee%where %\bea P_{0_{\rm loc}}=\sum_{n_2}\ |0,n_{2}\rangle_{\rm loc}\ {}_{\rm loc}\langle0,n_{2}|\eea%is the projection on the lowest eigenspace of $H_{1}$.A straightforward calculation gives %\be{\cal P} \ =\ {}_{\rm }\langle 0|P_{0_{\rm loc}}|0\rangle_{\rm }\ =\  1 -\frac{1}{16}\frac{\lambda^2}{\omega^4} +O(\lambda^4). \label{0p}\ee%This expression gives a quantitative evaluation of the ``error" thatwe make in confusing local particles with global particles: if thesystem is in the global vacuum state, there is a probability $1-{\calP}$ that a particle detector localized on the first oscillator detects a particle. \section{Chain of oscillators}As an intermediate step before going to field theory, let us considera chain of coupled harmonic oscillators.  This system allows us toemphasize several important points regarding the relation betweenlocal and global particle states.We study a system of $n$ harmonic oscillators $\mathbf q=(q^i), i=1,\ldots, n$with the same frequencies $\omega=1$ and coupled by a constant$\lambda$.  Each oscillator is coupled with its two neighboring (exceptthe first and the last oscillator that have only one coupling)%\bea H&=&  \frac{1}{2}\big(\; |\mathbf p|^2 +|\mathbf q|^2\, \big) + \lambda \sum_{i=1}^{n-1}  q^{i}q^{i+1}\label{H}\eea%where $|\mathbf q|^2=\sum_i (q^i)^2$.  Notice that we are notconsidering a ring but an open chain of oscillators.  Diagonalizingthe hamiltonian of the system we obtain the normal frequencies%\bea\omega_a= \sqrt{1+2 \lambda \cos \theta_a}, \  \ \; \; \hbox{where} \  \ \theta_a = \frac{a \pi}{n+1},\; \; \  \ \hbox{and} \   \ a=1,...,n.\eea%The normal modes $\mathbf Q=(Q_a), a=1,\ldots, n\ $ are given by$\textbf{Q}= U^{\scriptscriptstyle{(n)}} \textbf{q}$, where $ U^{\scriptscriptstyle{(n)}}$ is the orthogonal $n \times n$ matrix%\beaU^{\scriptscriptstyle{(n)}}_{ai}= \sqrt{\frac{2}{n+1}} \sin \left( \frac{ai\pi}{n+1} \right).\label{matrix}\eea%The vacuum state is%\bea \langle \mathbf q | 0 \rangle_{\rm } = \prod^{n}_{a=1}\left(\frac{\omega_a}{\pi}\right)^{1/4} e^{-\frac{1}{2} q^i D^{\scriptscriptstyle{(n)}}_{ij} q^j}. \label{Pserie}\eea%where%\beaD^{\scriptscriptstyle{(n)}}_{ij}= \sum_{a}  U^{\scriptscriptstyle{(n)}}_{ai}\omega_a U^{\scriptscriptstyle{(n)}}_{aj}\label{D}\eea%A basis that diagonalizes $H$ is given by the states $|\mathbfn\rangle= |n_{1}, \ldots n_n\rangle$ with $n_{a}$quanta in the $a$-th normal mode.  The number operator is%\bea N \  | \mathbf n \rangle_{\rm } = \left(\sum_{a=1}^n\ n_{a}\right)\ \  |\mathbf n\rangle\label{numberop}. \eea%Denote $|a\rangle$ the one particle state $|0, \ldots, 1,\ldots,0\rangle$ in which all normal modes are in the vacuum state except forthe $a$-th mode which is in its first excitation.  The state%\bea |i\rangle  = \sum_{a=1}^n\  U^{-1}_{ia}\ |a\rangle\label{gparticle}\eea%is the one particle state maximally concentrated on the $i$-thoscillator.  It is the analog of the global one-particle states(\ref{1glo}) and (\ref{fock1}).  This is the global one-particlestate, with the particle on the $i$-th oscillator.Now, consider a partition of the chain in two regions $R_{1}$ and$R_{2}$.  Let the region $R_{1}$ be formed by the first $n_{1}$oscillators, and the region $R_{2}$ be formed by the remaining $n_{2}$oscillators, with $n_{1}+n_{2}=n$.  We write $\mathbf q=(\mathbfq_{1},\mathbf{q}_{2})$, where $\mathbf q_{1}$ (respectively $\mathbfq_{2}$ ) is a vector with $n_{1}$ ($n_{2}$) components.  We regard thefirst region of oscillators as a generalization of the oscillator$q_{1}$ in the previous section, and the second region as the analog ofthe oscillator $q_{2}$.  The total Hilbert space of the systemfactorizes as ${\cal H}={\cal H}_{1}\otimes{\cal H}_{2}$.  We canrewrite the hamiltonian (\ref{H}) in the form%\bea H&=& H_{1}+H_{2}+V \\ \nonumber &=& \left(\frac{1}{2}\left(|\mathbf p_1|^2 + |\mathbf q_1|^2\right) + \sum_{i=1}^{n_{1}-1} \lambda \ q_{1}^iq_1^{i+1}\right)+ \left(\frac{1}{2}\left(|\mathbf p_2|^2 + |\mathbf q_2|^2\right) + \sum_{i=1}^{n_{2}-1} \lambda \ q_{2}^iq_2^{i+1}\right)+ \lambda\ q_1^{n_{1}}q_2^1.\eea%We ask what is a particle, or a quantum excitation of the system,localized in the region $R_{1}$.  As before, there are two possibleanswers.First, we can consider the vacuum state $|0 \rangle_{\rm }$ and definethe global one-particle states as a linear combination of singlequantum excitations of the normal modes of the system.  In particular,the linear combination can be chosen to be concentrated in the firstregion.  If $i<n_{1}$ is in the first region, (\ref{gparticle}) is astate representing a global particle state in $R_{1}$.If we make a measurement in the region $R_{1}$, however --namely ifwe measure a quantity that depends only on the variables $\mathbfq_{1}$ (and their momenta)-- we do not measure a state like(\ref{gparticle}), because this state is a state where the two regionsare correlated.  More precisely, this state is not an eigenstate of anobservable localized in $R_{1}$.  Suppose thus that we have only access to observables that arefunctions of the oscillator variables $\mathbf q_{1}$ in the firstregion.  For concreteness, suppose we measure the energy $H_{1}$contained in the first region.  Consider eigenvalues and eigenstatesof $H_{1}$ alone.  These are easy to find, since the calculation isthe same as above, only with $n$ replaced by $n_{1}$.  In particular,we must diagonalize $H_{1}$ in ${\cal H}_{1}$.  For this, \emph{weneed the normal modes of\ $H_{1}$ alone}.  These normal modes aregiven by $\mathbf{Q_1}= U^{\scriptscriptstyle{(n_{1})}}\mathbf{q_{1}}$, where $ U^{\scriptscriptstyle{(n_{1})}}$ is theorthogonal $n_{1} \times n_{1}$ matrix (\ref{matrix}).  Let$|0\rangle_{1}$ be the lowest eigenstate of $H_{1}$ in ${\cal H}_{1}$and $|\mathbfn_{1}\rangle_{1}= |n_{1}, \ldots n_{n_1}\rangle_{1}$ with $n_{a}$quanta in the $a$-th normal mode of $H_{1}$ is in its $n_{a}$-th level.   The local number operator %\bea N_{1} \ |\mathbf n \rangle_{1} = \left(\sum_{a=1}^{n_{1}}\n_{a}\right)\ \ |\mathbf n\rangle_{1} \label{numberop1}.  \eea%is defined on ${\cal H}_1$ and can be extended to the full $\cal H$(tensoring with the identity in ${\cal H}_2$).  We call local particlestates the eigenstates of the local number operator $N_{1}$.  Inparticular, let for instance $|0\rangle_{2}$ be the lowest eigenstateof $H_{2}$.  Then the states%\bea | 0 \rangle_{\rm loc} &=& | 0 \rangle_{1} \otimes | 0 \rangle_{2}, \\| i \rangle_{\rm loc} &=& | i\rangle_{1}\otimes | 0\rangle_{2} =   \left(\sum_{a=1}^{n_{1}}\ (U^{\scriptscriptstyle{(n_{1})}})^{-1}_{ia}\ |a\rangle_{1}\right)\otimes | 0\rangle_{2} ,\label{lparticle}\eea%where, as before, $ |a\rangle_{1}$ is the state with a singleexcitation of the $a$-th normal mode of $H_{1}$, are the local vacuumand the local one particle state -with the particle on the $i$-thoscillator, associated to the region $R_{1}$.The two states $| i \rangle$, defined in (\ref{gparticle}), and $| i\rangle_{\rm loc}$, defined in (\ref{lparticle}), are bothone-particle states where the particle is concentrated on the firstoscillator.  The first is the analog of the localized Fock particlestates used in QFT, the second is a state that can be detected by adetector localized in the region $R_{1}$.  Similarly, a detectorlocalized in $R_{1}$ will certainly detect no particles if the systemis in the state $| 0 \rangle_{\rm loc}$, while in QFT we usuallyinterpret a state where no particle has been measured by a localizeddetector as a global vacuum state analogous to $| 0 \rangle$.\section{Convergence between local and global states}What is the error we make in ignoring the difference between local andglobal states?  Clearly we should expect that the differencebetween $| i \rangle_{\rm loc}$ and $| i\rangle$ becomes negligibleif the region $R_{1}$ is sufficiently large, and if $i$ issufficiently distant from the boundary of the region $R_{1}$.  Thisfact allows us to ignore the difference, and to describe the outcomeof local detectors in terms of global particles without errors in ourpredictions.  Thus we expect that% \be| i \rangle_{\rm loc}\hspace{2em} \longrightarrow_{\scriptscriptstyle n,n_{1}\to\infty}\hspace{2em}   | i\rangle.\label{giusta} \ee%One might expect then that%\be\langle i| i \rangle_{\rm loc}\hspace{2em} \longrightarrow_{\scriptscriptstyle n,n_{1}\to\infty}\hspace{2em}   1;\label{sbagliata} \ee%perhaps surprisingly, however, (\ref{sbagliata}) is wrong, as can beshown by an explicit calculation.  The physical reason is that the twostates $| i \rangle_{\rm loc}$ and $| i\rangle$ are always physicaldistinguishable, irrespectively of the size of the regions.  This isbecause the second has correlations across the boundary of the tworegions, which are absent in the first.  Indeed (\ref{giusta}) iscorrect, but in a more subtle sense that (\ref{sbagliata}): whatconverges is just the expectation value of local measurements.  Let usillustrate this point in some detail. Let us illustrate this independence from the size of the regions bycomputing the probability $\cal P$ of finding no particles (namely$H_{1}$ in its lowest eigenstate), if the system is in the globalvacuum $|0 \rangle$, as we did in the case of the two oscillators. Let $P_{0_{\rm loc}}=|0\rangle_{1}\ {}_{1}\langle 0|$ be the projectoron the lowest eigenspace of $H_{1}$.  We have, indicating with $\psi_0({\mathbf{q}})$ the global vacuum state in the coordinate representation,% \bea {\cal P} = \left\langle 0 | P_{0_{\rm loc}}| 0\right\rangle = \int d  {\mathbf{q}} \ {\psi}_0^* ({\mathbf{q}}) \left(   P_{0_{\rm loc}} {\psi_0} \right)({\mathbf{q}}) \eea % where % \be  \left(   P_{0_{\rm loc}} {\psi_0} \right)({\mathbf{q}})= {\psi}_0^*  ({\mathbf{q_{1}}}) \int d\mathbf{q}_{1}' \left(\prod^{n}_{a=1}  \left(\frac{\omega_a}{\pi}\right)^{1/4}\right) \left(\prod^{n_{1}}_{a=1}  \left(\frac{\tilde{\omega}_a}{\pi}\right)^{1/4}\right)  e^{-\frac{1}{2}{\mathbf{q}_{1}'}^{T} D^{\scriptscriptstyle{(n_{1})}}{\mathbf{q}_{1}'}}  e^{-\frac{1}{2}{{\mathbf{q}}}^{'T}  D^{\scriptscriptstyle{(n)}}{{\mathbf{q}}'}}  \ee % where $\tilde\omega_{a}$ are the eigenfrequencies of $H_{1}$, and ${q^i}'=q^i$ for $i=n_1+1,...,n$.  Performing the integrations we obtain%  \be\left\langle 0 | P_{0_{\rm loc}}| 0\right\rangle =  \left(\prod^{n}_{a=1} \sqrt{\frac{\omega_a}{\pi}}\prod^{n_{1}}_{a=1}  \sqrt{\frac{\tilde{\omega}_a}{\pi}}\right)\left(\det\left(\frac{A}{2  \pi}\right)\right)^{-1} \left(\det\left(\frac{B}{2  \pi}\right)\right)^{-1/2} \ee % where $A$ is a $n_{1}\times n_{1}$ matrix and $B$ a $n_{2} \times n_{2}$  matrix with elements%  \bea A_{ij} &=& D^{\scriptscriptstyle{(n_{1})}}_{ij}+  D^{\scriptscriptstyle{(n)}}_{ij}, \   \ \; \; \; \; \;  i,j=1,...,n_{1}; \\ B_{kl} &=& \frac{1}{4} \left(D^{\scriptscriptstyle{(n)}} A^{-1}  D^{\scriptscriptstyle{(n)}}\right)_{kl} +  D^{\scriptscriptstyle{(n)}}_{kl}, \   \ \; k,l=n_{1}+1,...,n. \eea % Expanding for small values of the coupling $\lambda$ leads to  expression % \bea \prod^{n_{1}}_{a=1} \sqrt{\tilde\omega_a} &\approx& 1 - \lambda^2\  \frac{n_{1}-1}{4}  \\ \left(\det\left(\frac{A}{2 }\right)\right)^{-1} &\approx& 1 +  \lambda^2 \left(\frac{n_{1}}{2} - \frac{7}{16} \right) \\ \left(\det\left(\frac{B}{2}\right)\right)^{-1/2} &\approx& 1 +  \lambda^2 \left(\frac{n_{2}}{4} - \frac{1}{8} \right) \eea %And we obtain, to order$O(\lambda^2)$ %\bea {}_{\rm }\langle 0, P_{0_{\rm loc}} \, 0\rangle_{\rm } &=&\left[1- \lambda^2 \left(\frac{n_{1}+n-2}{4}\right)\right] \left[1+\lambda^2 \left(\frac{8n_{1}-7}{16}\right) \right] \left[ 1 +\lambda^2 \left(\frac{n_{2}}{4} - \frac{1}{8} \right)\right] \\&=& 1  - \frac{\lambda^2}{16}. \eea%Observe that this amplitude does not depend on the number ofoscillators in the two chains, in fact, it is equal to the oneobtained in the case $n_1=1$ and $n=2$ in (\ref{0p}). A similar result can be obtained for a one-particle state.  Let$P_{i_{\rm loc}}= | i \rangle_{1}\ {}_{1}\langle i |$ be the projectoron the local one-particle state $ | i \rangle_{1} $ in ${\cal H}_{1}$. The quantity $\langle i| P_{i_{\rm loc}}|i \rangle$ gives theamplitude of seeing the local particle in $i$ with a detectorlocalized in $R_{1}$ if the state is the global one-particle state$|i\rangle$.  Its difference from 1 expresses therefore the error wemake in neglecting the difference between local and global states. Assuming that $2<i<n_1-1$, a straightforward calculation yields, tosecond order in $\lambda$%\bea\langle i | P_{i_{\rm loc}}|i\rangle &=& 1  - \frac{\lambda^2}{16},\label{1p}\eea%which, again, does not depend on $n_{1}$ or $n$ either. In Appendix A, we illustrate this same point using a different technique. Why these results are not in contradiction with the possibility ofusing global states to approximate local states with arbitraryaccuracy?  Because to find an observable capable of distinguishingbetween the local and the global state we have to go to the boundaryof the regions.  The situation is similar to the well-known case oftotal charge in QCD: a state with vanishing total charge is alwaysorthogonal to a state with nonvanishing total charge.  But if a chargeis sufficiently far, its effect is irrelevant on local observables,hence states with different total charge can converge in the weaktopology defined by local observables \cite{Streater}.  Thus thecorrect convergence that describes the physical relation betweenglobal and local states is not in the Hilbert space norm.  It is in aweak topology determined by the local observables themselves.  As anexample, consider the two point function %\beaW(i,j)=\langle 0 | q_i q_j|0\rangle =\langle i | j \rangle \eea%expressing the correlation between two states. If $i$ and $j$ are in $R_{1}$ and sufficiently far from the border, then this quantity converges rapidly to the same quantity computed with local states %\beaW_{\rm loc}(i,j) &=& {}_{\rm loc}\langle i | j \rangle_{\rm loc}.  \eea%In fact, we have %\bea W(i,j) &=& \int d{\mathbf{q}}\prod^{N}_{j=1} \left(\frac{\omega_j}{\pi}\right)^{1/2} \; q_i \; q_j\;e^{-{\mathbf{q}}^{T}D^{\scriptscriptstyle{(n)}}{\mathbf{q}}} =\frac{1}{2} \left(D^{\scriptscriptstyle{(n)}} \right)^{-1}_{ij}.\eea%Therefore, clearly%\beaW(i,j)_{\rm loc} &=& \frac{1}{2} \left(D^{\scriptscriptstyle{(n_{1})}} \right)^{-1}_{ij}\eea%But since for $i,j<n_1$ and for small $\lambda$%\bea \left(D^{\scriptscriptstyle{(n)}} \right)^{-1}_{ij} =\left(D^{\scriptscriptstyle{(n_{1})}} \right)^{-1}_{ij} \approx \left(1+ \frac{3}{4} \lambda^2 \right) \delta_{i,j} - \frac{1}{2} \lambda \;\delta_{i,j\pm 1} + \frac{3}{8} \lambda^2 \; \delta_{i,j\pm 2} +\ldots \eea%it is clear that the two correlation functions $W(ij)$ and $W(ij)_{\rmloc}$ are equal to arbitrary high order in $\lambda$ if $i$ and $j$are sufficiently far from the border.  It is then clear that if theregion $R_{1}$ and we stay sufficiently far from the boundary, localand global particle states are indistinguishable by measuringlocal correlations.\section{Field theory}Finally, let us get to field theory.  We consider for simplicity afree scalar field $\phi(x)$ in two spacetime dimensions, confined in afinite spacial box of size $L$, with reflecting boundary conditions$\phi(0)=\phi(L)=0$.  Dynamics is governed by the hamiltonian  %\beH = \frac{1}{2} \int_{0}^L \left(\pi^2-(\partial\phi)^2+m^2\phi^2 \right).\ee%where $\pi(x)$ is the momentum conjugate to $\phi$. Let $k=1,2,\ldots$ label the (discrete) modes of the system and call$\omega_{k}$ their energy. These are given by %\bea u_{k}(x)= \frac{1}{\sqrt{L \omega_{k}}} \  \sin \left( \frac{k \pi x}{L}\right)\eea%where %\bea\omega_{k}^2 = \frac{k^2 \pi^2}{L^2} + m^2. \eea%Then %\bea u_k(x,t) = u_{k}(x)\ e^{i \omega_{k} t}\eea%is a complex solution of the equation of motion.  We can perform astandard quantization using the operators $a_{k}$ and $a_{k}^\dagger$,associated to these modes,%\bea a_{k} = \int dx\ u_{k}(x) \ \left( \sqrt{2\omega_{k}}\ \phi(x)+i\sqrt{\frac{2}{\omega_{k}}}\ \pi(x)\right)\eea%that give%\bea \phi(x)&=&\sum_{k} \sqrt{\frac{1}{2\omega_{k}}}\ (a_{k}+a^\dagger_{k})\  u_{k}(x)\\\pi(x)&=& i\sum_{k} \sqrt{\frac{\omega_{k}}{2}}\ (a_{k}-a^\dagger_{k}) \ u_{k}(x)\eea%in terms of which the hamiltonian operatorreads%\beH = \sum_{k}\ \omega_{k}\ a^\dagger_{k}\ a_{k}.\ee%Denote $|k\rangle$ the one-particle Fock states with momentum $k$. Global one-particle states are linear combinations of the states$|k\rangle$ %\be|f\rangle = \sum_{k}\ f_{k}\ |k\rangle.\ee%These are eigenstates of the number operator $N=\sum_{k}\a^\dagger_{k}\ a_{k}$ associated with $H$.  We can say that the``position" of the (global) particle is determined by the function%\bef(x,t)=  \sum_{k}\ f_{k}\ u_k(x,t). \ee%Now, consider a particle detector of size $R<L$, located in theregion $\cal R$ defined by $x\in [0,R]$.  Say the detector measuresthe energy contained in the region $\cal R$ defined by%\be H_{R} = \frac{1}{2} \int_{0}^{R}\left(\pi^2-(\partial\phi)^2+m^2\phi^2 \right).  \ee%The quantum operator $H_{R}$ can be written in terms of the operators $a_{k}$ and $a^\dagger_{k}$, giving %\be H_{R} = \sum_{k,k'} (A_{kk'} a^\dagger_{k}a^\dagger_{k'}+B_{kk'} a_{k}a_{k'}+C_{kk'} a^\dagger_{k}a_{k'})\ee%where the matrices $A, B$ and $C$ are easily computed from the eigen-energies and the overlaps%\be U_{kk'} =  \frac{1}{2} \int_{0}^{R} \ u_{k}(x)   \ u_{k'}(x).   \ee%For instance, we have easily%\be A_{kk'} = \frac{k^2k'{}^2+m^2-\sqrt{(k^2+m^2)(k'{}^2+m^2)}}{4\sqrt{\omega_{k}\omega_{k'}}}\ \  U_{kk'} .   \ee%This matrix does not vanish, hence the operator $H_{1}$ does notcommute with the number operator.  It contains $a^\dagger_{k}a^\dagger_{k'}$ terms that take out from the one-particlesubspace, and are analogous to the $ a^\dagger_{1}a^\dagger_{2}$terms that we have encountered in the $V$ term of the two-oscillatorexample.  It follows that one-particle Fock states cannot beeigenstates of this operator.  Therefore when we make a measurementwith a detector that measures the energy $H_{1}$ contained in a finiteregion, we project the state on a subspace of Fock space which is notan $n$-particle Fock state.  To find the eigenstates of $H_{R}$ we can simply compute the modes$u^{R}_{k}(x)$ of the field restricted to the $\cal R$ region, as wedid for the chain of oscillators in the previous section.  These havesupport on the region $\cal R$ where they are given by%\bea u^{R}_{k}(x)= \frac{1}{\sqrt{R \omega^{R}_{k}}} \  \sin \left( \frac{k \pi x}{R}\right)\eea%where%\bea \omega^{R}_{k}{}^2 = \frac{k^2 \pi^2}{R^2} + m^2.  \eea%It is obvious that the eigenstates of $H_{R}$ stillhave a particle like structure, given by the excitations of thesemodes.  In particular, we call \emph{local} one-particle state allsingle excitations of these modes (eigenstates of $H_{R}$) and theirlinear combinations.  More precisely, if $a^{R}_{k}$ and$a^\dagger{}^{R}_{k}$ are the creation and annihilation operatorsfor the  $u^{R}_{k}(x)$ modes, defined by %\bea a^R_{k} = \int dx\ u^R_{k}(x) \ \left( \sqrt{2\omega^R_{k}}\ \phi(x)+i\sqrt{\frac{2}{\omega^R_{k}}}\ \pi(x)\right)\eea%we have %\bea H_R =  \sum_{k}\ \omega^R_{k}\ a^R{}^\dagger_{k}\ a^R_{k}.\eea%We define the number operator%\be N_{R}= \sum_{k} \ a^\dagger{}^{R}_{k}\  a^{R}_{k}\ee%and we interpret it as the observable giving the number of particlesdetected by a detector confined in the region $R$.  The \emph{local}particle states are defined as the eigenstates of $N_{R}$. It is then clear from the discussion that local and global particlestates are distinct.  The firsts represent the states actuallymeasured by finite size detectors.  The seconds are the ones weroutinely use in QFT calculations.\subsection{Convergence between local and global particle states}As we did for the chain of oscillators, it is not hard to show thatlocal and global particle states do not converge in norm when $L$ and$R$ are large.  As before, however, the correlation functions definedby the local particle states converge to the ones defined by theglobal particle states.  We shall now show that this is indeed thecase.To see this, we show that correlation functions in a box of size $L$converge to the ones computed on Minkowski space as $L$ becomes large. It follows that both the correlation functions of the global and localparticles converge to the same value (the correlation of the freefield on Minkowski) for large $L$ and $R$, hence they converge to eachother.  We work below in the case of a field of mass $m$.  In thiscase, large $L$ and $R$ means large with respect to the Comptonwavelength $\lambda_{c}=1/m$ of the particle.  The convergence isexponential in the ratio $\lambda_{c}/R$.  Notice that this impliesthe convergence is extremely good for any macroscopic detector of size$R$.  The massless case is treated in Appendix C. For completeness, inAppendix B, we discuss also the case of a lattice field theory, whichbridges between the chain of oscillators considered above and fieldtheory.We want to compute the Green function for a scalar field with a mass$m$ quantized in a one-dimensional box of side $L$.  Let us indicatewith $| 0_L \rangle $ the vacuum state of the field, then the twopoints function is defined as%\bea \langle 0| \phi(x,t) \phi(x',t') | 0 \rangle &=&\sum_{k=-\infty}^{+\infty}\frac{1}{{L \omega_{k}}} \sin \left( \frac{k \pix}{L} \right) \sin \left( \frac{k \pi x'}{L} \right) e^{i \omega_{k} (t-t')}.\eea%Notice the Weyrich's formula%\bea \frac{e^{ik \sqrt{r^2+x^2}}}{\sqrt{r^2+x^2}} = \frac{i}{2}\int^{+ \infty}_{- \infty} e^{i \tau x} H^{(1)}_{0} \left( r\sqrt{k^2- \tau^2} \right) \, d \tau \eea%valid for $r$ and $x$ real and $0 \leq \arg \sqrt{k^2- \tau^2} < \pi$and $ 0 \leq \arg k < \pi$.  $H^{(1)}_{0}$ is the Hankel function ofthe first kind with index zero.  Using it, we can write%\bea \frac{\exp \left({i(t-t') \sqrt{m^2+(k \pi/L)^2}}\right)}{\sqrt{m^2+(k \pi /L)^2}} = \frac{i}{2} \int^{+\infty}_{- \infty} e^{i \tau k \pi /L} H^{(1)}_{0} \left( m\sqrt{(t-t')^2- \tau^2} \right) \, d \tau \eea%hence%\be \langle 0 | \phi(x,t) \phi(x',t') | 0 \rangle =\frac{i}{{L}} \sum_{k=0}^{+\infty} \int d \tau\ \sin\frac{k \pix}{L}\ \sin\frac{k \pi x'}{L}\ \cos\frac{\tau k \pi}{L}\ H^{(1)}_{0} \left( m\sqrt{(t-t')^2- \tau^2} \right). \ee%We focus the attention on the summation on $k$%\bea \sum_{k=0}^{+\infty} \sin\frac{k \pi x}{L}\ \sin\frac{k \pi x'}{L}\ \cos\frac{\tau k \pi}{L}\ &=& \frac{1}{4} \sum_{k=0}^{+\infty} \left[ \cos\left(\frac{k\pi}{L}(x-x'+ \tau) \right) +\cos\left(\frac{k \pi}{L}(x-x'- \tau)\right) + \right.  \nonumber\\&& \left.  - \cos\left(\frac{k \pi}{L}(x+x'+ \tau) \right)-\cos\left(\frac{k \pi}{L}(x+x'- \tau) \right)\right] .\eea%Consider the sum of the first cosine%\be \sum_{k=0}^{+\infty} \cos\frac{k \pi(x-x'+ \tau)}{L} =\frac{1}{2} \left( \frac{1}{1- \exp\left(i\frac{\pi}{L}(x-x'+\tau+i \epsilon) \right)} + \frac{1}{1- \exp\left(-i\frac{\pi}{L}(x-x'+ \tau-i \epsilon) \right)} \right) \ee%where a small imaginary part has been added in the exponential inorder to make the summation convergent. We have %\bea \sum_{k=0}^{+\infty} \sin \left( \frac{k \pi x}{L} \right) \sin\left( \frac{k \pi x'}{L} \right) \cos \left(\frac{\tau k \pi}{L}\right) &=& \frac{1}{8} \left( \frac{1}{1-e^{i\frac{\pi}{L}(x-x'+\tau+i \epsilon)} } + \frac{1}{1-e^{-i\frac{\pi}{L}(x-x'+\tau-i \epsilon)} } + \right.  \nonumber\\&& + \frac{1}{1- e^{i\frac{\pi}{L}(x-x'-\tau+i \epsilon)} } +\frac{1}{1- e^{-i\frac{\pi}{L}(x-x'-\tau-i \epsilon)} } + \nonumber\\&& - \frac{1}{1- e^{i\frac{\pi}{L}(x+x'+\tau+i \epsilon)} } -\frac{1}{1- e^{-i\frac{\pi}{L}(x+x'+\tau-i \epsilon)} } + \nonumber\\&& - \left.  \frac{1}{1- e^{i\frac{\pi}{L}(x+x'-\tau+i \epsilon)} } -\frac{1}{1- e^{-i\frac{\pi}{L}(x+x'-\tau-i \epsilon)} }\right). \eea%For simplicity we consider the case $t=t'$.  It is useful to expressthe Hankel function with the following integral representation%\bea H^{(1)}_{0} \left(i m |\tau | \right) = \frac{2}{i \pi} K_0 ( m|\tau |) = \frac{2}{i \pi} \int^{\infty}_{0}\frac{\cos (|\tau |y)}{\sqrt{m^2+y^2}}\ dy \eea%where $K_0$ is the MacDonald function.  The Green function becomes%\bea \langle 0 | \phi(x,t) \phi(x',t) | 0 \rangle = \frac{1}{4L\pi} \int^{\infty}_{0} dy \frac{1}{\sqrt{m^2+y^2}} \int^{+ \infty}_{-\infty} d \tau \left( \frac{\cos (|\tau|y)}{1-e^{i\frac{\pi}{L}(x-x'+\tau+i \epsilon)}} + similar \; terms \right).\label{2point}\eea%The integrals in $\tau$ are calculated going in the complex plane of$\tau$.  The choice of the closure of the contour depends on the signof $\tau$ in the exponential of the denominator.  So, for the firstintegral, we close the path on the lower half-plane obtaining thecontour $\cal{C}$, yielding%\bea \int^{+ \infty}_{- \infty} d \tau \frac{\cos (|\tau|y)}{1-e^{i\frac{\pi}{L}(x-x'+\tau+i \epsilon)}} = \int_{\mathcal{C}} d \tau\frac{\cos (|\tau|y)}{1- e^{i\frac{\pi}{L}(x-x'+\tau+i \epsilon)}}\eea%The integrand has an infinite number of poles in $\tau = x'-x + 2nL$,where $n\in N$.  Applying the theorem of residue we find%\bea \int_{\mathcal{C}} d \tau \frac{\cos (|\tau|y)}{1-e^{i\frac{\pi}{L}(x-x'+\tau+i \epsilon)}} = 2L \sum_n \cos\left(|x-x'+2nL|y \right).  \eea%And analogous results are found for the other integrals.  Insertingthese results in (\ref{2point}) we arrive at the final expression forthe Green function%\bea \langle 0_L | \phi(x,t) \phi(x',t) | 0_L \rangle &=& \frac{2}{\pi}\sum_n \left[ K_0 ( m |x-x'+ 2nL|) - K_0 ( m |x+x'+2nL|)\right].\label{2massive}\eea%When the size of the box is much greater than Compton wavelength ofthe scalar particle, namely $1/m$ in unit $\hbar=c=1$, we can distinguish two cases:$\bullet\ $ If $0 \ll x,x' \ll L$; in this case, in the limit $x \rightarrow x'$the main contribution to the correlation function comes from the firstMacDonald function in (\ref{2massive}).  In fact, for $n \neq 0$, wecan expand the MacDonald function for large argument, due to thecondition $mL \gg 1$%\bea K_0 ( m |x\pm x'+ 2nL|) \approx \sqrt{\frac{\pi}{2( m |x \pm x'+2nL|)}}e^{- m |x\pm x'+ 2nL|} \eea%so that $K_0$ has an exponential decay with the length scaleproportional to the Compton wavelength of the particle.  Thereforethis is a negligible contribution.  In contrast, for $n=0$, the factthat the MacDonald function diverges when the argument tends to zeroimplies%\bea K_0(m |x-x'|) \gg K_0(m |x+x'|) \eea%Consequently we can write%\bea \lim_{x \rightarrow x'} \langle 0 | \phi(x,t) \phi(x',t) | 0\rangle = \frac{2}{ \pi}K_0 ( m |x-x'|) \propto \langle 0_{M} | \phi(x,t)\phi(x',t) | 0_{M} \rangle \eea%where the state $|0_{M} \rangle$ is the vacuum state for the scalar fieldquantized in Minkowski spacetime.$\bullet\ $ If, on the other hand, $x \sim x' \sim 0$ or $x \sim x' \sim L$; inthis case, both Macdonald functions in (\ref{2massive}) contributesignificantly to the correlation function.  In particular, $K_0(m|x+x'+2nL|)$ is not negligible when $x,x' \approx 0$ for $n=0$, andwhen $x,x' \approx L$ when $n=-1$.  This means that when the twopoints considered are near the boundary of the box, the correlationfunction feels the present of the box, and differs from thecorrelation function defined in the whole Minkowski space.This result illustrate how, if we stay away from the boundary and ifthe region $R$ sufficiently large, correlation functions computedwith local states converge to the ones computed with global states. \section{Conclusion}We have argued that the particles detected by real measuring apparatusare local objects, in the sense that they are best represented by QFTstates that are eigenstates of local operators.  We have defined thesestates, and denoted them \emph{local particle states}.This is not what is usually done in QFT, where, instead, we representthe particles observed in particle detectors by means of a differentset of states: global particle states such as the $n$-particleFock states.  Global particle states provide a good approximation to local particlestates.  The convergence is not in the Hilbert space norm, but ina weak topology given by local observables.  The approximation isexponentially good with the ratio of the particle Compton wavelengthwith the size of the detector, and the distinction between global andlocal states can therefore be safely neglected in concreteutilizations of QFT. However, the distinction is conceptually important because it bears onthree related issues: (i) whether particles are local or globalobjects in conventional QFT; (ii) the extent to which the quantumfield theoretical notion of particle can be extended to generalcontexts where gravity cannot be neglected; and furthermore, more ingeneral, (iii) whether particles can be viewed asthe fundamental reality (the ``ontology") described by QFT. Let usdiscuss these three issues separately.(i) The distinction shows that in the context of conventional QFT theglobal properties of the particle states are an artifact of anapproximation taken, not an intrinsic property of physically observedparticles.  We view this as a simple and clear answer to the firstquestion we have addressed: whether particles are local or globalobjects in QFT.\footnote{To avoid misunderstanding, let us emphasizethe fact that we are not referring here to limitations of the theory. We have assumed here QFT to be exact.  Given a theoretical descriptionof the world, such as QFT on Minkowski space, it is important todistinguish three different levels: (i) the world, which is mostpresumably not exactly described by the theory (spacetime is curved,gravity is quantized\ldots), (ii) the ensemble of the empirical datato which we have access, with their given accuracy; (iii) the theory. We have then two distinct problems.  One is the empirical adequacy ofthe theory to the ensemble of data.  To be adequate, a theory does notneed to be an exact and complete description of the world; it issufficient that it correctly reproduces observations within theavailable accuracy.  For instance, we can describe observed waves on alake using a theory of waves on a \emph{flat} water surface, or on aspherical water surface, or on an ellipsoidal water surface\ldots Thefirst of these options can be perfectly empirically adequate, even isthe Earth isn't actually flat.  Once the theory is chosen, however,there is then a second issue: the precise identification betweentheoretical quantities and empirical data.  \emph{This} is the issuewe have discussed in this paper.  We have placed ourselves in aregime, or under the assumption, that flat space QFT is empiricallyadequate, but we have reconsidered how it should precisely beinterpreted.  Given a particle observed physically, what is the stateof the theory that best describes it?  Our suggestion is that theusual answer (a Fock particle state) can be replaced with another one(a local particle state) which is more coherent with the basic rulesof quantum mechanics (because the result of a local measurement has tobe interpreted as an eigenstate of the corresponding operator) andbears on the possibility of extending QFT methods to more generalcontexts.}%(ii) More importantly, the distinction bears on the general validityof the notion of particle, and on the possibility of utilizing it inthe context in which gravity cannot be neglected.  In so far asparticles are understood as global objects, tied to globalsymmetries of spacetime, their utilization outside flat space isdifficult.  In the context of a curved spacetime and, more in general,in a background independent context where there is no Poincar\'einvariant background spacetime, the notion of global particle state isambiguous, ill defined, or completely impossible to define.  Asmentioned in the introduction, this has lead several theoreticians toconsider interpretations of QFT where particles play no role.  But ifwe can understand particles as eigenstates of local operators, with noreference to global features, then it is clear that we have analternative notion of particle that has all the chances to be welldefined in general.  On a general curved spacetime, a finitely extended detector thatmeasures the energy $H_{R}$ contained in a finite region of space$\cal R$ and in a given reference frame, will detect local statesdetermined by eigenstates of (the Heisenberg operator) $H_{R}$.  Thesewill have a particle-like structure.  Indeed, they correspondprecisely to the states that best describe flat space QFT measurementsas well (instead of the Fock $n$-particle states).  Thus,\emph{global} particle states do not generalize, but \emph{local}particle states, that truly describe what we measure in a bubblechamber, do.  

The extension of our results to interacting theories should 
be trivial  when the absence of correlation between 
measurements performed in distant regions is assured by the 
cluster decomposition property \cite{weinberg}, but we expect 
the main point to hold in general. It particular, the results
of this paper strongly support the viability of the idea of using
a notion of particle also in the context of the boundary formulation
of quantum field theory \cite{boundary,book}, which is at the
root of the recent calculations of $n$-point functions in quantum gravity
\cite{transitions}.  This formalism provides the possibility to associate state spaces 
with arbitrary hypersurfaces of spacetime by encoding the information 
on the physical processes taking place within a spacetime region into 
the amplitude associated with states on its boundary hypersurface. 
Thus, our conclusion is that the absence of well-defined globalparticle states, Poincar\'e invariance, or a preferred vacuum state,has no bearing on the possibility of interpreting QFT in terms ofparticles.  Putting it vividly (but naively) we could say: localparticle detectors detect particles also on a curved or quantizedspacetime.  This leads us to the third issue, which is more``philosophical", and on which we offer only a few thoughts, without any pretension of rigor or completeness. (iii) Can we view QFT, in general, as a theory of particles?  Can wethink that reality is made by elementary objects -the particles- whoseinteractions are described by QFT? We think that our results suggestthat the answer is partially a yes and partially a no.  We have argued that \emph{local} particle states can be defined ingeneral.  In this sense, we share the point of view that QFT can beinterpreted as a theory of particles quite generally.On the other hand, however, it is clear from the discussion given thatthe particles described by the $n$-particle Fock states areidealizations that do not correspond to the real objects detected inthe detectors.  Moreover, they have unpalatable global properties. Therefore it is very difficult to view \emph{them} as the fundamentalobjects described by QFT. In particular, there is no reason forinterpreting the Fock basis as ``more physical" or ``more close toreality" than any other basis in the state space of QFT. Fockparticles aren't more fundamental objects than eigenstates of anyother operators.  If anything, they are \emph{less} fundamental,because we never measure the Fock number operator.  Interpreting QFTas the theory of physical objects described by the $n$-particle Fockstates, with their global features, is not only a stumbling bocktoward potentially useful generalizations of flat space QFT, but itis also in contradiction with what we have learned about the worldwith quantum theory.Can we base the ontology of QFT on local particles?  Yes, but localparticle states are very different from global particle states. Global particle states such as the Fock particle states are definedonce and for all in the theory, while each finite size detectordefines its own bunch of local particle states.  Since in general theenergy operators of different detectors do not commute($[H_{R},H_{R'}]\ne 0$), there is no unique ``local particle basis" inthe state space of the theory, as there is a unique Fock basis. Therefore, we cannot interpret QFT by giving a single list of objectsrepresented by a unique list of states.  In other words, we are in agenuine quantum mechanical situation in which distinct particlenumbers are complementary observables.  Different bases thatdiagonalize different $H_{R}$ operators have equal footing.  Whether aparticle exists or not depends on what I decide to measure.  In such acontext, there is no reason to select an observable as ``more real"than the others.The world is far more subtle than a bunch of particles that interact.\vskip2cm\appendix\section*{Appendix A: Density matrix}In this appendix, we give a different description of the relationbetween local and global states, by using a density matrix technique. We can obtain all probabilities for measurements performed in a region$R_{1}$ in terms of a reduced density matrix which is a function of thesole degrees of freedom in $R_{1}$.  If the state has correlationsbetween two regions $R_{1}$ and $R_{2}$, the corresponding reduceddensity matrix is not the one of a pure state.Consider the two oscillators system described in Section 2.  Thedensity matrix of the global vacuum is $\rho=|0\rangle\ \langle 0|$. In coordinate space, it reads%\bea\rho(q_1,q_2,q_1',q_2') &=& \langle q_1,q_2|0\rangle\ \langle 0|q_1',q_2'\rangle) \\&=& \frac{\sqrt{\omega_a \omega_b}}{{\pi}} \exp \left( -\frac{1}{2}\frac{\omega_a + \omega_b}{2} (q_1^2+q_2^2+q_1'{}^2+q_2'{}^2) -\frac{\omega_a  - \omega_b}{2} (q_1 q_2 +q_1' q_2') \right).\nonumber\eea%Tracing on the $q_{2}$ variable yields the reduced density matrix%\bea\rho_{red} (q_1,q_1') &=& \int dq_2 \; \rho(q_1,q_2,q_1',q_2) \\&=&  \sqrt{\frac{2\omega_a \omega_b}{\pi (\omega_a + \omega_b)}} \exp \left( -\frac{1}{2}\frac{\omega_a  + \omega_b}{2} (q_1^2+q_1'{}^2) -\frac{(\omega_a  - \omega_b)^2}{8 (\omega_a + \omega_b)} (q_1  +q_1' )^2 \right).\eea%This density matrix satisfies the properties $Tr(\rho_{red})=1$ as it must be for every density matrix and%\beaTr(\rho_{red}^2) =  \frac{2\sqrt{\omega_a \, \omega_b}}{\omega_a + \omega_b} \approx 1 -\frac{\lambda^2}{4 \omega^4}\eea%upon the expanding for small $\lambda$, showing that it is a densitymatrix of a mixed state, namely that there are vacuum correlationsbetween the two oscillators.  Suppose now that we disregard the second oscillator all together and weconsider the $q_{1}$ system alone, with the hamiltonian $H_{1}$.  If we measure the energy $E_{1}$ and find the system in the lowest eigenstate, then the system will be described by the density matrix %\bea\rho(q_1,q_1')=(\omega/\pi)^{1/2} \exp \left( -\omega (q_1^2+q_1'{}^ 2)/2\right)\eea%of the pure vacuum state of the single oscillator $q_1$.  The relationbetween the two density matrices is, to the first nontrivial order in$\lambda$%\bea\rho_{red} (q_1,q_1') \approx \rho (q_1,q_1') \; \left(1 - \frac{\lambda^2}{8 \omega^4}\left(1-\frac{\omega}{2}(q_1+q_1')^2 \right) \right).\eea%Thus, the expectation value of any observable calculated with thisreduced density matrix differs from the one evaluated with the densitymatrix of the pure state by a term proportional to $\lambda^2$.  Thisgives a precise general evaluation of the difference between the local and global vacuum states. Let us then repeat this calculation in the case of the chain ofoscillators.  The density matrix for the chain of $n$ oscillators canbe written as%\bea\rho(\mathbf q, \mathbf q') = \psi_0(\mathbf q) \psi^*_0( \mathbf q').\eea%To obtain the reduced density matrix on the region $R_{1}$ we musttrace over the variables $\mathbf q_2$, obtaining the reducedmatrix%\bea{\rho}_{red}(\mathbf q_1,\mathbf q_{1}') = \int d\mathbf q_2\;{\psi}_0(\mathbf q_1,\mathbf q_2) \;{\psi}^*_0(\mathbf q'_1,\mathbf q_2). \eea%Computing the integral and expanding for small values of $\lambda$, we find%\bea{\rho}_{red}(\mathbf q_1,\mathbf q_{1}') &=& \pi^{-n_{1}/2} \left( \det D^{\scriptscriptstyle{(n)}}\right)^{1/2}\left(\det C(D^{\scriptscriptstyle{(n)}}) \right)^{-1/2} \times \\&& \times \exp \left( -\frac{1}{2}\mathbf{q_1}^T D^{\scriptscriptstyle{(n)}} \mathbf{q_1}-\frac{1}{2}\mathbf{q_1}'{}^T D^{\scriptscriptstyle{(n_{1})}} \mathbf{q_1}_1'+\frac{\lambda^2}{16}(q_1^{n_1}+((q'_1)^{n_1})^2 \right) \nonumber\\&=&\pi^{-n_{1}/2} \left( \det D^{\scriptscriptstyle{(n)}} \right)^{1/2}\left(\det C( D^{\scriptscriptstyle{(n)}}) \right)^{-1/2} \times \nonumber\\&& \times \exp \left( -\frac{1}{2}\mathbf{q_1}^TD^{\scriptscriptstyle{(n_{1})}} \mathbf{q_1}-\frac{1}{2}\mathbf{q_1}'{}^TD^{\scriptscriptstyle{(n_{1})}}\mathbf{q_1}'+\frac{\lambda^2}{8}((q_1^{n_1})^2+((q_1^{n_1})')^2+q_{n_1} q_{n_1}') \right) \nonumber\eea%where $C(D^{\scriptscriptstyle{(n)}})$ is the minor of the matrix $D^{\scriptscriptstyle{(n)}}$ with first element $(D^{\scriptscriptstyle{(n_{1})}})_{n_{1}+1,n_{1}+1}$. It is not difficult to check that %\beaTr(\rho)= Tr({\rho}_{red})=1\eea%and $Tr(\rho^2)=1$ as should be for a pure state.  On the other hand,${\rho}_{red}$ is a density matrix of a mixed state, indeed%\beaTr({\rho}_{red}^2) \approx 1 - \frac{1}{8}\lambda^2.\eea%Its entropy  is%\beaS= -Tr \left( \rho_{red} \ln (\rho_{red}) \right)  \ \approx\ \frac{1}{16} \lambda^2.\eea%This entropy can be calculated expanding the density matrix in basisof the eigenfunction of the $N$ coupled oscillators and noting thatthe leading term in this expansion is the component on the vacuumstate.  This expression gives a quantitative expression of thedifference between the local and global state.  Notice that it doesnot go to zero for large $n$ and $n_{1}$.\section*{Appendix B: Lattice scalar field}A free scalar field can be modeled as a collection of coupled harmonicoscillators located on a lattice of space points $x$.  We consider thedynamical system defined by the hamiltonian%\bea H= \sum_{i=1}^{N}\frac{1}{2}\dot{q}_i^2 +\frac{1}{2}\sum_{i=1}^{N-1} \omega^2 \left( q_i+ q_{i+1} \right)^2\eea%where we have fixed the total number $N$ of oscillators.  The normalfrequencies of this system are%\bea \omega_{n} = \sqrt{2} \; \omega \sin \left(\frac{n \pi}{2(N+1)}\right).  \eea%Using the same notations as in Section 3, we express thecorrelation function as%\bea W(i,j) = \left\langle \hat{q}_i\hat{q}_j \right\rangle &=& \frac{1}{2}\left(D\right)^{-1}_{ij} = {\frac{1}{N+1}} \sum_{k} \frac{\sin \left( i \theta_k \right)\sin\left(j \theta_k \right)}{\sqrt{2} \; \omega \sin \left({\theta_k}/{2} \right)}\label{corrlatt}\eea%where $\theta_k ={k \pi}/{(N+1)}$.  The fraction in (\ref{corrlatt})can be reexpressed in the form%\bea \frac{\sin(i\theta_k) \sin(j \theta_k)}{\sin(\theta_k/2)}=\frac{\cos((i-j)\theta_k) - \cos((i+j) \theta_k)}{\sin(\theta_k/2)}. \eea%The following relation holds%\bea \frac{\cos((i-1)\theta_k) - \cos(i \theta_k)}{\sin(\theta_k/2)} =2 \sin((i-1/2)\theta_k); \eea%using which, we have%\bea \frac{\cos((i-j)\theta_k) - \cos((i+j)\theta_k)}{\sin(\theta_k/2)} = 2 \sum^{i+j}_{p=i-j+1}\sin((p-1/2)\theta_k).  \eea%Inserting the last expression into (\ref{corrlatt}) and inverting thesum over $p$ and $k$, we obtain%\bea W(i,j) &=&{\frac{\sqrt{2}}{\sqrt{\omega}(N+1)}} \sum_{k=1}^{N}\sum^{i+j}_{p=i-j+1} \sin((p-1/2)\theta_k) \\&=& {\frac{1}{\sqrt{2\omega}(N+1)}}\sum^{i+j}_{p=i-j+1} \cot\frac{(2p-1) \pi}{4(N+1)}.  \eea%For $N \gg i,j$ we can expand the cotangent function for smallargument, $\cot x = 1/x +o(x^{-3})$:%\beaW(i,j) &\approx& {\frac{4(N+1)}{\pi \sqrt{2\omega}(N+1)}}\sum^{i+j}_{p=i-j+1} \frac{1}{2p-1} \nonumber\\&\approx& \frac{2 }{\pi \sqrt{2\omega}} \left[ \Psi_0(i+j+1/2)-\Psi_0(i-j+1/2)\right], \eea%where $\Psi_0$ is the digamma function.  We obtain a correlationfunction independent from the total number $N$ of oscillators. \section*{Appendix C: Massless field}We extend the analysis of the massive scalar field correlationfunctions, given in section 5, to the massless case.  The masslesscase is more delicate, because of the infrared divergences, andbecause of the conformal invariance of the 2d massless theory.  Herewe give explicitly the correlation functions on a finite box, leavinga detailed physical discussion for further developments.A massless scalar field in the region $x\in[0,L]$ can be expanded inthe modes%\bea \phi(x,t)= \sum_{k=-\infty}^{+\infty} \left( a_k u_k(x,t) +a_k^*u_k^*(x,t) \right) \eea%where the functions $u_k(x,t)$ are solutions of the Klein-Gordonequation.  We want to quantize the field inside a (one dimensional)box of length $L$ and consequently we impose the following boundaryconditions on the modes of the field:%\bea u_k(0,t)=u_k(L,t)=0; \eea%therefore these functions result to be%\bea u_k(x,t) = \frac{1}{\sqrt{L E_k}} \sin \left( \frac{k \pi x}{L}\right) e^{i E_k t}; \qquad \hbox{with} \qquad E_k = \frac{|k| \pi}{L}\eea%The modes $u_k(x,t)$ form a complete orthonormal basis with respect tothe scalar product%\bea \left( u_k(x,t),u_{k'}(x,t) \right) = i \int_0 ^L \left( u_k(x,t)\partial_t u^* _{k'}(x,t) - [ \partial_t u_k(x,t)]u^* _{k'}(x,t)\right) dx.  \eea%The quantization of the scalar field promotes the coefficients $a_k$and $a_k^+$ to the status of annihilation and creation operatorsrespectively.  These operators define the vacuum state of the field inthe box, called $| 0 \rangle $.  The two-points function can bewritten as%\bea \langle 0 | \phi(x,t) \phi(x',t') | 0 \rangle &=&\sum_{k=-\infty}^{+\infty}\frac{1}{{L E_k}} \sin \left( \frac{k \pix}{L} \right) \sin \left( \frac{k \pi x'}{L} \right) e^{i E_k (t-t')}\\&=& \sum_{k=0}^{+\infty}\frac{1}{{L E_k}} \left[ \cos \left( \frac{k\pi (x-x')}{L} \right) - \cos \left( \frac{k \pi (x+x')}{L} \right)\right] e^{i E_k (t-t')} \nonumber\\&=& \frac{1}{2} \sum_{k=0}^{+\infty}\frac{1}{{k \pi}} \left[ e^{\frac{ik \pi (x-x')}{L} }+e^{ \frac{-ik \pi (x-x')}{L} } - e^{\frac{ik \pi (x+x')}{L} }- e^{ \frac{-ik \pi (x+x')}{L} } \right]e^{\frac{ik \pi (t-t')}{L} }. \nonumber \eea%We must compute sums of the kind%\bea \sum_{k=1}^{\infty} \frac{e^{iak}}{k} = -\ln (1-{e^{ia}}). \eea%Using this,%\bea \langle 0 | \phi(x,t) \phi(x',t') | 0 \rangle &=& \frac{-1}{2\pi} \left[ \ln (1-{e^{\frac{i\pi(x-x'+t-t') }{L}}})+\ln(1-{e^{\frac{i\pi(x'-x+t-t') }{L}}}) \right.  \nonumber\\&-& \left.  \ln (1-{e^{\frac{i\pi(x+x'+t-t') }{L}}})-\ln(1-{e^{\frac{i\pi(-x-x'+t-t') }{L}}}) \right].\eea%To simplify the discussion we focus on theequal time $(t=t')$ correlation function%\bea \langle 0 | \phi(x,t) \phi(x',t) | 0 \rangle &=& \nonumber \\ &&\hspace{-9em} = \ \frac{-1}{2\pi} \left[ \ln (1-{e^{\frac{i\pi(x-x') }{L}}})+\ln(1-{e^{-\frac{i\pi(x-x') }{L}}}) - \ln (1-{e^{\frac{i\pi(x+x')}{L}}})-\ln (1-{e^{-\frac{i\pi(x+x') }{L}}}) \right] \nonumber\\&&\hspace{-9em} =\ \frac{-1}{2 \pi} \left[ \ln \left(2-2 \cos \left( \frac{\pi(x-x')}{L}\right) \right) - \ln \left(2-2 \cos \left( \frac{\pi(x+x')}{L}\right)\right) \right].\label{2massless}\eea%When $0 \ll x,x' \ll L$ and in the limit $x \rightarrow x'$, the maincontribution comes from the first logarithm in (\ref{2massless}):%\be \langle 0 | \phi(x,t) \phi(x',t) | 0 \rangle \ \ \approx\ \ \frac{-1}{2 \pi} \ln \left({{\frac{\pi^2(x-x')^2 }{L^2}}} \right)=-\frac{1}{\pi} \ln|x-x'| + \frac{1}{\pi} \ln (L/\pi) \label{limit2massless}\ee%Notice that the dependence on the size of the box $L$ appears only asan additive constant.  To shed light on the meaning of this constant,recall that in Minkowski spacetime the massless correlation functionis a divergent quantity that needs to be regularized with theintroduction of an infrared cut-off, say $1/N$%\bea \langle 0 | \phi(x,t) \phi(x',t) | 0 \rangle = \frac{-1}{4 \pi}\ln \left( \frac{(x-x')^2}{N^2} \right) =-\frac{1}{2\pi} \ln|x-x'| + \frac{1}{2\pi} \ln N . \eea%One is then interested in physically observable cut-off independentquantities.  The cut-off dependence is precisely via an additiveconstant, namely the same as the dependence on the box size in(\ref{limit2massless}); in fact the box provides an infraredregularization of the correlation function for the massless field. Viceversa, an infrared cut-off can be interpreted precisely as thefinite detector size, as in this paper.  This relates the problem ofthe relation between (massless) local and global particles to theusual discussion of the relation between infrared divergences and(independence from) finite size detector effects.When the points considered are close to the boundaries of the box,i.e. $x \sim x' \sim 0$ or $x \sim x' \sim L$, the second logarithm in(\ref{2massless}) is no more negligible with to the first logarithm,and therefore the correlation function is sensibly different from theone defined in Minkowski space.\begin{thebibliography}{9}\bibitem{bd} ND Birrel, PCW Davies, {\it Quantum fields in curvedspace}, (Cambridge University Press, Cambridge 1982).\bibitem{wald} RM Wald, {\it Quantum Field Theory on Curved Spacetimeand Black Hole Thermodynamics} (University of Chicago Press, Chicago1994).
\bibitem{wigner} EP Wigner, ``On unitary representations of the inhomogeneous Lorentz group", Ann of Math {40} (1939), 149-204.\bibitem{Davies} PCW Davies, ``Particles do not exist'' in \textit{Quantum Theory of Gravity: Essays in honor of the 60th birthday of Brice DeWitt}, edited by SM Christensen,  pp 66-77, (Adam Hilger, Bristol 1984).\bibitem{weinberg} S Weinberg, \textit{The Quantum Theory of Fields}, (Cambridge University Press, 1995). \bibitem{book}C Rovelli, {\em Quantum Gravity}, (Cambridge University Press,Cambridge, 2004). 
T Thiemann, ``Modern canonical quantum general relativity,'' (Cambridge
  University Press, Cambridge, UK, 2007). 

\bibitem{lqg}  
C Rovelli, L Smolin,
 ``Knot theory and quantum gravity'' 
{Phys Rev Lett} {61} (1988) 1155-1158. 
%%CITATION = PRLTA,61,1155;%%
C Rovelli, L Smolin,
``Loop space representation for quantum general relativity", 
{Nucl Phys.} {B331}  (1990) 80-152.
 %%CITATION = NUPHA,B331,80;%%
 A Ashtekar, C Rovelli, L Smolin,
 ``Weaving a classical geometry with quantum threads",
{Phys Rev.\ Lett}  {69}  (1992) 237.
  %%CITATION = PRLTA,69,237;%%
C Rovelli, L Smolin,
 ``Discreteness of Area and Volume in Quantum Gravity", 
{Nucl Phys.} {B442} (1995)  593-619;
 %%CITATION = GR-QC 9411005;%%
{Nucl Phys.} {B456}  (1995) 734.
\bibitem{gravitons}
J Iwasaki, C Rovelli, ``Gravitons as embroidery on the weave", 
Int. J. Mod. Phys. D1 (1993) 533-557.
\bibitem{transitions} 
L Modesto, C Rovelli, ``Particle scattering inloop quantum gravity", Phys. Rev. Lett.  {95}, 191301 (2005).
C Rovelli:``Graviton propagator from background-independent quantum gravity", Phys Rev Lett 97 (2006) 151301. 
E Bianchi, L Modesto, C Rovelli, S Speziale: 
``Graviton propagator in loop quantum gravity", 
Classical and  Quantum Gravity 23 (2006) 6989-7028.
  E.~Alesci and C.~Rovelli,
  ``The complete LQG propagator: I. Difficulties with the Barrett-Crane
  vertex,''
  Phys. Rev.  D76, 104012 (2007).
E Alesci, C Rovelli,
  ``The Complete LQG propagator. II. Asymptotic behavior of the vertex",  
Phys. Rev. D77,  044024 (2008). 
C Perini, C Rovelli, S Speziale,
  ``Self-energy and vertex radiative corrections in LQG,''   arXiv:0810.1714.
  
\bibitem{boundary} 
 R Oeckl, ``A `general boundary' formulation for quantum mechanics and
 quantum gravity, Phys. Lett. B 575 (2003) 318--324.
R Oeckl,
``General boundary quantum field theory: Foundations and probability 
interpretation", Adv. Theo. Math. Phys. to appear,  hep-th/0509122.
R Oeckl, ``Probabilites in the general boundary formulation", J. Phys.: Conf. Ser.,
67 (2007) 012049.
D Colosi, R Oeckl, 
``Spatially asymptotic S-matrix from general boundary formulation",
Phys. Rev., D 78 (2008) 025020. 
D Colosi, R Oeckl, ``S-matrix at 
spatial infinity", Phys. Lett. B 665 (2008) 310--313.
  
  \bibitem{Unruh:db} WG Unruh, ``Notes On Black Hole Evaporation'', Phys Rev D {14} (1976) 870.\bibitem{Streater}RF Streater and AS Wightman, {\em PCT, Spin, Statistics, and all that}, (Benjamin Inc, New York 1964).\end{thebibliography}\end{document}gra