\documentclass[onecolumn,amsmath,amssymb,nofootinbib,12pt]{article}
\usepackage{../helpers/jheppub}
\usepackage{ifpdf}

% \usepackage[bb=boondox]{mathalfa}

%\usepackage[notcite,notref]{showkeys}

%\documentclass[11pt,a4paper,oneside]{article}
%[preprint]{revtex4}
%Changed subfigure to subcaption
\usepackage{graphicx,subcaption}
\graphicspath{ {figures/} }
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{tensor}
\usepackage{epsfig}
% \usepackage{BOONDOX-cal}
%\usepackage[usenames]{color}
\usepackage{pdfpages}
\usepackage{bbm}
\usepackage{graphicx,epstopdf}
\usepackage[numbers]{natbib}
\usepackage[makeroom]{cancel}
\usepackage{hyperref}
\usepackage{tikzsymbols}
\hypersetup{pdftex,colorlinks=true,allcolors=blue}
%\usepackage[margin=1.4in]{geometry}
\usepackage{array}
\usepackage[export]{adjustbox}

\usepackage[normalem]{ulem}

\usepackage[papersize={7.0in, 10.0in}, left=.5in, right=.5in, top=1in, bottom=.9in]{geometry}
\linespread{1.05}
\sloppy
\raggedbottom
\pagestyle{plain}

%####################################
% VINCENT'S MACROS
\newcommand{\Rbb}{\mathbb{R}}
\newcommand{\Lcal}{\mathcal{L}}
\newcommand{\Zbb}{\mathbb{Z}}
\newcommand{\Ocal}{\mathcal{O}}

\newcommand{\Brane}{{\rm brane}}
\newcommand{\brane}{\mt{brane}}
\newcommand{\bulk}{\mt{bulk}}
\newcommand{\ren}{\mt{ren}}
\newcommand{\area}{A}
\newcommand{\gen}{\mt{gen}}
\newcommand{\eff}{\mt{eff}}
\newcommand{\dive}{\mt{div}}
\newcommand{\CFT}{\mt{CFT}}
\newcommand{\hyperF}{\tensor[_2]{F}{_1}}
\newcommand{\CFTR}{\mathbf{R}}
\newcommand{\braneR}{R}
\newcommand{\AdS}{\mathrm{AdS}}
\newcommand{\bulkbndry}{\sigma}
\newcommand{\bulkreg}{\Sigma}
\newcommand{\RTsurf}{\Sigma_\xR} %{\bulkbndry_\CFTR}
\newcommand{\RTbulkreg}{\bulkreg_\CFTR}
\newcommand{\AdSdmetric}{{g}^{\AdS_d}} %{\bar{g}^{\AdS_d}}
\newcommand{\RTmetric}{h}
\newcommand{\QESmetricNoz}{\mathfrak{h}}
\newcommand{\twoDmetric}{f}
\newcommand{\curvK}{\mathcal{K}}
\newcommand{\inducedK}{\tilde{\curvK}}
\newcommand{\inducedR}{\tilde{R}}
\newcommand{\inducedg}{\tilde{g}}
\newcommand{\inducednabla}{\tilde{\nabla}}
\newcommand{\inducedBox}{\tilde{\Box}}
\newcommand{\UV}{\mathrm{UV}}
\newcommand{\simUnder}[1]{\underset{#1}{\sim}}
\newcommand{\simUV}{\simUnder{\mathrm{UV}}}
\newcommand{\extr}{{\rm ext}}
\newcommand{\islands}{\mathrm{islands}}
\newcommand{\RT}{\mt{RT}}
\newcommand{\DGP}{\mt{DGP}}
\newcommand{\IR}{\mt{IR}}
\newcommand{\cT}{c_\mt{T}}
\newcommand{\overscript}[2]{\overset{\scriptscriptstyle{(#2)}}{#1}{}}
%\DeclareMathOperator*{\extr}{ext}
%####################################

\newcommand{\xR}{\mathbf{R}}
\newcommand{\xV}{\mathbf{V}}
\newcommand{\EE}{\mt{EE}}
\newcommand{\tilh}{{\tilde h}}


\newcommand{\RN}[1]{%
	\textup{\uppercase\expandafter{\romannumeral#1}}%
}
\newcommand{\labell}[1]{\label{#1}} %{\label{#1}} %
\newcommand{\comment}[1]{\textcolor{red}{\bf [[[#1]]]}}

\newcommand{\del}{\partial}
\newcommand{\vac}{\text{vac}}
\newcommand{\surf}{\text{surf}}
\newcommand{\BH}{\text{BH}}
%\newcommand{\bulk}{\text{bulk}}
\newcommand{\tot}{\text{tot}}
\newcommand{\corner}{\text{corner}}
\newcommand{\mx}{\text{max}}
\newcommand{\past}{\text{past}}
\newcommand{\future}{\text{future}}
\newcommand{\jnt}{\text{jnt}}
\newcommand{\phys}{\rm phys}
%\usepackage{musixtex}
%\usepackage{lilyglyphs}

\definecolor{limegreen}{rgb}{0.2, 0.8, 0.2}
\definecolor{deepcarrotorange}{rgb}{0.91, 0.41, 0.17}
\definecolor{darkviolet}{rgb}{0.58, 0.0, 0.83}
\definecolor{cyan}{rgb}{0.0, 0.72, 0.92}
\definecolor{plum}{rgb}{0.67, 0.0, 0.55}

\newcommand{\schmm}[1]{\textcolor{red}{\bf #1}}
\newcommand{\rcm}[1]{\textcolor{red}{\bf [[Rob: #1]]}}
\newcommand{\josh}[1]{\textcolor{plum}{\bf [[Joshua: #1]]}}
\newcommand{\iar}[1]{\textcolor{blue}{\bf [[IR: #1]]}}
\newcommand{\vc}[1]{\textcolor{magenta}{\bf [[Vincent: #1]]}}
\newcommand{\dn}[1]{\textcolor{limegreen}{\bf [[Dominik: #1]]}}

\newcommand{\hd}[1]{\noindent{\bf #1}\ }

\newcommand{\eg}{{\it e.g.,}\ }
\newcommand{\ie}{{\it i.e.,}\ }
\newcommand{\reef}[1]{(\ref{#1})}
\newcommand{\ssc}{\scriptscriptstyle}
\newcommand{\mt}[1]{\textrm{\tiny #1}}

%\newcommand{\s}{\sigma}
\newcommand{\s}{z_\mt{B}}
\newcommand{\sir}{z_\mt{IR}}
\newcommand{\lir}{\ell_\mt{IR}}
\newcommand{\lb}{\ell_\mt{B}}
\newcommand{\lamb}{\lambda_{b}}
\newcommand{\lgb}{\lambda_\mt{GB}}
\newcommand{\thb}{\theta_\mt{CFT}}
\newcommand{\zeb}{\zeta_\mt{CFT}}
%\newcommand{\thb}{\theta_{\partial\CFTR}}
%\newcommand{\zeb}{\zeta_{\partial\CFTR}}
\newcommand{\SCFT}{\Sigma_\mt{CFT}}
%\newcommand{\SCFT}{\partial\CFTR}
\newcommand{\puv}{P_\mt{UV}}
\newcommand{\pb}{P_\mt{B}}
\newcommand{\pbo}{P_\mt{B,0}}
\newcommand{\veps}{\varepsilon}
\newcommand{\sgen}{S_\mt{gen}}
\newcommand{\Agen}{\mathcal{A}_\mt{gen}}


\newcommand{\CC}{\mathcal{C}}
\newcommand{\mS}{\mathcal{S}}
\newcommand{\Gn}{G_\mt{N}}
\newcommand{\eps}{\epsilon}
\newcommand{\tk}{{\tilde k}}
\newcommand{\mA}{\mathcal{A}}
\newcommand{\mB}{\mathcal{B}}
\newcommand{\mD}{\mathcal{D}}
\newcommand{\mM}{\mathcal{M}}
\newcommand{\mK}{{\cal K}} %{\mathcal{K}}
\newcommand{\mR}{{\cal R}} %{\mathcal{R}}
\newcommand{\ric}{{\tilde R}}

\newcommand{\ads}[1]{{\mt{AdS}_{\tiny #1}}}
\newcommand{\tg}{{\tilde g}}
\newcommand{\tR}{{\tilde R}}
\newcommand{\tdr}{{\tilde r}}


\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\beqa}{\begin{eqnarray}}
\newcommand{\eeqa}{\end{eqnarray}}

\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}
\newcommand{\nn}{\nonumber}
\newcommand{\pa}{\partial}
\newcommand{\vk}{{\vec k}}

\newcommand{\pen}{\frak{a}}
\newcommand{\tx}{\tilde{x}}
\newcommand{\tp}{\tilde{p}}
\newcommand{\hx}{\hat{x}}
\newcommand{\hp}{\hat{p}}
\newcommand{\hX}{{\widehat X}}
\newcommand{\hV}{{\widehat V}}
\newcommand{\hW}{{\widehat W}}
\newcommand{\hZ}{{\widehat Z}}

\newcommand{\lp}{\left(}
\newcommand{\rp}{\right)}
\newcommand{\op}{\mathcal{O}}
\newcommand{\cev}[1]{\reflectbox{\ensuremath{\vec{ \reflectbox{\ensuremath{#1}}}}}}
\newcommand{\tr}{{\rm tr}}
\newcommand{\bx}{\mathbf{x}}

%Some useful commands for QM
\newcommand{\bra}[1]{\left< #1 \right|}
\newcommand{\ket}[1]{\left| #1 \right>}
\newcommand{\expVal}[1]{\left< #1 \right>}
\newcommand{\braket}[2]{\left<#1|#2\right>}



\renewcommand{\(}{\left(}
\renewcommand{\)}{\right)}
\renewcommand{\[}{\left[}
\renewcommand{\]}{\right]}
\newcommand{\dslash}{\delta^{\!\!\!\!-}\!}
\def\Tr{{\text{Tr}}}

\newcommand{\psT}{\psi_\mt{T}}
\newcommand{\psR}{\psi_\mt{R}}
\newcommand{\wrr}{\omega_{\mt R}}
\newcommand{\id}{\mathbbm{1}}
\newcommand{\zero}{\mathbbm{0}}
\newcommand{\mC}{\mathcal{C}}
\newcommand{\mO}{\mathcal{O}}
\newcommand{\bR}{\mathbb{R}}
\newcommand{\mH}{\mathbb{H}}
\newcommand{\mV}{\mathcal{V}}
\newcommand{\mL}{\mathcal{L}}
\newcommand{\vq}{{\vec q}}
\newcommand{\vqp}{{\vec q}^{\,\, \prime}}
\newcommand{\Gbr}{G_\mt{brane}}
\newcommand{\Gbk}{G_\mt{bulk}}
\newcommand{\Geff}{G_\mt{eff}}
\newcommand{\leff}{\ell_\mt{eff}}

\newcommand{\RNum}[1]{\uppercase\expandafter{\romannumeral #1\relax}}


\usepackage[thinlines]{easytable}
%\usepackage{tikzsymbols}
\usepackage{textcomp}
%\usepackage{parskip}

%\newcommand{\max}{\text{max}}

\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{2}

\title{\boldmath Quantum Extremal Islands Made Easy, Part I:\\
Entanglement on the Brane}
%\\ Ryu-Takayanagi meets Randall-Sundrum}
%\title{\boldmath Life on the Planck Brane}


\author[a,b]{Hong Zhe Chen,}
\author[a]{Robert C. Myers,}
\author[a]{Dominik Neuenfeld,}
\author[c]{Ignacio A. Reyes}
\author[a,b]{and Joshua Sandor}

\affiliation[a]{Perimeter Institute for Theoretical Physics, Waterloo, ON N2L 2Y5, Canada}
\affiliation[b]{Dept.~of Physics $\&$ Astronomy, University of Waterloo, Waterloo, ON N2L 3G1, Canada}
\affiliation[c]{Max-Planck-Institut f\"ur Gravitationsphysik, Am M\"uhlenberg 1, 14476 Potsdam, Germany}

\emailAdd{hchen2@pitp.ca}
\emailAdd{rmyers@pitp.ca}
\emailAdd{dneuenfeld@pitp.ca}
\emailAdd{ignacio.reyes@aei.mpg.de}
\emailAdd{jsandor@perimeterinstitute.ca}




\abstract{Recent progress in our understanding of the black hole information paradox has lead to a new prescription for calculating entanglement entropies, which involves special subsystems in regions where gravity is dynamical, called \textit{quantum extremal islands}. We present a simple holographic framework where the emergence of quantum extremal islands can be understood in terms of the standard Ryu-Takayanagi prescription, used for calculating entanglement entropies in the boundary theory. Our setup describes a $d$-dimensional boundary CFT coupled to a ($d$--1)-dimensional defect, which are dual to global AdS${}_{d+1}$ containing a codimension-one brane. Through the Randall-Sundrum mechanism, graviton modes become localized at the brane, and in a certain parameter regime, an effective description of the brane is given by Einstein gravity on an AdS${}_d$ background coupled to two copies of the boundary CFT. Within this effective description, the standard RT formula implies the existence of quantum extremal islands in the gravitating region, whenever the RT surface crosses the brane. This indicates that islands are a universal feature of effective theories of gravity and need not be tied to the presence of black holes.}


\begin{document}
\maketitle

%\setcounter{section}{-1}
%\section{Proposed Outline}
%\vskip0.3cm


%\rcm{Use the notation of new note!! \eg $G_\mt{brane},\ G_\mt{bulk},\ G_\mt{eff},\ \ell_\mt{eff}$}\\

%\rcm{Use notation like `eq.' or `eqs.' in front of equation numbers. Use `figure' rather than `fig.' or `Figure'. Get bibtex references from INSPIRE.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage
\section{Introduction}\label{sec:introduction}
\input{sections/introduction}

\section{Brane Gravity}\label{sec:branegravity}
\input{sections/GravityBrane}
\input{sections/twodd}
\input{sections/DGP}

\section{Three perspectives: Bulk/Brane/Boundary}\label{face}
\input{sections/holoface_dom}

\section{Holographic EE on the Brane}\label{HEE}
\input{sections/HEE1}
\input{sections/HEE2}


\section{Discussion}\label{sec:discussion}
\input{sections/discussion}


\section*{Acknowledgments}
We would like to thank Ahmed Almheiri, Raphael Bousso, Xi Dong, Netta Engelhardt,  Zach Fisher, Greg Gabadadze, Juan Hernandez, Don Marolf, Shan-Ming Ruan, Edgar Shaghoulian, Antony Speranza and Raman Sundrum for useful comments and discussions. Research at Perimeter Institute is supported in part by the Government of Canada through the Department of Innovation, Science and Economic Development Canada and by the Province of Ontario through the Ministry of Colleges and Universities. RCM is supported in part by a Discovery Grant from the Natural Sciences and Engineering Research Council of Canada, and by the BMO Financial Group. HZC is supported by the Province of Ontario and the University of Waterloo through an Ontario Graduate Scholarship. RCM and DN also received funding from the Simons Foundation through the ``It from Qubit'' collaboration. The work of IR is funded by the Gravity, Quantum Fields and Information group at AEI, which is generously supported by the Alexander von Humboldt Foundation and the Federal Ministry for Education and Research  through the Sofja Kovalevskaja Award. IR also acknowledges the support of the Perimeter Visiting Graduate Fellows program and the hospitality of Perimeter Institute, where part of this work was done. JS acknowledges the support of the Natural Sciences and Engineering Research Council of Canada (NSERC).

\appendix

\section{Generalized Entropy on the Brane}\label{generalE}
\input{sections/general}

\section{RT Bubbles}
\label{bubble}
\input{sections/bubbles}

%\section{Thinking about Ensembles?}
%\label{ensign}
%\input{sections/ensemble-app}


%\section{Details of AdS/ICFT} \label{details}
%\input{sections/details}


\bibliography{bibliography}
\bibliographystyle{utphys}

\end{document}


%%% JUNK %%%%%

\dn{Extras for version 2}
Extras:
\begin{enumerate}
\item \rcm{intersection of brane and boundary is not a manifold??} 
\item \dn{Flat limit?}
\item \dn{de Sitter?}
\item \dn{Ensemble averaging!}
%\item {\it What to do with:} \rcm{Drop} While we specify the position and shape of the entangling surface in the asymptotic boundary, the position and shape of the RT surface as it crosses the brane, \ie the quantum extremal surface, is determined dynamically. There is a tension with \cite{Myers:2013lva}.
\end{enumerate}


\subsection{Old stuff}
\dn{Remove this subsection if not needed anymore}

\rcm{Tell the story of three descriptions of the same system; make the parallel with the three descriptions in \cite{Almheiri:2019hni}; old words:} There are two sides to the brane story that we will consider in this section. In the first perspective (section \ref{bulkgravity}) we view the brane as properly lying in the ($d+1$)-dimensional spacetime. We can use the Israel junction conditions to relate the bulk stress-energy tensor across the brane to the discontinuity of the extrinsic curvature across the brane, which determines the location of the brane as embedded in higher dimensional space. The second perspective (section \ref{inducedgravity}) applies the Randall-Sundrum technique to realize an effective action on the brane in which we are left only with a $d$-dimensional theory. We see the consistency between the pictures by observing how the equations of motion fix the brane position in the ambient spacetime.


\rcm{mention:} Our Planck brane is in the middle of the space and so there is no confusion about whether or not these degrees of freedom belong to the boundary or the bulk. It's the bulk!

\rcm{mention:} brane theory as an `ensemble', \ie not a CFT with a cutoff but an ensemble of CFTs \dn{So far I still disagree. I don't think that there are Randall-Sundrum modes for sources of scalar fields. The sources in the RS case correspond to the value of the Neumann BC of the non-gravitational fields towards the brane, which (at least in the $\mathbb Z_2$ orbifold) needs to be fixed.}

\rcm{Move the discussion of cutoff from section \ref{wormy} here:} Due to the large number of matter degrees of freedom, the formation of bubbles can occur above the Planck scale $A\sim G_\eff$. However, for the same reason, it becomes questionable whether the semiclassical description of gravity does in fact hold down to the Planck scale. Roughly speaking, $c$-many matter fields should backreact $c$-times as strongly: backreactions which are ordinarily suppressed by $G_\eff$ are now only suppressed by $c G_\eff$. In terms of Feynman diagrams, while loop order is ordinarily suppressed by $G_\eff$, matter loops now carry factors of $c$. This seems to suggest that with a large $c$ number of matter degrees of freedom, gravity can only be treated semiclassically down to scales of order\footnote{For related discussions, see \cite{Dvali:2007hz,Dvali:2007wp,Reeb:2009rm}. \vc{``Rob should take a look at them in more detail -- in particular, \cite{Dvali:2007wp}."}} $c G_\eff$ rather than $G_\eff$. Indeed, the exchange in dominance between `classical' area term $A/4G_\eff$ and the quantum entropy $S_\ren$ at scales $A\sim c G_\eff$ appears to support this claim. However, as argued in the previous paragraph, the formation of bubbles is not possible at scales $A\gtrsim c G_\eff$. In fact, the bubble size found in \eqref{eq:cookie} lies well below this scale. Hence, it would seem that the formation of bubbles is a mere by-product of attempting to probe length scales below the range of validity of semiclassical gravity on the brane.




%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
