\documentclass[11pt]{article}
\usepackage{jheppubmod}
\usepackage[showonlyrefs]{mathtools}
\usepackage{psfrag}
\usepackage{array}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{caption}
\usepackage[labelsep=quad]{subcaption}
\usepackage{epstopdf}
\let\normalcolor\relax	
\usepackage{epsfig}
\usepackage[punctsep]{collref}
\collectsep[]{;}	
\def\half{{1 \over 2}}	
\def\cn{{\mathcal N}}
\def\tr{{\rm Tr}}
\def\l{\lambda}
\def\la{\lambda}
\def\lad{\lambda^{\dagger}}
\def\lbd{\bar{\lambda}^{\dagger}}
\def\sl{\slashed}
\def\lb{\bar{\lambda}}
\def\m{\mu}
\def\mb{\bar{\mu}}
\def\wb{\bar{w}}
\def\wa{\widetilde{a}}
\def\e{\eta}
\def\ep{\epsilon}
\def\lvac{\langle 0 |}
\def\rvac{|0 \rangle}
\def\a{\alpha}
\def\ad{\dot{\alpha}}
\def\ab{\bar{a}}
\def\eb{\tilde{\eta}}
\def\Qb{\bar{Q}}
\def\Or[#1]{{\text{O}}\left({#1}\right)}
\def\dotl[#1,#2]{\left\langle #1,\, #2 \right\rangle}
\def\dotlb[#1,#2]{\left\langle #1,\, #2 \right\rangle}
\def\dotlm[#1,#2]{\left[ #1,\, #2 \right]}
\def\dotp[#1,#2]{(\vect{#1} \cdot\vect{#2})}
\def\aff[#1,#2]{\hat{#1}(#2)}
\def\n4sym{{\cal N}=4 SYM}
\def\>{\rangle}
\def\<{\langle}
\def\weight[#1,#2,#3]{\{(#1),#2,#3\}}
\def\ads[#1]{$\text{AdS}_{#1}$}
\def\cft[#1]{$\text{CFT}_{#1}$}
\def\front{\rm I}
\def\black{\rm II}
\def\other{\rm III}
\def\white{\rm IV}
\def\lrdel{\overset{\leftrightarrow}{\nabla}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\ba}{\begin{align}}
\newcommand{\ea}{\end{align}}
\newcommand{\bs}{\begin{split}}
\def\sess\end{split}
\def\zhor{z_0}
\newcommand{\hatbb}[1]{\hat{#1}}
\newcommand{\vect}[1]{{\boldsymbol{#1}}}
\newcommand{\norm}[1]{|{\boldsymbol{#1}}|}
\newcommand{\sr}{\color{red}}
\newcommand{\state}[1]{{|#1\rangle}}
\newcommand{\zconstant}{Z^2}
\newcommand{\normfact}{\Upsilon}
\title{An Infalling Observer in AdS/CFT}
\author[a,b]{Kyriakos Papadodimas}
\author[c,d,e]{and Suvrat Raju}
\affiliation[a]{Centre for Theoretical Physics, University of Groningen, Nijenborgh 4, The Netherlands.}
\affiliation[b]{Theory Group, Physics Department, CERN, CH-1211 Geneva 23,
Switzerland.}
\affiliation[c]{International Centre for Theoretical Sciences, IISc Campus, Bengaluru 560012, India.}
\affiliation[d]{Harish-Chandra Research Institute, Chatnag Marg, Jhunsi,
Allahabad 211019, India.}
\affiliation[e]{School of Natural Sciences, Institute for Advanced Study,
Princeton, NJ 08540, USA.}
\emailAdd{k.papadodimas@rug.nl}
\emailAdd{suvrat@icts.res.in}



\date{}





\abstract{We describe the experience of an observer falling into a
  black hole using the AdS/CFT correspondence.  In order to do this,  we
  reconstruct the local bulk operators measured by the observer along
  his trajectory outside the black hole. We then extend our
  construction beyond the black hole horizon. We show that this is
 possible because of an effective doubling of the observables in the
  boundary theory, when it is in a pure state that is close to the
  thermal state. Our construction allows us to rephrase questions
  about information-loss
and the structure of the metric at the horizon in terms of more
familiar CFT correlators. It suggests that to precisely
identify black-hole microstates, the observer would need to conduct
measurements to an accuracy of $e^{-S_{\text{BH}}}$. This appears to be
inconsistent with the ``fuzzball'' proposal, and other recent proposals
in which pure states in the ensemble of the black hole are represented
by macroscopically distinct geometries. Furthermore, our description of the black hole interior
in terms of CFT operators provides a natural realization of
black hole complementarity and a method of preserving
unitarity without ``firewalls.''
}
\setcounter{tocdepth}{2}
\preprint{\parbox{3cm}{HRI/ST/1107 \\  ICTS/2012/10}}
\keywords{AdS-CFT, Information Paradox, Black Holes}
\listfiles
\begin{document}
\maketitle
\input{s1_introduction.tex}
\input{s3_emptyads.tex}
\input{gr_background.tex}
\input{s4_blackholefront.tex}
\input{s5_blackholebehind.tex}
\input{s6_applications.tex}
\input{subtleties.tex}
\input{s7_conclusions.tex}



\section{Acknowledgments}
We would like to thank  Ofer Aharony, Luis Alvarez-Gaume, Nima
Arkani-Hamed, Jan de Boer, Borun Chowdhury, Sheer El-Showk, Ben
Freivogel, Rajesh Gopakumar, Monica Guica, Dan Harlow, Gary Horowitz,
Veronika Hubeny, Norihiro Iizuka, Tom Hartman, Elias Kiritsis, Neil
Lambert, Klaus Larjo, Gilad Lifschytz, Loganayagam,  David Lowe, Juan Maldacena, Don
Marolf, Paul McFadden, Shiraz Minwalla, Boris Pioline, Joe Polchinski,
Mukund Rangamani, Ashoke Sen, Masaki Shigemori, Joan Simon, Sandip
Trivedi, Erik Verlinde, Spenta Wadia, Edward Witten and Xi Yin for helpful discussions. We would
also like to thank all the participants of the discussion meeting on
string theory at ICTS, Bangalore, in June 2012 and members of the quantum information group at HRI. KP would like to thank the organizers and participants
of the ``Bits, Branes and Black Holes'' workshop at KITP 2012, the Cosmology and Complexity 2012 meeting in Hydra, 
the Amsterdam summer string workshop 2012, the ``Holographic Way'' workshop in
Nordita and the Crete Center for Theoretical Physics for hospitality. SR is partially supported by a
Ramanujan fellowship of the Department of Science and Technology
(India).  SR gratefully acknowledges the hospitality of Brown
University, where part of this work was done. 


\appendix

\input{appendix_2dthermal.tex}
\input{rindler_quantization.tex}
\bibliographystyle{JHEP}
\bibliography{references}
\end{document}
