
\chapter{Acknowledgements}

The ideas presented here have been at the centre of my research for most of
the last six years, during which time I have been influenced---for the
better, I think---by more people than I can thank formally here.  Above all
I feel moulded by the people who were around me during my time in
Cambridge.  I want to thank Martin Hyland in particular: his ways of
thought have been a continual inspiration, and I feel very lucky to have
been his student and to have shared a department with him for so long.  I
have also gained greatly from conversations with Eugenia Cheng, Peter
Johnstone, and Craig Snydal.  Ivan Smith and Dick Thomas have cheerfully
acted as consultants for daft geometry questions.

Ian Grojnowski suggested this Lecture Notes series to me, and I am very
glad he did: I could not have wished for a more open-minded, patient and
helpful editor than Jonathan Walthoe at CUP.  I would also like to thank
Nigel Hitchin, the series editor.

I am very grateful to Andrea Hesketh for sound strategic advice.

Some of the quotations starting the chapters were supplied, directly or
indirectly, by Sean Carmody, David Corfield, and Colin Davey.  Paul-Andr\'e
Melli\`es gave me important information on Swiss cheese.

Almost all of the software used in the preparation of this book was free,
not just financially, but also in the libertarian sense: freedom to take it
apart, alter it, and propagate it, like a piece of mathematics.  I am
grateful to the thousands of developers who brought about the truly
remarkable situation that made this possible.  I also acknowledge the use
of Paul Taylor's excellent commutative diagrams package.

I started writing this when I was Laurence Goddard Research Fellow at St
John's College, Cambridge, and finished when I was William Hodge Fellow at
the Institut des Hautes \'Etudes Scientifiques.  The index was compiled
while I was visiting the University of Chicago at the invitation of Peter
May.  I am immensely grateful to St John's and the IH\'ES for the
opportunities they have given me, and for their unwavering dedication to
research.  In particular, I thank the IH\'ES for giving me the chance to
live in Paris and absorb some Russian culture.

