
\begin{thebibliography}{c9999}

\addcontentsline{toc}{chapter}{Bibliography}

\small

\bibitem[1994]{AR}
Ji\v{r}\'\i\ Ad\'amek, Ji\v{r}\'\i\ Rosick\'y,
\emph{Locally Presentable and Accessible Categories},
London Mathematical Society Lecture Note Series~189,
Cambridge University Press, 1994.

\bibitem[1978]{Ad}
J.F.~Adams, 
\emph{Infinite Loop Spaces},
Princeton University Press, 1978.

\bibitem[1971]{And}
D.W.~Anderson,
Spectra and $\Gamma$-sets,
in \emph{Algebraic Topology (Proceedings of Symposia in Pure Mathematics
22)}, AMS, 1971.  

\bibitem[1982]{Laurie}
Laurie Anderson, 
Let $X=X$, 
on \emph{Big Science}, Warner Brothers, 1982.

\bibitem[1980]{Athenian}
\emph{The Athenian}, That wretched kilt-wearer: the paintings of Theofilos,
May 1980.

\bibitem[1997]{BzIN}
John C.~Baez,
An introduction to $n$-categories,
e-print \url{q-alg\slsh 9705009}, 1997;
also in 
\emph{Category Theory and Computer Science (Santa Margherita Ligure,
1997)},
Lecture Notes in Computer Science~1290,
Springer, 1997.

\bibitem[1995]{BDHDA0}
John C.~Baez, James Dolan,
Higher-dimensional algebra and topological quantum field theory,
\emph{Journal of Mathematical Physics} 36 (11) (1995), 6073--6105;
also e-print \url{q-alg\slsh 9503002}, 1995 (without diagrams).

\bibitem[1997]{BDHDA3}
John C.~Baez, James Dolan,
Higher-dimensional algebra III: $n$-categories and the algebra of opetopes,
e-print \url{q-alg\slsh 9702014}, 1997;
also 
\emph{Advances in Mathematics} 135 (1998), no.~2, 145--206.

\bibitem[1998a]{BatMGC}
M.A.~Batanin,  
Monoidal globular categories as a natural environment for the theory of weak
$n$-categories, 
\emph{Advances in Mathematics} 136 (1998), no.~1, 39--103.

\bibitem[1998b]{BatCFM}
M.A.~Batanin,
Computads for finitary monads on globular sets,
in
\emph{Higher Category Theory (Evanston, IL, 1997)},
Contemporary Mathematics~230, 
AMS, 1998.

\bibitem[2002a]{BatPMW}
M.A.~Batanin,  
On the Penon method of weakening algebraic structures, 
\emph{Journal of Pure and Applied Algebra} 172 (2002), no.~1, 1--23.

\bibitem[2002b]{BatCSO}
M.A.~Batanin,
Computads and slices of operads,
e-print \url{math\dt CT\slsh 0209035}, 2002.

\bibitem[2000]{BSUPM}
Michael Batanin, Ross Street,
The universal property of the multitude of
trees, 
\emph{Journal of Pure and Applied Algebra} 154 (2000), no.~1-3, 3--13.

\bibitem[1997]{BJT}
Hans-Joachim Baues, Mamuka Jibladze, Andy Tonks,
Cohomology of monoids in monoidal categories,
in J-L.~Loday, J.~Stasheff, A.~Voronov (eds.), \emph{Operads: Proceedings
of Renaissance Conferences}, Contemporary Mathematics 202, AMS, 1997.

\bibitem[1969]{Beck}
Jon Beck,
Distributive laws,
in 
\emph{Seminar on Triples and Categorical Homology Theory (ETH, Z\"urich,
1966/67)},  
Springer, 1969.

\bibitem[\latin{c.}1997]{BeDr}
Alexander Beilinson, Vladimir Drinfeld, Chiral algebras, preprint,
\latin{c.}1997. 

\bibitem[1967]{Ben}
Jean B\'enabou, 
Introduction to bicategories,
in
J.~B\'enabou et al (eds.), \emph{Reports of the Midwest Category Seminar},
Lecture Notes in Mathematics~47, Springer, 1967.

\bibitem[2002]{Ber}
Clemens Berger,
A cellular nerve for higher categories,
\emph{Advances in Mathematics} 169 (2002), no.~1, 118--175.

\bibitem[1983]{BCSW}
Renato Betti, Aurelio Carboni, Ross Street, Robert Walters,
Variation through enrichment,
\emph{Journal of Pure and Applied Algebra} 29 (1983), 109--127.

\bibitem[1989]{BKP}
R.~Blackwell, G.M.~Kelly, A.J.~Power,
Two-dimensional monad theory,
\emph{Journal of Pure and Applied Algebra} 59 (1989), no.~1, 1--41.

\bibitem[1973]{BV}
J.M.~Boardman, R.M.~Vogt, 
\emph{Homotopy Invariant Algebraic Structures on
Topological Spaces},
Lecture Notes in Mathematics~347,
Springer, 1973.

\bibitem[1994a]{Borx1}
Francis Borceux, 
\emph{Handbook of Categorical Algebra 1: Basic Category Theory},
Encyclopedia of Mathematics and its Applications~50,
Cambridge University Press, 1994.

\bibitem[1994b]{Borx2}
Francis Borceux, 
\emph{Handbook of Categorical Algebra 2: Categories and
Structures},
Encyclopedia of Mathematics and its Applications~51,
Cambridge University Press, 1994.

\bibitem[1997]{Borch}
Richard Borcherds, 
Vertex algebras, 
e-print \url{q-alg\slsh 9706008}, 1997;
also in 
\emph{Topological Field Theory, Primitive Forms and Related Topics
(Kyoto, 1996)},
Progress in Mathematics~160,
Birkh\"auser Boston, 1998.

\bibitem[1999]{Bri}
Susie Bright,
\emph{Full Exposure},
HarperCollins, 1999.

\bibitem[1999]{BrownJ}
James Robert Brown,
\emph{Philosophy of Mathematics: an Introduction to the World of Proofs and
Pictures},
Routledge, 1999. 

\bibitem[1971]{Bur}
Albert Burroni,  
$T$-cat\'egories (cat\'egories dans un triple),
\emph{Cahiers de Topologie et G\'eom\'etrie Diff\'erentielle} 
12 (1971), 215--321.

\bibitem[1991]{BurHDW}
Albert Burroni,
Higher-dimensional word problem,
in \emph{Category Theory and Computer Science (Paris, 1991)},
Lecture Notes in Computer Science~530,
Springer, 1991.
 
\bibitem[1993]{BurHDWA}
Albert Burroni,
Higher-dimensional word problems with applications to equational logic,
\emph{Theoretical Computer Science} 115 (1993), no.~1, 43--62.

\bibitem[1995]{CJ}
A.~Carboni, P.~Johnstone,
Connected limits, familial representability and Artin glueing,
\emph{Mathematical Structures in Computer Science} 5 (1995), 441--459.

\bibitem[2002]{CJcorr}
A.~Carboni, P.~Johnstone, 
Corrigenda to `Connected limits, familial
representability and Artin glueing', to appear.

\bibitem[1987]{CKW}
Aurelio Carboni, Stefano Kasangian, Robert Walters,
An axiomatics for bicategories of modules,
\emph{Journal of Pure and Applied Algebra} 45 (1987), 127--141.

\bibitem[2003a]{CheWOM}
Eugenia Cheng,
Weak $n$-categories: opetopic and multitopic foundations,
also e-print \url{math\dt CT\slsh 0304277}, 2003;
\emph{Journal of Pure and Applied Algebra}, to appear.

\bibitem[2003b]{CheWCO}
Eugenia Cheng,
Weak $n$-categories: comparing opetopic foundations,
also e-print \url{math\dt CT\slsh 0304279}, 2003;
\emph{Journal of Pure and Applied Algebra}, to appear.

\bibitem[2003c]{CheCOCOS}
Eugenia Cheng,
The category of opetopes and the category of opetopic sets,
e-print \url{math\dt CT\slsh 0304284}, 2003.

\bibitem[2003d]{CheOBC}
Eugenia Cheng,
Opetopic bicategories: comparison with the classical theory,
e-print \url{math\dt CT\slsh 0304285}, 2003.

\bibitem[2003e]{CheACU}
Eugenia Cheng,
An alternative characterisation of universal cells in opetopic
$n$-categories,
e-print \url{math\dt CT\slsh 0304286}, 2003.

\bibitem[1999]{CraTPG}
Sjoerd E.~Crans,
A tensor product for \textbf{Gray}-categories,
\emph{Theory and Applications of Categories} 5 (1999), no.~2, 12--69.

\bibitem[2000]{CraBSS}
Sjoerd Crans, 
On braidings, syllepses and symmetries,
\emph{Cahiers de Topologie et G\'eom\'etrie Diff\'erentielle Cat\'egorique} 
41 (2000), no.~1, 2--74.

\bibitem[2001]{DusSM2}
John W.~Duskin, 
A simplicial-matrix approach to higher dimensional category theory II:
bicategory morphisms and simplicial maps, 
preprint, 2001.

\bibitem[2002]{DusSM1}
John W.~Duskin, 
Simplicial matrices and the nerves of weak $n$-categories I: nerves of
bicategories, 
\emph{Theory and Applications of Categories} 9 (2002), no.~10, 198--308.

\bibitem[1989]{DKS}
W.G.~Dwyer, D.M.~Kan, J.H.~Smith,
Homotopy commutative diagrams and their realizations,
\emph{Journal of Pure and Applied Algebra} 57 (1989), no.~1, 5--24.

\bibitem[1962]{EH}
B.~Eckmann, P.J.~Hilton, 
Group-like structures in general categories, I: 
multiplications and comultiplications,
\emph{Mathematische Annalen} 145 (1962), 227--255.

\bibitem[1995]{Eco}
Umberto Eco, 
\emph{The Search for the Perfect Language}, 
Blackwell, 1995.

\bibitem[1965]{Ehr}
Charles Ehresmann, 
\emph{Cat\'egories et Structures},
Dunod, 1965.

\bibitem[1967]{GZ}
P.~Gabriel, M.~Zisman,
\emph{Calculus of Fractions and Homotopy Theory},
Springer, 1967.

\bibitem[2002]{Gan}
Wee Liang Gan,
Koszul duality for dioperads,
e-print \url{math\dt QA\slsh 0201074}, 2002.

\bibitem[1994]{GeKaMO}
E.~Getzler, M.M.~Kapranov,
Modular operads,
e-print \url{dg-ga\slsh 9408003}, 1994; 
also
\emph{Compositio Mathematica} 110 (1998), no.~1, 65--126.

\bibitem[1995]{GeKaCOC}
E.~Getzler, M.M.~Kapranov,
Cyclic operads and cyclic homology,
in \emph{Geometry, Topology, \& Physics},
Conference Proceedings and Lecture Notes in Geometry and Topology~IV,
International Press, 1995.

\bibitem[1994]{GiKa}
Victor Ginzburg, Mikhail Kapranov, 
Koszul duality for operads,
\emph{Duke Mathematical Journal} 76 (1994), no.~1, 203--272.

\bibitem[1995]{GPS}
R.~Gordon, A.J.~Power, Ross Street, 
\emph{Coherence for Tricategories}, 
Memoirs of the AMS~117, no.~558, 1995.

\bibitem[1999]{GPLDC}
Marco Grandis, Robert Par\'e, 
Limits in double categories,
\emph{Cahiers de Topologie et G\'eom\'etrie Diff\'erentielle Cat\'egorique}
40 (1999), no.~3, 162--220.

\bibitem[2003]{GPADC}
Marco Grandis, Robert Par\'e, 
Adjoints for double categories, 
preprint, 2003.

\bibitem[1983]{GroPS}
A.~Grothendieck, 
Pursuing stacks,
manuscript, 1983.

\bibitem[2000]{HerRM}
Claudio Hermida,  
Representable multicategories,
\emph{Advances in Mathematics} 151 (2000), no.~2, 164--225.

\bibitem[1998]{HMPHDM}
Claudio Hermida, Michael Makkai, John Power,
Higher-dimensional multigraphs,
in
\emph{Thirteenth Annual IEEE Symposium on Logic in Computer Science
(Indianapolis, IN, 1998)},
IEEE Computer Society, 1998.

\bibitem[2000]{HMP1}
Claudio Hermida, Michael Makkai, John Power,
On weak higher dimensional categories I.1, 
\emph{Journal of Pure and Applied Algebra} 154 (2000), no.~1-3, 221--246.

\bibitem[2001]{HMP2}
Claudio Hermida, Michael Makkai, John Power,
On weak higher-dimensional categories I.2, 
\emph{Journal of Pure and Applied Algebra} 157 (2001), no.~2-3, 247--277.

\bibitem[2002]{HMP3}
Claudio Hermida, Michael Makkai, John Power,
On weak higher-dimensional categories I.3,
\emph{Journal of Pure and Applied Algebra} 166 (2002), no.~1-2, 83--104.

\bibitem[1981]{Hut}
John E. Hutchinson, 
Fractals and self-similarity,
\emph{Indiana University Mathematics Journal} 30 (1981), no.~5, 713--747.

\bibitem[1986]{JoyFAE}
Andr\'e Joyal,
Foncteurs analytiques et esp\`eces de structures,
in
\emph{Combinatoire \'Enum\'erative (Montreal, Quebec, 1985)},
Lecture Notes in Mathematics~1234,
Springer, 1986.

\bibitem[1997]{JoyDDT}
A.~Joyal,
Disks, duality and $\Theta$-categories,
preprint, 1997.

\bibitem[1993]{JSBTC}
Andr\'e Joyal, Ross Street,
Braided tensor categories, 
\emph{Advances in Mathematics} 102 (1993), no.~1, 20--78.

\bibitem[1994]{KV}
M.M.~Kapranov, V.A.~Voevodsky,
2-categories and Zamolodchikov tetrahedra equations,
in 
\emph{Algebraic Groups and their Generalizations: Quantum and
Infinite-Dimensional Methods}, Proceedings of Symposia in Pure Mathematics 56,
Part 2, AMS, 1994.

\bibitem[1972]{KelMVFI}
G.M.~Kelly, 
Many-variable functorial calculus I,
in
\emph{Coherence in Categories},
Lecture Notes in Mathematics~281,
Springer, 1972.

\bibitem[1974]{KelCD}
G.M.~Kelly,
On clubs and doctrines,
in \emph{Category Seminar}, Lecture Notes in Mathematics 420,
Springer, 1974.

\bibitem[1980]{KelUTT}
G.M.~Kelly,
A unified treatment of transfinite constructions for free algebras, free
monoids, colimits, associated sheaves, and so on,
\emph{Bulletin of the Australian Mathematical Society} 22 (1980), 1--83.

\bibitem[1982]{KelBCE}
Gregory Maxwell Kelly, 
\emph{Basic Concepts of Enriched Category Theory},
London Mathematical Society Lecture Note Series~64,
Cambridge University Press, 1982.

\bibitem[1992]{KelCDC}
G.M.~Kelly, 
On clubs and data-type constructors,
in 
\emph{Applications of Categories in Computer Science (Durham, 1991)},
London Mathematical Society Lecture Note Series~177,
Cambridge University Press, 1992.
 
\bibitem[1974]{KSRE2}
G.M.~Kelly, Ross Street, 
Review of the elements of $2$-categories, 
in
\emph{Category Seminar (Sydney, 1972\slsh 1973)},
Lecture Notes in Mathematics~420,
Springer, 1974.

\bibitem[2001]{KhoFVI}
Mikhail Khovanov,
A functor-valued invariant of tangles,
e-print \url{math\dt QA\slsh 0103190}, 2001;
also
\emph{Algebraic and Geometric Topology} 2 (2002), 665--741.

\bibitem[1999]{KonOMD}
Maxim Kontsevich, 
Operads and motives in deformation quantization,
e-print \url{math\dt QA\slsh 9904055}, 1999; 
also 
\emph{Letters in Mathematical Physics} 48 (1999), no.~1, 35--72.

\bibitem[1994]{KMGWC}
M.~Kontsevich, Yu.~Manin, 
Gromov-Witten classes, quantum cohomology and enumerative geometry,
e-print \url{hep-th\slsh 9402147}, 1994; 
also
\emph{Communications in Mathematical Physics} 164 (1994), no.~3, 525--562.

\bibitem[2003]{KosMAP}
J.~Koslowski,
A monadic approach to polycategories,
\emph{Electronic Notes in Theoretical Computer Science} 69 (2003).

\bibitem[1969]{LamDSCII}
Joachim Lambek, 
Deductive systems and categories II: standard constructions
and closed categories,
in
P.~Hilton (ed.), 
\emph{Category Theory, Homology Theory and their Applications, I
(Battelle Institute Conference, Seattle, 1968, Vol. One)}
Lecture Notes in Mathematics~86,
Springer, 1969.

\bibitem[1996]{Law}
R.J.~Lawrence,
An introduction to topological field theory,
in \emph{The Interface of Knots and Physics}, 
Proceedings of Symposia in Applied Mathematics~51 (1996), AMS, 89--128.

\bibitem[1955]{Laz}
Michel Lazard,
Lois de groupes et analyseurs,
\emph{Annales Scientifiques de l'\'Ecole Normale Sup\'erieure (3)}
72 (1955), 299--400.

\bibitem[1998a]{GOM}
Tom Leinster,
General operads and multicategories,
e-print \url{math\dt CT\slsh 9810053}, 1998.

\bibitem[1998b]{BB}
Tom Leinster,
Basic bicategories,
e-print \url{math\dt CT\slsh 9810017}, 1998.

\bibitem[1999a]{GECM}
Tom Leinster,
Generalized enrichment for categories and multicategories,
e-print \url{math\dt CT\slsh 9901139}, 1999.

\bibitem[1999b]{UTHM}
Tom Leinster,
Up-to-homotopy monoids,
e-print \url{math\dt QA\slsh 9912084}, 1999.

\bibitem[2000a]{HAO}
Tom Leinster,
Homotopy algebras for operads,
e-print \url{math\dt QA\slsh 0002180}, 2000.

\bibitem[2000b]{OHDCT}
Tom Leinster,
Operads in higher-dimensional category theory,
e-print \url{math\dt CT\slsh 0011106}, 2000.

\bibitem[2001a]{SHDCT}
Tom Leinster,
Structures in higher-dimensional category theory,
e-print \url{math\dt CT\slsh 0109021}, 2001.

\bibitem[2001b]{SDN}
Tom Leinster, 
A survey of definitions of $n$-category,
e-print \url{math\dt CT\slsh 0107188}, 2001;
also
\emph{Theory and Applications of Categories} 10 (2002), no.~1, 1--70.

\bibitem[2002]{GEC}
Tom Leinster, 
Generalized enrichment of categories,
e-print \url{math\dt CT\slsh 0204279}, 2002; 
also
\emph{Journal of Pure and Applied Algebra} 168 (2002), no.~2-3, 391--406.

\bibitem[2003]{FM}
Tom Leinster, 
Fibrations of multicategories,
to appear.

\bibitem[1972]{Lew} 
Geoffrey Lewis, 
Coherence for a closed functor, 
in
G.M.~Kelly, M.~Laplaza, G.~Lewis, S.~Mac Lane (eds.), 
\emph{Coherence in Categories}, Lecture Notes in Mathematics~281, 
Springer, 1972.

\bibitem[1997]{Ren}
Jean-Louis Loday, James D. Stasheff, Alexander A. Voronov (eds.), 
\emph{Operads: Proceedings of Renaissance Conferences}, AMS, 1997.

\bibitem[1963]{MacNAC}
Saunders Mac Lane, 
Natural associativity and commutativity,
\emph{Rice University Studies} 49 (1963), no.~4, 28--46.

\bibitem[1971]{MacCWM}
Saunders Mac Lane,
\emph{Categories for the Working Mathematician},
Graduate Texts in Mathematics~5, Springer, 1971, revised edition 1998.

\bibitem[1992]{MM}
Saunders Mac Lane, Ieke Moerdijk,
\emph{Sheaves in Geometry and Logic},
Springer, 1992.

\bibitem[1996]{MakAAC}
M.~Makkai,
Avoiding the axiom of choice in general category theory,
\emph{Journal of Pure and Applied Algebra} 108 (1996), no.~2, 109--173.

\bibitem[1999]{MakMOC}
M.~Makkai,
The multitopic $\omega$-category of all multitopic $\omega$-categories,
preprint, 1999.

\bibitem[2001]{MZ}
Mihaly Makkai, Marek Zawadowski, 
Duality for simple $\omega$-categories and disks,
\emph{Theory and Applications of Categories} 8 (2001), no.~7, 114--243.

\bibitem[1976]{Manes}
Ernest G. Manes, 
\emph{Algebraic Theories},
Graduate Texts in Mathematics~26,
Springer, 1976.

\bibitem[2002]{Marcus}
Ben Marcus,
\emph{Notable American Women},
Vintage, 2002.

\bibitem[2002]{MSS}
Martin Markl, Steve Shnider, Jim Stasheff, 
\emph{Operads in Algebra, Topology and Physics},
Mathematical Surveys and Monographs~96,
AMS, 2002.

\bibitem[1972]{MayGIL}
J.P.~May, 
\emph{The Geometry of Iterated Loop Spaces},
Lectures Notes in Mathematics~271, Springer, 1972.

\bibitem[1997]{MayDOA}
J.P.~May,
Definitions: operads, algebras and modules, 
in 
J-L.~Loday, J.~Stasheff, A.~Voronov (eds.), \emph{Operads: Proceedings
of Renaissance Conferences}, 
Contemporary Mathematics 202, AMS, 1997. 

\bibitem[2001]{MayOCA}
J.P.~May, 
Operadic categories, $A_\infty$-categories and $n$-categories,
preprint, 2001.

\bibitem[1989]{MoSe}
Gregory Moore, Nathan Seiberg, 
Classical and quantum conformal field theory,
\emph{Communications in Mathematical Physics} 123 (1989), no.~2, 177--254.

\bibitem[1999]{Pen}
Jacques Penon, 
Approche polygraphique des $\infty$-categories non strictes, 
\emph{Cahiers de Topologie et G\'eom\'etrie Diff\'erentielle Cat\'egorique}
40 (1999), no.~1, 31--80.

\bibitem[1954]{Rea}
Pauline R\'eage, 
\emph{Story of O},
1954.

\bibitem[1999]{Rue}
David Ruelle, 
Conversations on mathematics with a visitor from outer space,
in V.~Arnold, M.~Atiyah, P.~Lax, B.~Mazur (eds.), 
\emph{Mathematics: Frontiers and Perspectives}, International Mathematical
Union/AMS, 1999.

\bibitem[2000]{Sal}
Paolo Salvatore, 
The universal operad of iterated loop spaces,
talk at \emph{Workshop on Operads},
Osnabr\"uck, 2000.

\bibitem[1992]{SV}
R.~Schw\"anzl, R.~Vogt, 
Homotopy homomorphisms and the hammock localization,
\emph{Bolet\'\i n de la Sociedad Matem\'atica Mexicana (2)} 37 (1992), 
no.~1-2, 431--448.

\bibitem[1968]{SegCSS}
Graeme Segal, 
Classifying spaces and spectral sequences, 
\emph{Institut des Hautes \'Etudes Scientifiques Publications Math\'ematiques}
34 (1968), 105--112.

\bibitem[1974]{SegCCT}
Graeme Segal, 
Categories and cohomology theories,
\emph{Topology} 13 (1974), 293--312.

\bibitem[1596]{Sha}
William Shakespeare, 
\emph{King John},
1596.

\bibitem[1994]{Shum}
Mei Chee Shum, 
Tortile tensor categories,
\emph{Journal of Pure and Applied Algebra} 93 (1994), no.~1, 57--110.

\bibitem[1997]{SimCMS}
Carlos Simpson,
A closed model structure for $n$-categories, internal
\textit{\underline{Hom}}, $n$-stacks and generalized Seifert-Van Kampen,  
e-print \url{alg-geom\slsh 9704006}, 1997.

\bibitem[1998]{SimOBB}
Carlos Simpson,
On the Breen--Baez--Dolan stabilization hypothesis for Tamsamani's weak
$n$-categories,  
e-print \url{math\dt CT\slsh 9810058}, 1998.

\bibitem[1999a]{SnyEBG}
Craig. T.~Snydal, 
Equivalence of Borcherds $G$-vertex algebras and axiomatic vertex algebras, 
e-print \url{math\dt QA\slsh 9904104}, 1999.

\bibitem[1999b]{SnyRMC}
Craig T.~Snydal, 
Relaxed multicategory structure of a global category of rings and modules,
e-print \url{math\dt CT\slsh 9912075}, 1999;
also 
\emph{Journal of Pure and Applied Algebra} 168 (2002), no.~2-3, 407--423.

\bibitem[1997]{SoiMTC}
Yan Soibelman,
Meromorphic tensor categories,
e-print \url{q-alg\slsh 9709030}, 1997.

\bibitem[1999]{SoiMBC}
Yan Soibelman, 
The meromorphic braided category arising in quantum affine
algebras,
e-print \url{math\dt QA\slsh 9901003}, 1999;
also
\emph{International Mathematics Research Notices} 1999, no.~19, 1067--1079.

\bibitem[1998]{SB}
Alan D.~Sokal, Jean Bricmont, 
\emph{Fashionable Nonsense},
Picador, 1998.

\bibitem[1963a]{StaHAHI}
James Dillon Stasheff,
Homotopy associativity of $H$-spaces I,
\emph{Transactions of the AMS} 108 (1963), 275--292.

\bibitem[1963b]{StaHAHII}
James Dillon Stasheff,
Homotopy associativity of $H$-spaces II,
\emph{Transactions of the AMS} 108 (1963), 293--312.

\bibitem[1972]{StrFTM}
Ross Street,
The formal theory of monads,
\emph{Journal of Pure and Applied Algebra} 2 (1972), 149--168.

\bibitem[1976]{StrLIC}
Ross Street, 
Limits indexed by category-valued 2-functors,
\emph{Journal of Pure and Applied Algebra} 8 (1976), no.~2, 149--181.

\bibitem[1987]{StrAOS}
Ross Street,
The algebra of oriented simplexes,
\emph{Journal of Pure and Applied Algebra} 49 (1987), no.~3, 283--335.

\bibitem[1988]{StrFN}
Ross Street,
Fillers for nerves,
in
\emph{Categorical Algebra and its Applications (Louvain-la-Neuve, 1987)},
Lecture Notes in Mathematics~1348,
Springer, 1988.

\bibitem[1996]{StrCS}
Ross Street,
Categorical structures, 
in
M.~Hazewinkel (ed.), 
\emph{Handbook of Algebra Vol.~1},
North-Holland, 1996.

\bibitem[1998]{StrRMB}
Ross Street,
The role of Michael Batanin's monoidal globular categories,
in
\emph{Higher Category Theory (Evanston, IL, 1997)},
Contemporary Mathematics~230,
AMS, 1998.

\bibitem[2003]{StrWOC}
Ross Street,
Weak $\omega$-categories,
in 
\emph{Diagrammatic Morphisms and Applications}, 
Contemporary Mathematics~318,
AMS, 2003.

\bibitem[1975]{Sza}
M.E.~Szabo, 
Polycategories,
\emph{Communications in Algebra} 3 (1975), no.~8, 663--689.

\bibitem[1995]{TamSNN}
Zouhair Tamsamani,
Sur des notions de $n$-cat\'egorie et $n$-groupo\"{\i}de non strictes via des
ensembles multi-simpliciaux,  
e-print \url{alg-geom\slsh 9512006}, 1995;
also
\emph{$K$-Theory} 16 (1999), no.~1, 51--99.

\bibitem[1998]{Till}
Ulrike Tillmann,
$\mathcal{S}$-structures for $k$-linear categories and the definition of a
modular functor,
e-print \url{math\dt GT\slsh 9802089}, 1998; 
also
\emph{Journal of the London Mathematical Society (2)} 58 (1998), no.~1,
208--228.

\bibitem[2002]{TV}
Bertrand To\"en, Gabriele Vezzosi, 
Segal topoi and stacks over Segal categories,
e-print \url{math\dt AG\slsh 0212330}, 2002.

\bibitem[1963]{Von}
Kurt Vonnegut, 
\emph{Cat's Cradle},
1963.

\bibitem[1998]{VorSCO}
Alexander A.~Voronov, 
The Swiss-cheese operad,
e-print \url{math\dt QA\slsh 9807037}, 1998;
also in
\emph{Homotopy Invariant Algebraic Structures (Baltimore, MD, 1998)},
Contemporary Mathematics~239,
AMS, 1999.

\bibitem[1981]{Wal}
R.F.C.~Walters,
Sheaves and Cauchy-complete categories.
\emph{Cahiers de Topologie et G\'eom\'etrie Diff\'erentielle Cat\'egorique} 
22 (1981), no.~3, 283--286.

\bibitem[2001]{Web}
Mark Weber, 
Symmetric operads for globular sets,
Ph.D. thesis, Macquarie University, 2001.

\bibitem[1975]{Wor}
Sol Worth, 
Pictures can't say `ain't',
\emph{Versus: Quaderni di studi semiotici} 12 (1975), 85--105.

\bibitem[2001]{Yet}
David N.~Yetter, 
\emph{Functorial Knot Theory},
Series on Knots and Everything~26,
World Scientific Publishing Co., 2001.



\end{thebibliography}

\normalsize