\begin{thebibliography}{99}
\bibitem{Milligan} Milligan, Spike. `Adolf Hitler: my part in his downfall',
Penguin Books, 1972.
\bibitem{Decker} Decker, M., and W. A. Woyczynski, {\bf Introductory
Statistics and Random Phenomena}, Birkhauser, 1998.
\bibitem{Tolman} Tolman, R. C., {\bf Principles of Statistical Mechanics},
Oxford Univ. Press, 1938.
\bibitem{Kolmogorov} Kolmogorov, A. N., {\bf Grundbegriffe der
Wahrscheinlichkeitsrechnung}, Springer-Verlag, Berlin, 1933.
\bibitem{Krylov} Krylov, N. S., {\bf Works on the Foundations of
Statistical Physics}, Princeton University Press, 1979.
\bibitem{Ruelle} Ruelle. D. `Characteristic Exponents for a viscous
fluid..', {\em Commun. Math. Phys.}, {\bf 93}, 285-300, 1984.
{\bf Thermodynamic Formalism}, Encyc. of Math., {\bf 5},
Addison-Wesley, 1978.
\bibitem{Birkhoff} Birkhoff, G. D., {\em Univ. Nac. Tucuman Rev.}, Ser.
{\bf A5}, 147-, 1946.
\bibitem{Ando} Ando, T. {\em Linear Algebra and its Appl.}, {\bf 118},
163-248, 1983.
\bibitem{Streater} R. F. Streater, {\bf Statistical Dynamics}, Imperial
College Press, 1995.
\bibitem{Ingarden} Ingarden, R. S., `Information theory and variational
principles in statistical physics', {Bull. Acad. Pol. Sci.}, {\bf 11},
541-547, 1963.
\bibitem{Jaynes} Jaynes, E. T., `Information theory and
statistical mechanics'', {\em Phys. Rev.}, {\bf 106}, 620-630 and
{\bf 108}, 171-190, 1957.
\bibitem{Rao} Rao, C. R. `Information and accuracy attainable in the
estimation of statistical parameters', {\em Bull. Calcutta Math. Soc.},
{\bf 37}, 81-91, 1945.
\bibitem{Fisher} Fisher, R. A., `Theory of statistical estimation',
{\em Proc. Camb. Phil. Soc.}, {\bf 22}, 700-725, 1925.
\bibitem{Ingarden2} Ingarden, R. S., `Information geometry in functional
spaces' {\em Int. J. Engineering Sci.}, {\bf 19}, 1609-1633, 1981.
\bibitem{Amari} Amari, S.-I., {\bf Differential-Geometric Methods in
Statistics}, {\em Lecture Notes in Statistics}, {\bf 28}, Springer-Verlag,
1985.
\bibitem{Pistone} Pistone, G. and C. Sempi, {\em Annals of Stats.}, {\bf 23},
1543-1561, 1995.
\bibitem{Chentsov} Chentsov, N. N., {\bf Statistical Decision and Optimal
Inference}, {\em Nauka}, Moscow, 1972; {\em Trans. Amer Math Soc}, {\bf 53},
1982.
\bibitem{Petz} Hasagawa, H. and D. Petz, `Non-commutative extension of
information geometry, II', pp 109-118 in
{\bf Quantum Communication, Computing
 and Measurement}, Eds. Hirota et al., Plenum Press, N. Y., 1997.
\bibitem{Ray4} Streater, R. F. `Information manifold for relatively
bounded forms', to appear in the commemoration volume for N. N. Bogoliubov,
(ed. A. A. Slavnov), Steklov Institute. Archive: math-ph/9910035.
\bibitem{Ray5} Streater, R. F., `The analytic quantum info manifold',
to appear in {\bf Stochastic processes, physics and geometry: new
interplays}, eds. F. Geszetesy, S. Paycha and H. Holden; Can. Math. Soc.
Archive math-ph/9910036.
\bibitem{Ray6} Grasselli, M., and R. F. Streater, `Quantum info manifold
for epsilon-bounded forms', Archive math-ph/9910031.
\bibitem{Bachelier} Bachelier, L., `Th\'{e}orie de la speculation',
{\em Ann. Sci. Ecole Norm. Sup.}, {\bf 17}, 21-86, 1900. Reprint
edited by Jacques Gabay, Gauthier-Villars, Paris, 1995.
\bibitem{Einstein1} Einstein, A., {\em Annalen der Physik}, {\bf 17}, 549-560,
1905.
\bibitem{Smol} Chandrasekhar, S., M. Kac and R. Smoluchowski, {\bf
Marian Smoluchowski}, Polish Scientific Publ., Warsaw, 1986.
\bibitem{RFS1} Streater, R. F. {\em Jour. Stat. Phys.}, {\bf 88}, 447-469,
1997. Rep. on Math. Phys., {\bf 40}, 557-564, 1997.
\bibitem{Legesgue} H. Lebesgue, `Lecons sur l'integration et la
recherche des fonctions primitive', Gauthier-Villars, Paris, 1904.
\bibitem{Vitale} G. Vitale, Sul problema della mesura dei gruppe di
punti di una retta, Bologna, 1905.
\bibitem{Hausdorff} F. Hausdorff, {\bf Grundz\"{u}ge der Mengenlehre},
W. de Gruyter, Leipzig, 1914. Reprint, Chelsea, N. Y. 1949, 1965.
\bibitem{Williams} Williams, D., {\bf Probability with Martingales},
Cambridge Univ. Press, 1991.
\bibitem{Wiener} Wiener, N., {\em J. Maths. and Phys.}, {\bf 2},
132, 1923.
\bibitem{Segal} Segal, I. E., `Tensor algebras over Hilbert space I',
{\em Trans. Amer. Math. Soc.,} {\bf 81}, 106-143, 1956.
\bibitem{Wiener2} Wiener, N., `The homogeneous chaos', {\em Amer.
J. Math.}, {bf 55}, 897-936, 1938.
\bibitem{Wick} Wick, G. C., `Evaluation of the collision matrix',
{\em Phys. Rev.}, {\bf 80}, 268-, 1950.
\bibitem{Einstein2} Einstein, A., B. Podolsky, and N. Rosen, {\em Phys.
Rev.} {\bf 47}, 777-80, 1935.
\bibitem{Schilpp} Schilpp, P. A., (ed) {\bf Albert Einstein:
Philosopher-Scientist}, Open Court, La Salle Ill., 1969. p. 210.
\bibitem{Landau} Landau, L. J., `On the Violation of Bell's inequality
in quantum theory', {\em Phys. Lett.}, {\bf A120}, 54-56, 1987.
\bibitem{Kochen} Kochen, S., and E. P. Specker, {\em Jour. Math. Mech.}
{\bf 17}, 59-67, 1967.
\bibitem{Garden} Garden, R. {\bf Modern Logic and Quantum Mechanics}, Adam
Hilger, Bristol, 1984.
\bibitem{Birkhoff2} Birkhoff, G. D., and J. von Neumann, `Logic of
 quantum mechanics', {\em Ann. of Math.}, {\bf 37}, 823-843, 1936.
\bibitem{Jauch} Jauch, J. M., {\bf Foundations of Quantum Mechanics},
Addison Wesley, 1968.
\bibitem{Nelson4} Nelson, E., {\bf Dynamical Theories of Brownian Motion},
Princeton Univ. Press, 1967.
\bibitem{Emch} Emch, G. G., {\bf Algebraic Methods in Statistical Mechanics
and Quantum Field Theory}, Wiley, 1972.
\bibitem{Haag} Haag, R., {\bf Local Quantum Physics}, $2^{\rm nd}$ ed.,
Springer-Verlag, 1996
\bibitem{Horuzhy} Horuzhy, S. S., {\bf Introduction to Algebraic Quantum
Field Theory}, Kluwer Academic, 1990.
\bibitem{Holevo} Holevo, A. S. {\bf Probabilistic and Statistical Aspects
of Quantum Theory}, North Holland, 1982.
\bibitem{Ohya} Ohya, M., and D. Petz, {\bf Quantum Entropy and its Use},
Springer-Verlag, Heidelberg, 1993.
\bibitem{Ingarden4} Ingarden, R. S., H. Janyszek, A. Kossakowski and
T. Kawaguchi, {\em Ibid}, {\bf 37}, 105-111, 1982.
\bibitem{Petz2} Petz, D., and G. Toth, Lett. Math. Phys., {\bf 27}, 205-216,
1993.
\bibitem{Gross} Gross, L., `Abstract Wiener spaces',
{\em Proc. $5^{\rm th}$ Berkeley Symp. in math. stat. and prob. theory},
31-42, 1965-66.
\bibitem{Schwartz} Schwartz, L., {\bf Radon Measures on Arbitrary
Toplological Space and Cylindrical Measures}, Oxford Univ. Press, 1973.
\bibitem{Grimmett} Grimmett, G. R., and D. R. Stirzaker, {\bf
Probability and Random Processes}, Oxford, 1982.
\bibitem{Varadhan} Varadhan, S. R. S., {\bf Large Deviations and
Applications}, SIAM, Philadelphia, 1984.
\bibitem{Donsker} Donsker, M. D., and S. R. S. Varadhan, {\em Phys. Rep.},
{\bf 3}, 235-237, 1981.
\bibitem{Lewis} Lewis, J. T. `Large deviation principle in statistical
mechanics'. pp 141-155, LNM {\bf 1325}, Ed. A. Truman and I. M. Davies,
Springer-Verlag, 1988.
\bibitem{Stroock2} Stroock, D. W., {\bf An Introduction to the Theory
of Large Deviations}, Springer-Verlag, 1984.
\bibitem{GelfandV} Gelfand, I. M. and N. Ya. Vilenkin, {\bf Generalised
Functions IV}, Academic Press, 1964.
\bibitem{Stroock} Stroock, D. and S. R. S. Varadhan, {\bf Multidimensional
Diffusion Processes}, Springer-Verlag, 1979.
\bibitem{Williams2} Williams, D., `To begin at the beginning', 1-55 in
{\bf Stochastic Integrals}, Ed. D. Williams, LNM {\bf 851}, Springer-Verlag,
1981.
\bibitem{McShane} McShane, E. J., {\bf Stochastic Calculus and Stochastic
Models}, Academic Press, N. Y., 1974.
\bibitem{Barnett3} Barnett, C., R. F. Streater and I. F. Wilde, `Quantum
stochastic integrals under standing hypotheses', {\em J. of Math.
Anal. and Applications}, {\bf 127}, 181-192, 1987.
\bibitem{Kac} Kac, M., `Some connections between probability theory
and differential and integral equations', {\em Proc. $2^{\rm nd}$ Berkeley
Symp.} J. Neyman (ed.), University of Calif Press, Berkeley, 1951.
\bibitem{Feynman} Feynman, R. P., {\em Rev. Mod. Phys.}, {\bf 20}, 267-,
1948.
\bibitem{Nelson} Nelson, E., `Feynman Integrals and the
Schr\"{o}dinger Equation', {\em Jour. Math. Phys.,} {\bf 5}, 332-343, 1964.
\bibitem{Simon} Simon, B., {\bf Functional Integration and Quantum
Physics}, Academic Press, N. Y., 1979.
\bibitem{HIK} Hudson, R. L., P. D. F. Ion and K. R. Parthasarathy,
`Time-orthogonal unitary dilations...'{\em Commun. Math. Phys.},
{\bf 83}, 261-280, 1982.
\bibitem{Glimm} Glimm, J., and A. M. Jaffe, {\bf Quantum Physics},
Springer-Verlag, Second Ed., 1987.
\bibitem{Glimm2} Glimm, J., and A. M. Jaffe, `Probability applied to
physics', {\em Univ. Arkansas Lect. Notes in Math.}, {\bf 2},
Fayetteville, 1978.
\bibitem{Frohlich} Fr\"{o}hlich, J., `Schwinger functions and their
generating functionals, I', {\em Helv. Phys. Acta}, {\bf 47}, 265-306, 1974.
II, `Markovian and generalized path space measures on ${\cal S}$',
{\em Adv. in Math.}, {\bf 23}, 119-180, 1977.
\bibitem{Guerra} Guerra, F., L. Rosen and B. Simon, `The $P(\varphi)_2$
Euclidean quantum field theory as classical statistical mechanics',
{\em Annals of Math.}, {\bf 101}, 111-259, 1975.
\bibitem{Dyson} Dyson, F. J., `The S-matrix in quantum electrodynamics',
{\em Phys. Rev.}, {\bf 75}, 1736-1755, 1949.
\bibitem{Schwinger} Schwinger, J. {\em Phys. Rev}, {\bf 82}, 664-, 1951.
\bibitem{Jost} Jost, R. {\bf General Theory of Quantized Fields}, Amer.
Math. Soc., Providence, 1965.
\bibitem{SW} Streater, R. F., and A. S. Wightman, {\bf PCT, Spin and Statistics,
and All That}, Benjamin-Cummings, 1964.
\bibitem{Symanzik} Symanzik, K., `Euclidean quantum field theory, pp 152-226
in {\bf Local Quantum Theory}, R. Jost (ed.), Academic Press, 1969.
\bibitem{Minlos} Minlos, R. A. Trudy Mosk. Mat. Obs. {\bf 8}, 497-518, 1959.
\bibitem{Wong} Wong, E., {\em Ann. Math. Stat.}, {\bf 40}, 1625-1634, 1969.
\bibitem{Nelson2} Nelson, E., `The Free Markov Field', {\em J. Functl.
Anal.}, {\bf 12}, 211-227, 1973.
\bibitem{Simon2} Simon, B., {\bf The $P(\varphi)_2$ Euclidean (Quantum)
Field Theory}, Princeton Univ. Press, 1974.
\bibitem{Gross2} Gross, L., `The free Euclidean Proca and electromagnetic
fields', pp 69-82 in: {\bf Functional Integration and its
Applications}, ed. A. M. Arthurs, Oxford, 1975.
\bibitem{Senitzky} Senitzsky, I. R., `Dissipation in Quantum Mechanics:
The harmonic oscillator I,II'. {\em Phys. Rev.}, {\bf 119}, 670-679,
1960; {\em ibid}, {\bf 124}, 642-648, 1961.
\bibitem{RFS5} Streater, R. F., `Damped oscillator with quantum noise',
{\em J. Phys.}, {\bf A15}, 1477-1486, 1982.
\bibitem{Lax} Lax, M., {\em Phys. Rev.}, {\bf 145}, 111-129, 1965.
\bibitem{Hudson3} Hudson, R. L., and K. R. Parthasarathy, `Quantum Ito's
formula and stochastic evolutions', {\em Commun. Math. Phys.}, {\bf 93},
301-303, 1984.
\bibitem{Partha2} Parthasarathy, K. {\bf An Introduction to Quantum
Stochastic Calculus}, Birkh\"{a}user, Basel, 1992.
\bibitem{Ford} Ford, G. W., M. Kac and P. Mazur, {\em J. Math. Phys.},
{\bf 6}, 505-515, 1965.
\bibitem{Lewishb} Lewis, J. T., and L. C. Thomas, `How to make a heat
bath', 97-123 in {\bf Functional Integration and its Applications},
Ed. A. M. Arthurs, Clarendon Press, Oxford, 1975.
\bibitem{HLK} Hasagawa, H., J. R. Klauder and M. Lakshmanan, {\em J. of
Phys.}, {\bf A14}, L123-L128, 1985.
\bibitem{Accardi} Accardi, L., A. Frigerio and J. T. Lewis, `Quantum
stochastic processes', {\em Proc. Res. Inst. Math. Sci.}, {\bf 18},
97-133, 1982.
\bibitem{Ford2} Ford, G. W., `Temperature-dependent Lamb shift of a quantum
oscillator', pp 202-206 in {\bf Quantum Probability and Applications II},
LNM {\bf 1136}, Eds. L. Accardi and W. von Waldenfels, Springer-Verlag, 1985.
\bibitem{ArakiW} Araki, H., and E. J. Woods, {\em Proc. R. I. M. S., Kyoto},
{\bf 2}, No.2, 1966.
\bibitem{Streater2} Streater, R. F. {\em Nuovo Cimento,} {\bf 53}, 487-495,
1968.
\bibitem{RFS3} Streater, R. F. {\em Nuovo Cimento} {\bf 53}, 487-495, 1968.
\bibitem{Guichardet} Guichardet, A., {\em Commun. Math. Phys.}, {\bf 5},
262-, 1967.
\bibitem{Dubin} Dubin, D. A. and R. F. Streater, {\em Nuovo Cimento} {\bf 50},
154-157, 1967.
\bibitem{RFS2} Streater. R. F. `Current commutation relations, continuous
tensor products, and infinitely divisible group representations', pp
247-263 in {\bf Local Quantum Theory}, Ed. R. Jost, Academic Press, 1969.
\bibitem{Araki} Araki, H. `Factorizable representations of current algebras'
{\em Proc. R. I. M. S., Kyoto}, {\bf 5}, 361-422, 1970/71.
\bibitem{Partha} Parthasarathy, K., and Schmidt, K., `Positive definite
kernels, continuous tensor products, and and central limit theorems of
probability theory', {\bf Lecture Notes in Maths.}, {\bf 272}, 1972.
\bibitem{Wigner} Wigner, E. P., {\bf Group Theory and Applications to 
Atomic Spectroscopy}, First Ed. (German), Viewig., Brannschweig, 1931.
\bibitem{Gelfand} Gelfand, I. M., M. I. Graev and A. M. Vershik,
`Representations of the group $SL(2,{\bf R})$, where ${\bf R}$ is a ring
of functions.' {\em Uspehi-Mat-Nauk} {\bf 28}, 83-128, 1973.
\bibitem{Guichardet2} Guichardet, A., {\bf Symmetric Hilbert Space and Related
Topics}, LNM {\bf 261}, Springer-Verlag, 1972.
\bibitem{Erven} Erven, J., and B.-J. Falkowski, {\bf Low Order Cohomology
and Applications}, Lecture Notes in Mathematics, 877, Springer-Verlag, 1981.
\bibitem{RFS4} Streater, R. F. `Infinitely divisible representations of Lie
algebras', {\em Wahrschein. ver. Geb.}, {\bf 19}, 67-80, 1971.
\bibitem{Mathon} Mathon, D., `Infinitely divisible projective
representations of the Lie algebras', {\em Proc. Camb. Phil. Soc.}, {\bf 72},
357-368, 1972.
\bibitem{Mathon2} Mathon, D. and R. F. Streater, `Infinitely divisible
representations of Clifford algebras', {\em Zeits Wahr. verw. Geb.},
{\bf 20}, 308-316, 1971.
\bibitem{Cushen} Cushen, C. D., and R. L. Hudson, `A quantum mechanical
central limit theorem', {\em J. Appl. Prob.}, {\bf 8}, 454-469, 1971.
\bibitem{Hudson7} Hudson, R. L., `A quantum-mechanical central limit
theorem for anti-commuting variables', {\em J. Appl. Prob.}, {\bf 10},
502-509, 1973.
\bibitem{Hegerfeldt2} Hegerfeldt, G. C., `Prime field decompositions and
infinitely divisible states on Borchers' tensor algebra', {\em Commun.
Math. Phys.}, {\bf 45}, 137-151, 1975.
\bibitem{Schurmann} Sch\"{u}rmann, M., `Positive and conditionally positive
linear functionals on coalgebras', 475-492 in: {\em Quantum Probability II},
Eds. L. Accardi and W. von Waldenfels, LNM {\bf 1136}, 1985.
`Infinitely divisible states on cocommutative
bialgebras', Proc. {\em probability Measures on Groups IX}, Oberwohlfach,
1988. {\bf White Noise on Bialgebras}, {\em Lecture Notes in Math.},
{\bf 1544}, Springer-Verlag, 1993.
\bibitem{Arakithesis} Araki, H., Ph. D. Thesis, Princeton, 1960 (unpublished).
\bibitem{Klauder} Klauder, J. R., `Fock space revisited', {\em Jour.
Math. Phys.}, {\bf 11}, 609-630, 1970; `Ultralocal scalar field
models', {\em Commun. Math Phys.}, {\bf 18}, 307-318, 1970.
\bibitem{Goldin} Goldin, G. A., R. Menikoff and D. H. Sharp, {\em J. Math.
Phys.}, {\bf 21}, 650-, 1980.
\bibitem{Meyer} Meyer, P.-A., {\bf Quantum Probability for Probabilists}
Lecture Notes in Mathematics {\bf 1538}, Springer-Verlag, 1991.
\bibitem{Voiculescu} Voiculescu, D., `Addition of certain non-commuting
random variables', {\em J. Functl. Anal.}, {\bf 66}, 323-346, 1986.
\bibitem{Albeverio} Albeverio, S. and Hoegh-Krohn, `Some Markov processes
and Markov fields', pp 497-540, in {\bf Stochastic Integrals},
ed. D. Williams, LNM {\bf 851}, Springer-Verlag, 1981.
\bibitem{Davies} Davies, E. B., {\bf Quantum Theory of Open Systems},
Academic Press, 1976.
\bibitem{Kraus} Kraus, K., `General State Changes in Quantum Theory',
{\em Ann. Phys.}, {\bf 64}, 311-335, 1971.
\bibitem{Gorini} Gorini, V., A. Kossakowski and E. C. G. Sudarshan,
`Completely positive dynamical semigroups of $n$-level systems',
{\em J. Math. Phys.}, {\bf 17}, 821-825, 1976.
\bibitem{Lindblad} Lindblad, G., {\em Commun. Math. Phys.}, {\bf 48},
119-130, 1976.
\bibitem{Landau2} Landau, L. J., and R. F. Streater, `On Birkhoff's
theorem...', {\em Linear Algebra and its applications}, {\bf 193},
107-127, 1993.
\bibitem{Fannes} Fannes, M., and J. Quaegebure, `Infinite divisibility
and central limit theorems for completely positive mappings', {\bf Quantum
Probability and Applications}, LNM 1136, Ed. L. Accardi and W. von
Waldenfels, Springer-Verlag, 1985.
\bibitem{Stinespring} Stinespring, W. F., `Positive functions on
$C^*$-algebras', {\em Proc. Amer. Math. Soc.}, {\bf 6}, 242-247, 1955.
\bibitem{Umegaki} Umegaki, H., `Conditional expectation in an operator
algebra', {\em Tohoku Math. J.}, {\bf 8}, 86-100, 1956.
\bibitem{Barnett} Barnett, C., `Supermartingales on semi-finite von
Neumann algebras', {\em J. Lond. Math. Soc.}, {\bf 24}, 175-181, 1981.
\bibitem{Cuculescu} Cuculescu, I., `Martingales on von Neumann algebras',
{\em J. Multivariate Analysis}, {\bf 1}, 17-27, 1971.
\bibitem{Hudson} Hudson, R. L. and R. F. Streater, `Examples of
Quantum Martingales', {\em Phys. Lett.}, {\bf 85A}, 64-67, 1981.
\bibitem{Wulfsohn} Streater, R. F., and A. Wulfsohn, `Continuous tensor
products of Hilbert spaces and generalised random fields', {\em Nuovo Cimento},
{\bf 57}, 330-339, 1968.
\bibitem{Hudson4} Hudson, R. L. and R. F. Streater, `Noncommutative
martingales and
stochastic integrals in Fock space', pp 216-222 in {\bf Stochastic Processes
in Quantum Theory and Statistical Physics}, {\em Lecture Notes in Physics},
{\bf 173}, eds S. Albeverio, Ph. Combe and M. Sirugue-Collin,
Springer-Verlag, 1981.
\bibitem{Barnett2} Barnett, C., R. F. Streater and I. F. Wilde, `The
Ito-Clifford integral', {\em J. Functl. Anal.}, {\bf 48}, 172-212, 1982.
\bibitem{Segal0} Segal. I. E. `A non-commutative extension of abstract
integration', {\em Ann. of Math.}, {\bf 57}, 401-457, and 595-596, 1953.
\bibitem{Segal2} Segal, I. E., `Tensor algebras over Hilbert space II',
{\em Annals of Math.}, {\bf 63}, 160-175, 1956.
\bibitem{Kunze} R. A. Kunze and I. E. Segal, {\bf Integrals and
Operators}, Springer-Verlag, 1978.
\bibitem{Barnett5} Barnett, C., R. F. Streater and I. F. Wilde,
`The Ito-Cilfford integral II' {\em J. Lond. Math. Soc.}, {\bf 27},
373-384, 1983.
\bibitem{Barnett6} Barnett, C., R. F. Streater and I. F. Wilde, `The
Ito-Clifford integral IV' {\em J. Operator Theory}, {\bf 11}, 255-271, 1984.
\bibitem{Pisier} Pisier, G., and Q. Xu, `Non-commutative martingale
inequalities', {\em Commun. Math. Phys.}, {\bf 189}, 667-698, 1997.
\bibitem{Barnett4} Barnett, C., R. F. Streater and I. F. Wilde, `Quasi-free
quantum stochastic integrals for the {\em CAR} and {\em CCR}',
{\em J. Funct. Anal.}, {\bf 52}, 19-47, 1983.
\bibitem{Lindsay} Lindsay, M. and I. F. Wilde, `On non-Fock boson
stochastic integrals', {\em J. Functl. Anal.}, {\bf 65}, 76-82, 1986.
\bibitem{Meyer2} Meyer, P.-A., Private communication, Oberwohlfach, 1985.
\bibitem{Evans} Evans, M. P., and R. L. Hudson, `Multidimensional quantum
diffusions', LNM, {\bf 1303}, 69-88, Springer-Verlag, 1988.
\bibitem{Smith} Vincent-Smith, G. F., `Unitary quantum stochastic evolutions',
{\em Proc. London Math. Soc.} {\bf 63}, 401-425, 1991.
\bibitem{Belavkin} Belavkin, V. P., `Quantum stochastic positive
evolutions', {\em Commun. Math. Phys.}, {\bf 184}, 533-566, 1997. `On stochastic
generators of completely positive cocycles', {\em Russian J. of Math.
Phys.}, {\bf 3}, 523-528, 1995.
\bibitem{Frigerio} Frigerio, A., `Construction of stationary quantum Markov
processes', pp 207-222 in {\bf Quantum Probability and Applications II},  
LNM {\bf 1136}, (eds. L. Accardi and W. von Waldenfels), Springer-Verlag,
1985.
\bibitem{Hudson2} Hudson, R. L. and R. F. Streater, `Ito's formula is
the chain rule with Wick ordering,' {\em Phys. Lett.}, {\bf 86A}, 277-279,
1981.
\bibitem{Applebaum} Applebaum, D. and R. L. Hudson, `Fermion Ito's formula
and stochastic evolutions', {\em Commun. Math Phys.}, {\bf 96},
473-496, 1984.
\bibitem{Maassen} Maassen, H. `Quantum Markov processes on Fock space
described by integral kernels', {\bf Quantum Probability II}, 361-374, 1985,
Lecture Notes in Mathematics 1136, eds. L. Accardi and W. von Waldenfels,
Springer-Verlag.
\bibitem{Obata} Obata, N., 'Wick product of white noise operators and
quantum stochastic differential equations', {\em J. Math. Soc. Japan},
{\bf 51}, 613-641, 1999.
\bibitem{Talkner} Talkner, P. `Failure of the quantum regression
theorem', {\em Ann. of Phys.}, {\bf 167}, 390-436, 1986.
\bibitem{Lindblad2} Lindblad, G. `Brownian Motion of quantum harmonic
oscillators', {\em J. Math. Phys.}, {\bf 39}, 2763-2780, 1998.
\bibitem{Alicki} Alicki, R. and K. Lendl, {\bf Quantum Dynamical Semigroups
and Applications}, Lecture Notes in Physics, {\bf 286}, Springer-Verlag,
1987.
\bibitem{AlickiM} Alicki, R., and J. Messer, `Nonlinear quantum dynamical 
semigroups for many-body open systems', {\em J. Stat. Phys.}, {\bf 32},
299-312, 1983.
\bibitem{Balian} Balian, R., Y. Alhassid and R. Reinhardt,
`Dissipation in many-body systems: a geometric approach based on
information theory', {\em Physics Reports}, {\bf 131}, 1-146, 1986.
\bibitem{Ingarden3} Ingarden, R. S., Y. Sato, K. Sagura and T. Kawaguchi,
{\em Tensor}, {\bf 33}, 347-, 1979.
\bibitem{RFS7} Streater, R. F. `Onsager symmetry in statistical dynamics',
{\em Open Systems and Information Dynamics}, {\bf 6}, 87-100, 1999.
`The Soret and Dufour Effects in Statistical Dynamics', {\em Proc. Roy.
Soc.}, A {\bf 456}, 205-221, 2000. Archive math-ph/9910043.

\end{thebibliography}


\end{document}
