
\chapter*{Preface}
 
Mathematical analysis is a standard course which introduces students to rigorous reasonings in mathematics, as well as the theories needed for advanced analysis courses. It is a compulsory course for all mathematics majors. It is also strongly recommended for students that major in computer science, physics, data science, financial analysis,  and other areas that require a lot of analytical skills. Some standard textbooks in mathematical analysis include the classical one by  Apostol \cite{Apostol} and Rudin \cite{Rudin}, and the modern one by Bartle \cite{Bartle}, Fitzpatrick \cite{Fitzpatrick}, Abbott \cite{Abbott}, Tao \cite{Tao_1, Tao_2} and Zorich \cite{Zorich_1, Zorich_2}.
 
 This book is the first volume of the textbooks intended for a one-year course in mathematical analysis.   We introduce the fundamental concepts in a pedagogical way. Lots of examples are given to illustrate the theories. 
We assume that students are familiar with the material of calculus such as those in the book \cite{Stewart}. Thus, we do not emphasize on the computation techniques. Emphasis is put on building up  analytical skills through rigorous reasonings. 

Besides calculus, it is also assumed that students have taken introductory courses in discrete mathematics and linear algebra, which covers topics such as logic, sets, functions, vector spaces, inner products, and quadratic forms. Whenever needed, these concepts would be briefly revised. 

In this book, we have defined all the mathematical terms we use carefully. While most of the terms have standard definitions, some of the terms may have definitions  defer from authors to authors. The readers are advised to check the definitions of the terms used in this book when they encounter them. This can be easily done by using the search function provided by any PDF viewer. The readers are also encouraged to fully utilize the hyper-referencing provided.




 \vspace{0.9cm}
~\hfill\bookauthor

 

%-----------------------------------------------------------------------------
% End of preface.tex
%-----------------------------------------------------------------------------
