\chapter*{Acknowledgements}

Many people have given me encouragement and the benefit of their insight,
wisdom, and expertise.  I am especially grateful to John Baez, Jim Borger,
Tony Carbery, Jos\'e Figueroa-O'Farrill, Tobias Fritz, Herbert Gangl, Heiko
Gimperlein, Dan Haydon, Richard Hepworth, Andr\'e Joyal, Joachim Kock,
Barbara Mable, Louise Matthews, Richard Reeve, Emily Roff, Mike Shulman,
Todd Trimble, Simon Willerton, and X\={\i}l\'{\i}ng Zh\={a}ng, as well as
Roger Astley at CUP.

My heartfelt thanks go to Christina%
%
\index{Cobbold, Christina} 
%
Cobbold and Mark%
%
\index{Meckes, Mark} 
% 
Meckes, not only for many thought-provoking mathematical conversations over
the years, but also for taking the time to read drafts of this text, which
their perceptive and knowledgeable comments helped to improve.

I thank all of those involved in the original functional equations seminar
course in Edinburgh for their good humour and friendly participation.
Interactions during those seminars were formative for my understanding.  So
too were many conversations at \emph{The $n$-Category Caf\'e}, a research
blog that has been invaluable for my learning of new mathematics.

I owe a great deal to the Boyd Orr Centre for Population and Ecosystem
Health, an interdisciplinary research centre at the University of Glasgow,
of which I have been happy to have been a member since 2010.  It is a model
of the interdisciplinary spirit: welcoming, collaborative, informal, and
scientifically ambitious.  Warm thanks to all who foster that positive
atmosphere.

This work was supported at different times by an EPSRC Advanced Research
Fellowship, a BBSRC FLIP award (BB/P004210/1), and a Leverhulme Trust
Research Fellowship.  Critical research advances were made during a 2012
programme on the Mathematics of Biodiversity at the Centre
de Recerca Matem\`atica (CRM) in Barcelona, which was supported by the CRM,
a Spanish government grant, and a BBSRC International Workshop Grant; I was
also supported personally there by the Carnegie Trust for the Universities
of Scotland.  Finally, I thank the Department of Mathematics at the
Universitat Aut\`onoma de Barcelona for their hospitality during part of
the writing of this book.

Some parts of Section~\ref{sec:max} first appeared in Leinster and
Meckes~\cite{MDBB}, and are reproduced with the second author's permission.



