
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{\renewcommand{\thechapter}{}\renewcommand{\chaptername}{}
\addtocounter{chapter}{0}
\section*{Published articles derived from this thesis}\markboth{\sl PUBLICATIONS}{\sl PUBLICATIONS}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\addcontentsline{toc}{chapter}{Publications} 

The research included in this thesis has given rise to the following articles:
\vspace{0.5cm}
\begin{itemize}
 \item  \emph{Redistribution of particle and anti-particle entanglement in non-inertial frames}.
    E. Mart\'in-Mart\'inez and I. Fuentes. Phys. Rev. A 83, 052306 (2011)
    \vspace{0.3cm} 
\item   \emph{Using Berry's phase to detect the Unruh effect at lower accelerations}.
    E. Mart\'in-Mart\'inez and I. Fuentes, R. B. Mann. arXiv:1012.2208. Submitted to Phys. Rev. Letters
    \vspace{0.3cm}
\item   \emph{The entangling side of the Unruh-Hawking effect}
    M. Montero and E. Mart\'in-Mart\'inez. arXiv:1011.6540. Submitted to JHEP
\vspace{0.3cm}
\item   \emph{The Unruh effect in quantum information beyond the single-mode approximation}
    D. E. Bruschi, J. Louko, E. Mart\'in-Mart\'inez, A. Dragan and I. Fuentes. Phys. Rev. A, 82, 042332 (2010)
\vspace{0.3cm}
\item   \emph{Quantum entanglement produced in the formation of a black hole}
    E. Mart\'in-Mart\'inez, L. J. Garay and J. Le\'on. Phys. Rev. D, 82, 064028 (2010)
\vspace{0.3cm}
\item   \emph{Unveiling quantum entanglement degradation near a Schwarzschild black hole}
    E. Mart\'in-Mart\'inez, L. J. Garay and J. Le\'on. Phys. Rev. D, 82, 064006 (2010)
\vspace{0.3cm}
\item   \emph{Entanglement of Dirac fields in an expanding spacetime}
    I. Fuentes, R.B. Mann, E. Mart\'in-Mart\'inez and S. Moradi. Phys. Rev. D, 82, 045030 (2010)
\vspace{0.3cm}
\item   \emph{Population bound effects on bosonic correlations in non-inertial frames}
    E. Mart\'in-Mart\'inez and J. Le\'on, Phys. Rev. A, 81, 052305 (2010)
\vspace{0.3cm}
\item   \emph{Quantum correlations through event horizons: Fermionic vs bosonic entanglement}
    E. Mart\'in-Mart\'inez and J. Le\'on, Phys. Rev. A, 81, 032320 (2010)
\vspace{0.3cm}
\item   \emph{Fermionic entanglement that survives a black hole}
    E. Mart\'in-Mart\'inez and J. Le\'on, Phys. Rev. A, 80, 042318 (2009)
\vspace{0.3cm}
\item   \emph{Spin and occupation number entanglement of Dirac fields for non-inertial observers}
    J. Le\'on and E. Mart\'in-Mart\'inez, Phys. Rev. A, 80, 012314 (2009)
\end{itemize}

Also, during the development of this thesis another article  has been published by its author

\begin{itemize}
\item   \emph{Physical qubits from charged particles: IR divergences in quantum information}.
    J. Le\'on and E. Mart\'in-Mart\'inez, Phys. Rev. A, 79, 052309 (2009)
\end{itemize}



\cleardoublepage
