\begin{thebibliography}{99}

%
% Chap_intro 
%

%\cite{Maldacena:1997re}
\bibitem{Maldacena:1997re2}
  J.~M.~Maldacena,
  ``The Large N limit of superconformal field theories and supergravity,''
  Adv.\ Theor.\ Math.\ Phys.\  {\bf 2 } (1998)  231
  [hep-th/9711200].

\bibitem{MTW}
%no database entry
C.~W.~Misner, K.~S.~Thorne and J.~A.~Wheeler, 
\textit{Gravitation} 
(W. H. Freeman, 
%San Francisco, 
1973).
%［邦訳：重力理論（丸善，2011）］．

%\cite{Aharony:2000ti}
\bibitem{Aharony:2000ti}
O.~Aharony, S.~S.~Gubser, J.~Maldacena, H.~Ooguri and Y.~Oz,
``Large N field theories, string theory and gravity,''
Phys.\ Rept.\  {\bf 323} (2000) 183
[hep-th/9905111].
%%CITATION = HEP-TH 9905111;%%

%\cite{Green:1987sp}
\bibitem{Green:1987sp}
  M.~B.~Green, J.~H.~Schwarz and E.~Witten,
  \textit{Superstring theory}
  (Cambridge Univ.\ Press, 1987).
    
%\cite{Polchinski:1998rq}
\bibitem{Polchinski:1998rq}
  J.~Polchinski,
  \textit{String theory}
  (Cambridge Univ.\ Press, 1998).

%\cite{Zwiebach:2004tj}
\bibitem{Zwiebach:2004tj}
  B.~Zwiebach,
  \textit{A first course in string theory, second edition}
  (Cambridge Univ.\ Press, 2009).

%\cite{Becker:2007zj}
\bibitem{Becker:2007zj}
  K.~Becker, M.~Becker and J.~H.~Schwarz,
  \textit{String theory and M-theory: A modern introduction}
  (Cambridge Univ.\ Press, 2007).

%\cite{Kiritsis:2007zz}
\bibitem{Kiritsis:2007zz}
  E.~Kiritsis,
  \textit{String theory in a nutshell}
  (Princeton Univ.\ Press, 2007).
  
%\cite{Johnson:2003gi}
\bibitem{Johnson:2003gi}
  C.~V.~Johnson,
  \textit{D-branes}
  (Cambridge Univ.\ Press, 2003).   

%\cite{Schutz:1985jx}
\bibitem{Schutz:1985jx}
B.~F.~Schutz,
\textit{A first course in general relativity, second edition}
(Cambridge Univ.\ Press, 2009).
%［邦訳：バーナード・Ｆ・シュッツ，相対論入門 第2版（丸善，2010）］．

%\cite{Zee:2013dea}
\bibitem{Zee:2013dea}
A.~Zee,
\textit{Einstein Gravity in a Nutshell}
(Princeton Univ.\ Press, 2013).

%\cite{Wald:1984rg}
\bibitem{Wald:1984rg}
R.~M.~Wald,
\textit{General relativity}
(The Univ.\ of Chicago Press, 1984).

% added
%\cite{Ryder:1985wq}
\bibitem{Ryder:1985wq}
  L.~H.~Ryder,
  \textit{Quantum Field Theory, second edition} 
  (Cambridge Univ.\ Press, 1996).
  
% added
%\cite{Zee:2003mt}
\bibitem{Zee:2003mt}
  A.~Zee,
  \textit{Quantum field theory in a nutshell}
  (Princeton Univ.\ Press, 2010).

% added
%\cite{Srednicki:2007qs}
\bibitem{Srednicki:2007qs}
  M.~Srednicki,
  \textit{Quantum field theory}
  (Cambridge Univ.\ Press, 2007).
  
% added
%\cite{Peskin:1995ev}
\bibitem{Peskin:1995ev}
  M.~E.~Peskin and D.~V.~Schroeder,
  \textit{An Introduction to quantum field theory}
  (Addison-Wesley, 1995).
  
% added
%\cite{Weinberg:1995mt}
\bibitem{Weinberg:1995mt}
  S.~Weinberg,
  \textit{The Quantum theory of fields}, Vol. 1-3
  (Cambridge Univ.\ Press, 1995-2000).
%The Quantum theory of fields. Vol. 1: Foundations,''

%\cite{Weinberg:1996kr}
%\bibitem{Weinberg:1996kr}
%  S.~Weinberg,
%``The quantum theory of fields. Vol. 2: Modern applications,''

%\cite{Weinberg:2000cr}
%\bibitem{Weinberg:2000cr}
%  S.~Weinberg,
  %``The quantum theory of fields. Vol. 3: Supersymmetry,''

%\cite{Papantonopoulos:2011zz}
\bibitem{Papantonopoulos:2011zz}
\textit{From Gravity to Thermal Gauge Theories: The AdS/CFT Correspondence}, 
eds. E. Papantonopoulos, Lecture Notes in Physics 828 (Springer, 2011).

%\cite{Horowitz:2012nnc}
\bibitem{Horowitz:2012nnc}
\textit{Black Holes in Higher Dimensions}, 
eds. G. Horowitz (Cambridge University Press, 2012).

%\cite{DeWolfe:2013cua}
\bibitem{DeWolfe:2013cua}
  O.~DeWolfe, S.~S.~Gubser, C.~Rosen and D.~Teaney,
  ``Heavy ions and string theory,''
  arXiv:1304.7794 [hep-th].
  %%CITATION = ARXIV:1304.7794;%%

% added
%\cite{Hartnoll:2009sz}
\bibitem{Hartnoll:2009sz}
  S.~A.~Hartnoll,
  ``Lectures on holographic methods for condensed matter physics,''
  Class.\ Quant.\ Grav.\  {\bf 26} (2009) 224002
  [arXiv:0903.3246 [hep-th]].
  %%CITATION = ARXIV:0903.3246;%%
  
% added
%\cite{McGreevy:2009xe}
\bibitem{McGreevy:2009xe}
  J.~McGreevy,
  ``Holographic duality with a view toward many-body physics,''
  Adv.\ High Energy Phys.\  {\bf 2010} (2010) 723105
  [arXiv:0909.0518 [hep-th]].
  %%CITATION = ARXIV:0909.0518;%%

%\cite{Iqbal:2011ae}
\bibitem{Iqbal:2011ae}
  N.~Iqbal, H.~Liu and M.~Mezei,
  ``Lectures on holographic non-Fermi liquids and quantum phase transitions,''
  arXiv:1110.3814 [hep-th].
  %%CITATION = ARXIV:1110.3814;%%

%\cite{CasalderreySolana:2011us}
\bibitem{CasalderreySolana:2011us}
  J.~Casalderrey-Solana, H.~Liu, D.~Mateos, K.~Rajagopal and U.~A.~Wiedemann,
  \textit{Gauge/String Duality, Hot QCD and Heavy Ion Collisions} (Cambridge Univ.\ Press, 2014)
  [arXiv:1101.0618 [hep-th]].
  %%CITATION = ARXIV:1101.0618;%%

% added
%\cite{Ammon:2015wua}
\bibitem{Ammon:2015wua}
  M.~Ammon and J.~Erdmenger,
  \textit{Gauge/gravity duality : Foundations and applications}
  (Cambridge Univ.\ Press, 2015).
  %%CITATION = INSPIRE-1376202;%%

% added
%\cite{Zaanen:2015oix}
\bibitem{Zaanen:2015oix}
  J.~Zaanen, Y.~W.~Sun, Y.~Liu and K.~Schalm,
  \textit{Holographic Duality in Condensed Matter Physics}
  (Cambridge Univ.\ Press, 2015).
  %%CITATION = INSPIRE-1384852;%%

\end{thebibliography}
